### 2018-05-09 Backup.tex
Initial ShareLaTeX Import

\documentclass[a4paper,11pt]{article}
\pdfoutput=1 % if your are submitting a pdflatex (i.e. if you have
             % images in pdf, png or jpg format)

\usepackage{jcappub} % for details on the use of the package, please
                     % see the JCAP-author-manual

\usepackage[T1]{fontenc} % if needed
\usepackage{eucal}
\usepackage{float}

\title{\boldmath PyUltraLight: A Pseudo-Spectral Code for the Dynamics of Ultra-Light Dark Matter}

%% %simple case: 2 authors, same institution
%% \author{A. Uthor}
%% \author{and A. Nother Author}
%% \affiliation{Institution,\\Address, Country}

% more complex case: 4 authors, 3 institutions, 2 footnotes
\author[a,b,1]{F. Irst,\note{Corresponding author.}}
\author[c]{S. Econd,}
\author[a,2]{T. Hird\note{Also at Some University.}}
\author[a,2]{and Fourth}

% The "\note" macro will give a warning: "Ignoring empty anchor..."
% you can safely ignore it.

\affiliation[a]{One University,\\some-street, Country}
\affiliation[b]{Another University,\\different-address, Country}
\affiliation[c]{A School for Advanced Studies,\\some-location, Country}

% e-mail addresses: one for each author, in the same order as the authors
\emailAdd{first@one.univ}
\emailAdd{second@asas.edu}
\emailAdd{third@one.univ}
\emailAdd{fourth@one.univ}

\abstract{P\textsc{y}U\textsc{ltra}L\textsc{ight} is a Python based program that solves the Schr{\"o}dinger-Poisson coupled differential equations, which govern the dynamics of ultra-light scalar field dark matter in the non-relativistic regime. P\textsc{y}U\textsc{ltra}L\textsc{ight} utilises a symmetrised split-step pseudo-spectral Fourier algorithm to solve these differential equations, and demonstrates excellent energy conservation. Being Python-based, the program is particularly easy to set up, and the accompanying Jupyter notebook provides a simple interface in which to augment simulation parameters and create visualisations of output data. Using P\textsc{y}U\textsc{ltra}L\textsc{ight}, we are able to reproduce many previously studied dynamical characteristics of ultra-light dark matter, even when running the simulations at relatively low spatial resolution from a standard user-PC.}

\begin{document}
\maketitle
\flushbottom

\section{Introduction}
\label{sec:intro}

The true nature of dark matter (DM) is one of the most important outstanding problems in theoretical physics. WIMP models of dark matter are favoured by many, due in part to their consistency with supersymmetric extensions to the standard model of particle physics. However, with increasingly advanced direct-detection experiments failing to find evidence for WIMP candidates, the viable parameter space continues to shrink \cite{Tan2016, Akerib2017}. Furthermore, the $\Lambda$CDM cosmologies consistent with WIMP models suffer from a number of widely-publicised small-scale problems \cite{Bull2016}. While baryonic physics is proposed to offer solutions to a number of these problems, the perceived issues within $\Lambda$CDM are increasingly motivating novel research into alternative DM models, one such class of model being ultra-light dark matter (ULDM), also known as fuzzy dark matter (FDM), or BEC dark matter. ULDM models are well motivated by fundamental theories possessing approximate shift symmetries, and offer natural resolutions to the small-scale problems of $\Lambda$CDM without the need to invoke baryonic physics. This is essentially because below the de Broglie wavelength of the ULDM particle, the Heisenberg uncertainty principle suppresses gravitational collapse. Since the mass of the ULDM particle is postulated to be of order $10^-22$ eV, this amounts to suppression of gravitational collapse on kilo parsec scales \cite{Hu2000}. A recent review of ULDM/FDM is given in \cite{Hui2016}.

The computational modelling of ULDM systems in the non-relativistic regime can be distilled into numerically solving the dynamical Schr{\"o}dinger-Poisson (SP) system of coupled differential equations for given initial conditions. Approaches to this problem have been varied, and have given rise to modifications of existing cosmological simulation codes as well as the development of new codes specifically for the modelling of ULDM systems, though public releases of the latter are  lacking. One approach has been to use the Madelung fluid formulation of the SP system \cite{Madelung1926}, in which there arises a quantum pressure term which can be treated numerically in a variety of ways. In \cite{Zhang2018}, the cosmological code \textsc{gadget} \cite{Springel2005} has been modified in order to treat this quantum pressure using an effective particle-particle interaction. This modified code, \textsc{axion-gadget}, has been made publicly available \cite{axion-gadget}. In \cite{Nori2018}, a non-public extension of \textsc{gadget}, \textsc{p-gadget3}, has been modified to treat the quantum pressure term using smoothed-particle hydrodynamics (SPH) routines. The SPH approach has also been used in \cite{Mocz2015}, while a particle-mesh approach was implemented in \cite{Veltmaat2016} to treat the quantum pressure term. Other cosmological simulation codes which have been recruited to solve ULDM dynamics (though not necessarily using the Madelung formulation) are N\textsc{yx} \cite{Almgren2013}, G\textsc{alacticus} \cite{Benson2012}, \textsc{arepo} \cite{Springel2010}, and \textsc{gamer} \cite{Schive2010, gamer}. N\textsc{yx} was modified in \cite{Schwabe2016} to study merging ULDM solitonic cores, G\textsc{alacticus} was modified in \cite{Du2017} to study the effects of tidal stripping and dynamical friction on ULDM halos, \textsc{arepo} was modified in \cite{Mocz2017} to study the core-mass relationship and turbulence characteristics of ULDM halos, and \textsc{gamer} was modified in \cite{Schive2014_b} to perform a detailed study of structure formation in ULDM cosmologies.

While these modified cosmological simulation codes have produced promising results in the study of ULDM physics, at the time of writing only one such code, \textsc{axion-gadget}, has been made publicly available. The aim of this paper is to introduce P\textsc{y}U\textbf{ltra}L\textsc{ight}, a stand-alone Python-based pseudo-spectral SP solver, and to demonstrate that this code is able to reproduce a number of key findings from the more complicated cosmological simulation codes without the need for high-performance computing environments. Consequently, we anticipate that as a publicly available resource, P\textsc{y}U\textbf{ltra}L\textsc{ight} will serve as a valuable cross-check for more complex implementations and may serve as a basis for further development of such codes within the computational cosmology community.

P\textsc{y}U\textbf{ltra}L\textsc{ight} makes use of a symmetrised-split-step pseudo-spectral Fourier algorithm (also known as leap-frog) to compute the dynamics of the SP system. We note that a similar methodology has been described in \cite{Paredes2016}, though to the authors' knowledge this code has not been released publicly. The symmetrised-split-step method offers $2^{nd}$ order accurate time integration steps and sub-percent level energy conservation while normalisation of the wavefunction is conserved to machine precision. As a pseudo-spectral code, linear differential operators are computed by direct multiplication in the Fourier domain, while non-linear terms are evaluated in position space. This means that P\textsc{y}U\textbf{ltra}L\textsc{ight} is free from noise associated with finite-differencing methods of computing spatial derivatives, though there is a necessary computational cost associated with the Fourier and inverse Fourier transforms. These transforms are optimised in P\textsc{y}U\textbf{ltra}L\textsc{ight} through the use of the pyFFTW pythonic wrapper around the C-based FFWTW subroutine library \cite{pyfftw} \cite{fftw}. While the code is currently implemented to run on a singe machine, the FFTW libraries offer full parallelisation, so there is no barrier to running the code in a cluster environment.

This paper is organised as follows. We first provide a short review of the physics of ULDM, including a derivation of the SP equations from an appropriate scalar-field Lagrangian. We then describe the details of the implementation of P\textsc{y}U\textbf{ltra}L\textsc{ight}, before moving on to a testing and verification section, in which we reproduce a selection of results from a variety of recent ULDM simulations and discuss energy conservation within the code.

\section{The Physics of ULDM: The Schr{\"o}dinger-Poisson System}\label{sec:physics}

The existence of an extremely light scalar field, minimally coupled to gravity, is the central premise on which ULDM models are predicated. Within the ULDM framework, it is proposed that large-scale cosmological structure formation proceeds through the Bose-Einstein condensation of this scalar field. In this section we derive the (non-relativistic) dynamical equations of the single macroscopic condensate wavefunction, namely, the Schr{\"o}dinger-Poisson equations. We begin by considering the following action functional for a scalar field, $\phi$, minimally coupled to gravity and in the absence of self-interactions:
\begin{equation}\label{eq:action}
    S=\int \frac{d^4x}{\hbar}\sqrt{-g}\bigg\{\frac{1}{2}g^{\mu\nu}\partial_\mu\phi\partial_\nu\phi-\frac{m^2}{2\hbar^2}\phi^2\bigg\},
\end{equation}
where we have taken $c=1$ but retain factors of $\hbar$ at this stage. Applying the variational principle to this action yields the Euler-Lagrange equations
\begin{equation}\label{eq:e-l}
    \frac{1}{\sqrt{-g}}\partial_\mu\big[\sqrt{-g}g^{\mu\nu}\partial_\nu\phi\big]-\frac{m^2}{\hbar^2}\phi=0.
\end{equation}
We evaluate equation \ref{eq:e-l} using linear perturbation theory, adopting the perturbed FRW metric in the Newtonian gauge:
\begin{equation}\label{eq:pFRW}
    ds^2=-\big(1+2\Phi(\vec{r})\big)dt^2+a(t)^2\big(1-2\Phi(\vec{r})\big)d\vec{r}^{\,2}.
\end{equation}
To linear order in $\Phi(\vec{r})$ we obtain 
\begin{equation}\label{eq:linear}
    \Ddot{\phi}-\frac{\big(1+4\Phi(\vec{r})\big)}{a(t)^2}\nabla^2\phi+3\big(1+2\Phi(\vec{r})\big)H\Dot{\phi}+\big(1+2\Phi(\vec{r})\big)\frac{m^2}{\hbar^2}\phi=0,
\end{equation}
where $H=\Dot{a}(t)/a(t)$. At late times, $H\ll m$ and it is sufficient to set $H=0$ and $a(t)=1$ in equation \ref{eq:linear}. In the non-relativistic limit, we may use the WKB approximation to obtain an ansatz solution to equation \ref{eq:linear}. We take 
\begin{equation}\label{eq:ansatz}
    \phi=\frac{\hbar}{\sqrt{2}m}\big(\psi e^{-imt/\hbar}+\psi^* e^{imt/\hbar}\big),
\end{equation}
where $\psi$ is a slowly varying complex function so that terms involving $\Ddot{\psi}$ may be neglected. We may also use $\nabla^2\phi\ll m^2/\hbar^2\phi$, such that our final result takes the form
\begin{equation}\label{eq:schrodinger}
    i\hbar\Dot{\psi}=-\frac{\hbar^2}{2m}\nabla^2\psi+m\Phi\psi.
\end{equation}
Thus, by making suitable approximations we have derived the Schr{\"o}dinger equation for $\psi$, which we now interpret as the unique macroscopic wavefunction of a Bose-Einstein condensate. It follows that the particle density of the condensate is given by $\vert\psi\vert^2$, such that the mass density is simply $m\vert\psi\vert^2$. Using this, we may therefore write down the Poisson equation as
\begin{equation}\label{eq:poisson}
    \nabla^2\Phi=4\pi G m \vert\psi\vert^2,
\end{equation}
where $G$ is Newton's gravitational constant. The coupled equations \ref{eq:schrodinger} and \ref{eq:poisson} together form the nonlinear Schr{\"o}dinger-Poisson (SP) system which describes the dynamics of ULDM in the non-relativistic regime. It is this system of equations which P\textsc{y}U\textsc{ltra}L\textsc{ight} is designed to solve.

\section{Algorithm and Implementation}\label{sec:implementation}

Summary of this section TBW

\subsection{Field Rescalings}

It is helpful to recast the SP system (equations \ref{eq:schrodinger} and \ref{eq:poisson}) in terms of adimensional quantities. In keeping with \cite{Schive2014} and \cite{Paredes2016}, we introduce length, time, and mass scales as follows:
\begin{align}
    \CMcal{L}&=\left(\frac{8\pi\hbar^2}{3 m^2H_0^2\Omega_{m_0}}\right)^{\frac{1}{4}}\approx121\left(\frac{10^{-23}\operatorname{eV}}{m}\right)\operatorname{kpc},\label{eq:length}\\
    \CMcal{T}&=\left(\frac{8\pi}{3 H_0^2\Omega_{m_0}}\right)^{\frac{1}{2}}\approx75.5 \operatorname{Gyr},\label{eq:time}\\
    \CMcal{M}&=\frac{1}{G}\left(\frac{8\pi}{3 H_0^2\Omega_{m_0}}\right)^{-\frac{1}{4}}\left(\frac{\hbar}{m}\right)^{\frac{3}{2}}\approx 7\times 10^7\left(\frac{10^{-23}\operatorname{eV}}{m}\right)\operatorname{M}_{\odot},\label{eq:mass}
\end{align}
where $m$ is the mass of the ultra-light scalar field, $H_0$ is the present-day Hubble parameter, $G$ is Newton's gravitational constant and $\Omega_{m_0}$ is the present-day matter fraction of the energy density of the universe. We now recast equations \ref{eq:schrodinger} and \ref{eq:poisson} in terms of the following dimensionless quantities:
\begin{equation}\label{eq:dimensionless}
    t'=\frac{t}{\CMcal{T}},\quad
    \vec{x}^{\,'}=\frac{\vec{x}}{\CMcal{L}},\quad
    \Phi'=\frac{m\CMcal{T}}{\hbar}\Phi,\quad
    \psi'=\CMcal{T}\sqrt{mG}\psi.
\end{equation}
Dropping the primes for notational convenience, we see that the coupled differential equations of the SP system are reduced to 
\begin{align}
    i\Dot{\psi}(\vec{x},t)&=-\frac{1}{2}\nabla^2\psi(\vec{x},t)+\Phi(\vec{x},t)\psi(\vec{x},t),\label{eq:s-adim}\\
    \nabla^2\Phi(\vec{x},t)&=4\pi\vert\psi(\vec{x},t)\vert^2,\label{eq:p-adim}
\end{align}
where it is understood that all quantities involved are dimensionless. We can recover any dimensionful quantities of interest through simple transformations using equations \ref{eq:length} to \ref{eq:mass}. For example, the integrated mass of the system, $M_{tot}$, is given by
\begin{equation}\label{eq:integrated-mass}
    M_{tot}=\CMcal{M}\int d^3x\vert\psi\vert^2.
\end{equation}
Likewise, the mass density at any point is given by
\begin{equation}\label{eq:density}
    \rho=\frac{\CMcal{M}}{\CMcal{L}^3}\vert\psi\vert^2.
\end{equation}
All computations performed by P\textsc{y}U\textsc{ltra}L\textsc{ight} are carried out using the dimensionless quantities defined in \ref{eq:dimensionless}, while provisions are made within the code for initial conditions and outputs to be converted to and from the desired unit system specified by the user. Henceforth, we will often refer to the quantity $\vert\psi\vert^2$ simply as the density, where it is understood that this is in fact a dimensionless quantity related to the physical mass density through a constant of proportionality given by equation \ref{eq:density}.

\subsection{Dynamical Evolution}\label{sec:dynamics}

Dynamical evolution within P\textsc{y}U\textsc{ltra}L\textsc{ight} progresses via a symmetrised split-step Fourier process on an $N\times N\times N$ simulation grid with spatial periodic boundary conditions. To understand this method, we first consider the exact expression for the unitary time evolution of the wavefunction according to equation \ref{eq:s-adim}, namely
\begin{equation}\label{eq:exact-time-ev}
    \psi(\vec{x},t+h)=\mathcal{T}\operatorname{exp}\left[-i\int_t^{t+h}dt'\left\{-\frac{1}{2}\nabla^2+\Phi(\vec{x},t')\right\}\right]\psi(\vec{x},t),
\end{equation}
where $\mathcal{T}$ is the time-ordering symbol. For sufficiently small timestep $h$, we can use the trapezoidal rule make the approximation
\begin{equation}
    \int_t^{t+h}dt'\Phi(\vec{x},t')\approx \frac{h}{2}\Big(\Phi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)+\Phi(\vec{x},t)\Big).
\end{equation}
We can therefore write the approximate form of equation \ref{eq:exact-time-ev} as
\begin{equation}\label{eq:approx-time-ev}
    \psi(\vec{x},t+h)\approx \operatorname{exp}\left[i\frac{h}{2}\Big(\nabla^2-\Phi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)-\Phi(\vec{x},t)\Big)\right]\psi(\vec{x},t).
\end{equation}
It is acknowledged that the exponential in equation \ref{eq:approx-time-ev} omits the time-ordering symbol, and is only equivalent to its time-ordered counterpart to order $h^2$.

Given that the linear differential operator in equation \ref{eq:approx-time-ev} acts naturally in Fourier space, while the nonlinear potential term is simplest to evaluate in position space, we are motivated to split the exponential so as to separate these terms. This will then allow us to employ a pseudo-spectral algorithm to compute the time evolution. Such a splitting of the exponential is valid when the timestep $h$ is small, as the dispersive and nonlinear terms can be assumed to act independently. In this implementation, we split the exponential as follows:
\begin{equation}\label{eq:split-step}
    \psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)\approx\operatorname{exp}\left[-\frac{ih}{2}\Phi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)\right]\operatorname{exp}\left[\frac{ih}{2}\nabla^2\right]\operatorname{exp}\left[-\frac{ih}{2}\Phi(\vec{x},t)\right]\psi(\vec{x},t).
\end{equation}
This splitting can be understood thusly: first, a half timestep is taken in which only the nonlinear potential operator acts, followed by a full timestep in the linear term. The potential field is then updated, and a final half timestep in the nonlinear term is performed. We may use the Baker-Campbell-Hausdorff formula to express the product of exponentials in equation \ref{eq:split-step} as a single exponential. Keeping only terms to order $h^2$ we find:
\begin{equation}\label{eq:BCH}
    \operatorname{exp}\left[i\frac{h}{2}\Big(\nabla^2-\Phi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)-\Phi(\vec{x},t)\Big)+\frac{h^2}{8}\Big[\nabla^2,\Phi(\vec{x},t)\Big]-\frac{h^2}{8}\Big[\nabla^2,\Phi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)\Big]\right].
\end{equation}
Making use of the fact that $\Phi(\vec{x},t+h)\approx\Phi(\vec{x},t)+h\Dot{\Phi}(\vec{x},t)$, we see that the commutators in equation \ref{eq:BCH} cancel at $\mathcal{O}(h^2)$, such that this expression is equivalent to \ref{eq:approx-time-ev}, with the dominant error term appearing at $\mathcal{O}(h^3)$.

Evaluation of equation \ref{eq:split-step} within P\textsc{y}U\textsc{ltra}L\textsc{ight} proceeds as follows:
Initially, the nonlinear term acts in position space for one half-timestep. The result is then Fourier transformed, and a full timestep is taken in which the differential operator is applied in the Fourier domain. The potential field is then updated in accordance with equation \ref{eq:p-adim}. An inverse Fourier transform is then performed, and a final half timestep is taken in which the updated nonlinear term acts in position space to give the new $\psi$ field configuration. This procedure is known as the symmetrised split-step Fourier method, and is used widely in fields such as nonlinear fiber optics \cite{Agrawal2013}. It can be represented schematically as follows:
\begin{equation}\label{eq:schematic-psi}
    \psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)=\operatorname{exp}\left[-\frac{ih}{2}\Phi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)\right]\CMcal{F}^{-1} \operatorname{exp}\left[\frac{-ih}{2}k^2\right]\CMcal{F}\operatorname{exp}\left[-\frac{ih}{2}\Phi(\vec{x},t)\right]\psi(\vec{x},t),
\end{equation}
where the order of operations is read from right to left, $\CMcal{F}$ and $\CMcal{F}^{-1}$ represent the discrete Fourier transform and its inverse, and $k$ is the wavenumber in the Fourier domain. The potential field is updated following the inverse Fourier transform in equation \ref{eq:schematic-psi}, and this update procedure may be represented schematically as
\begin{equation}\label{eq:schematic-phi}
    \Phi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)=\CMcal{F}^{-1}\left(-\frac{1}{k^2}\right)\CMcal{F}\ 4\pi\vert\psi(\vec{x},t_{i})\vert^2,
\end{equation}
where $\psi(\vec{x},t_{i})$ represents the field configuration at this intermediary point in the timestep. The final operation in equation \ref{eq:schematic-psi} changes only the phase of the $\psi$ field, and thus we could replace $\psi(\vec{x},t_{i})$ with $\psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)$ in equation \ref{eq:schematic-phi} with no change in meaning.

In the P\textsc{y}U\textsc{ltra}L\textsc{ight} implementation, we make use of an additional simplification of the symmetrised split-step Fourier method. Namely, consecutive half-steps in the nonlinear term may be combined into a single full step in each case such that when iterated, only the first and last operations involve half steps. This can be represented schematically as 
\begin{equation}
    \psi(t\hspace*{-.2em}+\hspace*{-.2em}nh)=\operatorname{exp}\left[+\frac{ih}{2}\Phi\right]\left(\prod^n\operatorname{exp}\left[-\frac{ih}{2}\Phi\right]\operatorname{exp}\left[\frac{ih}{2}\nabla^2\right]\right)\operatorname{exp}\left[-\frac{ih}{2}\Phi\right]\psi(t),
\end{equation}
where the $\Phi$ field is updated at each step in accordance with equation \ref{eq:schematic-phi} and attention is drawn to the sign difference between the first and last operators.

The discrete Fourier transform (DFT) and its inverse are handled within P\textsc{y}U\textsc{ltra}L\textsc{ight} using pyFFTW. pyFFTW is a pythonic wrapper around the C-based FFTW subroutine library, and allows for efficient evaluation of both real and complex DFTs \cite{pyfftw} \cite{fftw} \cite{Frigo2005}.  Multithreading is supported by pyFFTW, and is implemented within P\textsc{y}U\textsc{ltra}L\textsc{ight} by using the multiprocessing package to determine the number of online CPUs and feed this directly to the threads argument of the pyfftw.FFTW class.

In addition to the use of pyFFTW, performance within P\textsc{y}U\textsc{ltra}L\textsc{ight} is further improved through the use of the NumExpr package, which provides routines for the efficient manipulation of array objects within the simulation using a vector-based virtual machine \cite{numexpr}.

\subsection{Initial Conditions: Soliton Profiles}\label{sec:soliton-profiles}

The initialisation procedure within P\textsc{y}U\textsc{ltra}L\textsc{ight} can be easily augmented to accommodate the requirements of the user. In the current implementation, the initial field configuration is built by loading a NumPy array file encoding a solitonic solution to the Schr{\"o}dinger-Poisson system. Using this method, solitons may be placed at the desired locations within the simulation grid, with the corresponding mass, velocity, and phase parameters each specified by the user within the accompanying Jupyter notebook.

The solitonic solution to the SP system is a natural choice of initial field configuration; these stable, localised objects are ubiquitous in ULDM models \cite{Marsh2015}.

%In particular, it is proposed that ULDM halos exhibit Navarro-Frenk-White profiles at large radii, with solitonic cores at smaller radii. Because solitons have a constant central density, it is suggested that this offers a solution to the core-cusp problem of $\Lambda$CDM \cite{Marsh2015}.

%Remove this section if we're going to talk about solitons merging to form NFW/soltion profiles, and compare to e.g. Schwabe.

In this section, we discuss the method used to generate the soliton profile. Following \cite{Paredes2016}, we begin by assuming a spherically symmetric solution to the SP system:
\begin{equation}
    \psi(\vec{x},t)\rightarrow e^{i\beta t} f(r), \quad \Phi(\vec{x},t)\rightarrow \varphi(r),
\end{equation}
where $r=\vert\vec{x}\vert$. Introducing $\tilde{\varphi}(r):=\varphi(r)+\beta$, equations \ref{eq:s-adim} and \ref{eq:p-adim} are reduced to
\begin{align}
    0 &= -\frac{1}{2}f^{\prime\prime}(r)-\frac{1}{r}f^\prime(r)+\tilde{\varphi}(r)f(r)\label{eq:s-spherical}\\
    0 &= \tilde{\varphi}^{\prime\prime}(r)+\frac{2}{r}\tilde{\varphi}^\prime(r)-4\pi f(r)^2\label{eq:p-spherical}
\end{align}
where primes denote derivatives with respect to $r$. To generate profiles for $\tilde{\varphi}(r)$ and $f(r)$ which satisfy these equations, we use a fourth-order Runge-Kutta algorithm, included as the auxiliary soliton\texttt{\_}solution.py program. We make the arbitrary choice $f(r=0)$=1, while the smoothness criterion demands that the first derivatives of $f(r)$ and $\tilde{\varphi}(r)$ vanish at the origin. We then search for solutions of $f(r)$ and $\varphi(r)$ which satisfy the boundary conditions $\operatorname{lim}(r\rightarrow\infty)\varphi(r)=0$ and $\operatorname{lim}(r\rightarrow\infty)f(r)=0$. To do this we use a shooting method, varying $\tilde{\varphi}(r=0)$ until we obtain a solution of $f(r)$ which approaches zero at the maximal specified radius, $r_m$. The value of $\beta$ is then calculated by assuming that $\varphi(r)$ goes as $-c/r$ at large radii, where $c$ is a constant. Under this assumption, we can write
\begin{equation}
    \tilde{\varphi}(r_m)=-\frac{c}{r_m}+\beta, \quad c=r_m^2\tilde{\varphi}^{\ \prime}(r_m).
\end{equation}
We thus obtain the full solution $\psi(\vec{x},t)=e^{i\beta t}f(r)$.
Having initially chosen $f(r=0)=1$, we may then generalise to $f(r=0)=\alpha$, where $\alpha$ is an arbitrary positive real number. It is easy to verify that if $e^{i\beta t}f(r)$ is a solution to the spherically symmetric SP system, then $g(r)$ is also a solution, where
\begin{equation}
    g(r)=e^{i\alpha\beta t}\alpha f(\sqrt{\alpha}\ r).
\end{equation}
Hence we obtain a family of spherically symmetric soliton solutions to the dimensionless SP system, for which the dimensionless soliton mass is proportional to $\sqrt{\alpha}$ and the full width at half maximum is proportional to $1/\sqrt{\alpha}$. Because the size of the soliton scales inversely with the mass, it is important to consider the maximum soliton mass which can be accurately modeled in a simulation with finite spatial resolution.

While the Schr{\"o}dinger equation is not trivially form invariant under Galilean boosts, we can enforce Galilean covariance by augmenting the solutions through the addition of a velocity-dependent phase factor. Namely,
\begin{equation}\label{eq:covariant-solution}
    \psi(\vec{x},t)=\alpha f\big(\sqrt{\alpha}\vert\vec{x}-\vec{v}t\vert\big)e^{i\left(\alpha\beta t+\vec{v}\cdot\vec{x}-\frac{1}{2}\vert\vec{v}\vert^2t\right)}.
\end{equation}
To construct the initial field configuration within P\textsc{y}U\textsc{ltra}L\textsc{ight}, the NumPy array encoding the radial profile $f(r)$ for the $f(r=0)=1$ case is loaded, where the spatial resolution of this profile must agree with the expected spatial resolution within the main code. Equation \ref{eq:covariant-solution} is then used to augment the configuration such that the soliton is furnished with the position, mass, and velocity specified by the user in the accompanying Jupyter notebook. The user may also add an additional constant phase factor if desired.

While the current implementation of P\textsc{y}U\textsc{ltra}L\textsc{ight} is designed such that the initial field configuration is built from solitons, it may be simply modified to take in alternative field configurations in future.

\subsection{Timestep Constraints}

When using finite-differencing methods to solve certain partial differential equations, the causality-preserving Courant-Friedrichs-Lewy (CFL) condition on the size of the simulation timestep is a necessary condition for convergence \cite{Ajaib2013}. As such, this condition is often appears in numerical analysis of ULDM physics governed by the SP system, an example being \cite{Schwabe2016}. However, because P\textsc{y}U\textsc{ltra}L\textsc{ight} does not use the method of finite differences to compute spatial derivatives, but rather computes them exactly via explicit multiplication in the Fourier domain, the CFL condition in its usual formulation is not applicable to this code.

In place of the CFL condition, the constraint on the timestep within P\textsc{y}U\textsc{ltra}L\textsc{ight} is set based on the fluid interpretation of the SP system. In dimensionless code units we can write 
\begin{equation}
    \psi\equiv\sqrt{\rho}e^{i\theta}, \quad \vec{v}\equiv\boldsymbol{\nabla}\theta.
\end{equation}
From here we can define a `maximum velocity' for the simulation grid. This is the velocity at which the fluid appears to move backward, which occurs when the phase change between successive grid points exceeds $\pi$. The timestep is then constrained such that a fluid travelling at this maximum grid velocity will traverse a distance of one grid space per timestep.

This proves to be a highly restrictive constraint on the timestep, and can be relaxed significantly at the user's discretion for simulations in which the expected velocities within the system will not approach this maximum grid velocity. Convergence testing on a case-by-case basis is encouraged to determine the timestep constraint necessary for a given degree of accuracy.

\section{Testing and Verification}\label{sec:test}

\subsection{Interference Patterns During Soliton Collisions}\label{sec:interference}

Hierarchical structure formation within the ULDM framework proceeds through both gravitational overdensity collapse as well as through sequential binary mergers of solitonic cores to form individual virialised objects \cite{Schive2014}. It is therefore important that ULDM simulations are able to accurately reproduce the expected characteristics of binary soliton interactions.

Soliton collisions can be classified into two categories depending on whether the total energy of the isolated binary system is positive or negative. In the case of positive energy, the solitons pass through each other, emerging largely undisturbed from their initial configurations following the interaction. During the collision, the wavefunctions describing the solitons are superposed, yielding distinctive interference patterns. In this section we demonstrate the ability of P\textsc{y}U\textsc{ltra}L\textsc{ight} to reproduce the interference patterns predicted by theoretical models during binary collisions.

Following \cite{Schwabe2016}, we consider the head-on collision of two solitions with mass ratio $\mu=2$ and high relative velocity. Using equation \ref{eq:covariant-solution}, we can write the total wavefunction of the binary system in terms of dimensionless quantities along the collision axis as
\begin{align}
    \psi(x,t)=& \ \alpha_1 f\big(\sqrt{\alpha_1}\vert x-x_1-v_1 t\vert\big)e^{i\left(\alpha_1\beta t+v_1(x-x_1)-\frac{1}{2}v_1^2 t+\delta\right)}\nonumber\\
    &+\alpha_2 f\big(\sqrt{\alpha_2}\vert x-x_2-v_2 t\vert\big)e^{i\left(\alpha_2\beta t+v_2(x-x_2)-\frac{1}{2}v_2^2 t\right)}
\end{align}
where $x_1$ and $x_2$ are the initial positions of the centres of the solitions, $v_1$ and $v_2$ are the velocities of the solitons, $\delta$ is a constant relative phase term, and $\alpha_1=1/4 \ \alpha_2$ parameterise the density profiles as discussed in section \ref{sec:soliton-profiles}. Because of the symmetry of the configuration, we take $v_1=-v_2$ and $x_1=-x_2$. We are interested in the density profile of the system at the time of maximal interference, where we expect the superposed wavefunction to possess pronounced features. Therefore, we choose the time at which the two components of the wavefunction are centred at the same location, such that $x_1+v_1t=-x_2-v_2t=0$. This corresponds to a time $t_{max}=\vert x_1/v_1\vert = \vert x_2/v_2\vert$. The dimensionless density is then given by
\begin{align}\label{eq:predicted-interference}
    \vert\psi(x,t_{max})\vert^2= \ \alpha_1^2\bigg[&f(\sqrt{\alpha_1}x)^2+16f(2\sqrt{\alpha_1}x)^2+\nonumber\\
    &8f(\sqrt{\alpha_1}x)f(2\sqrt{\alpha_1}x)\operatorname{cos}\left(-3\alpha_1\beta\left\vert\frac{x_1}{v_1}\right\vert+2v_1x+\delta \right)\bigg]
\end{align}
In Figure \ref{fig:interference}, we plot the dimensionless density profile at the time of maximal interference for two solitons with mass ratio $\mu=2$, and phase difference $\delta=0$. We see that the numerical result obtained using P\textsc{y}U\textsc{ltra}L\textsc{ight} closely matches the theoretical prediction of equation \ref{eq:predicted-interference}. Small disparities between the numerical and theoretical profiles may be attributed to the effect of gravitational contraction not included in the theoretical prediction of equation \ref{eq:predicted-interference}, and to a small offset in the true time of maximal interference. Neglecting these small deviations, we can verify that the code is well suited to faithfully reproduce the wave interference effects of the ULDM model.

\begin{figure}[tbp]
\centering % \begin{center}/\end{center} takes some additional vertical space
\includegraphics{interference_patterns}
\label{fig:interference}
\caption{Comparison of theoretical and numerical density profiles at time of maximal interference for head-on collision of two solitons with mass ratio $\mu=2$ and no relative phase difference. The solitons were chosen with dimensionless masses 5 and 10, with an initial separation of 4 code units and relative velocity of 20 code units. The simulation was run at $256^3$ for a box of side-length 8 code units.}
\end{figure}

\subsection{Effective Forces Due to Destructive Interference}

As demonstrated in \cite{Paredes2016}, the wavelike properties of ULDM can give rise to effective forces which can dramatically affect the dynamics of core collisions. These effective forces arise as a result of interference phenomena, and do not require the ULDM model to incorporate explicit local interactions. Figure \ref{fig:repulsion} shows the results of two simulations using P\textsc{y}U\textsc{ltra}L\textsc{ight} in which two solitons of equal mass undergo a head-on collision. Contours of the density profile along the plane of symmetry are displayed. In the first simulation, there is no phase offset between the initial solitons, while in the second simulation the phases differ by $\pi$. We can see that the result of this phase shift is to create an effective repulsive force between the two solitons as they approach one another. Further discussion of this phenomenon and its possible observational consequences can be found in \cite{Paredes2016}.

\begin{figure}[tbp]
\centering % \begin{center}/\end{center} takes some additional vertical space
\includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{phase_comparison}
\caption{Head-on collisions between two solitons of mass 5, relative velocity 6, and initial separation 4 in code units. Time progresses from left to right across each row.}\label{fig:repulsion}
\end{figure}

% \subsection{Soliton Mergers}

% Want to show radial profile looking like NFW at the outer edges. See http://zulassung.hlrn.de/kurzbeschreibungen/nip00040.pdf

\subsection{Tidal Disruption of Solitons Orbiting a Central Potential}\label{sec:disruption}

P\textsc{y}U\textsc{ltra}L\textsc{ight} is designed such that a central potential may be included by way of a point-mass located at the centre of the simulation grid. While this setup is somewhat unphysical, it provides a starting point for a study of the stability of satellite dark matter halos orbiting a much larger object. Studies of this type can be used to make predictions regarding the lifespans of satellite galaxies around the Milky Way, offering a potential resolution of the so-called missing satellites problem \cite{Weinberg2015}.

An extensive study of the tidal disruption of ULDM solitonic cores orbiting a central potential has recently been undertaken in \cite{Du2018}. Here we reproduce just one of their results as a validity check for P\textsc{y}U\textsc{ltra}L\textsc{ight}. In particular, it is found in \cite{Du2018} that an initially spherical solitonic core without self-rotation will gradually spin up to form a tidally-locked irrotational Riemann -S ellipsoid when orbiting a host mass. Figure \ref{fig:riemann} shows the internal velocity field of a solitonic satellite after one complete revolution around a host mass, obtained using P\textsc{y}U\textsc{ltra}L\textsc{ight} on a $256^3$ simulation grid. In this case the ratio of host to satellite mass was $\approx 55$. It can be seen that the soliton has become elongated along the radial line connecting to the host, indicating that it is tidally locked. Meanwhile the internal velocity field can be seen to be that of an irrotational Riemann-S ellipsoid \cite{Daller2012}.

\begin{figure}[tbp]
\centering 
\includegraphics[width=1.\textwidth,trim=1cm 1cm 0 1cm,clip]{riemann}
\caption{Solitonic core after one revolution of a central potential, demonstrating the internal velocity field of an irrotational Riemann-S ellipsoid (where the collective motion of the soliton has been subtracted from the velocity field). Deformation of the spherical density profile into that of an ellipsoid can be seen, with extension along the radial direction (red arrow).}\label{fig:riemann}
\end{figure}

\subsection{Energy Conservation}\label{sec:energy}

In addition to the evolution of the condensate wavefunction $\psi$ and the potential $\Phi$, P\textsc{y}U\textsc{ltra}L\textsc{ight} also evaluates the energy of the SP system. An important achievement of the code is sub-percent level energy conservation, even at relatively low spatial resolution, for all dynamical scenarios tested. In this section we derive the expression for the total energy of the SP system and discuss its decomposition into individual constituents calculated separately within the code. We then present the numerical energy evolution of a variety of configurations.

We begin by defining a suitable action which yields the full SP system through its corresponding Euler-Lagrange equations. We find that variation of
\begin{equation}\label{eq:sp-action}
    S=\int dt\int_{\mathbb{R}^3} d^3x \  -\bigg\{\frac{1}{2}\vert\nabla\Phi\vert^2+\Phi\vert\psi\vert^2+\frac{1}{2}\vert\nabla\psi\vert^2+\frac{i}{2}(\psi\Dot{\psi}^*-\Dot{\psi}\psi^*)\bigg\}
\end{equation}
with respect to $\Phi$, $\psi^*$ and $\psi$ yields equations \ref{eq:p-adim}, \ref{eq:s-adim}, and the conjugate of equation \ref{eq:s-adim}, respectively. The integrand of equation \ref{eq:sp-action} is the Lagrangian density, $\mathcal{L}$, from which we can derive the conserved energy in the usual way:
\begin{equation}
    E_{tot}=\int_{\mathbb{R}^3}d^3x \ \bigg\{\frac{\partial \mathcal{L}}{\partial \Dot{\psi}}\Dot{\psi}+\frac{\partial \mathcal{L}}{\partial \Dot{\psi}^*}\Dot{\psi}^*+\frac{\partial \mathcal{L}}{\partial \Dot{\Phi}}\Dot{\Phi}-\mathcal{L}\bigg\}.
\end{equation}
Evaluating this expression, we obtain:
\begin{align}
    E_{tot}&=\int_{\mathbb{R}^3}d^3x \ \bigg\{\frac{1}{2}\vert\nabla\Phi\vert^2+\Phi\vert\psi\vert^2+\frac{1}{2}\vert\nabla\psi\vert^2\bigg\}\\
    &=\int_{\mathbb{R}^3}d^3x \ \bigg\{\frac{1}{2}\nabla(\Phi\nabla\Phi)-\frac{1}{2}\Phi\nabla^2\Phi+\Phi\vert\psi\vert^2+\frac{1}{2}\nabla(\psi^*\nabla\psi)-\frac{1}{2}\psi^*\nabla^2\psi\bigg\}\\
    &=\int_{\mathbb{R}^3}d^3x \ \bigg\{\frac{1}{2}\Phi\vert\psi\vert^2-\frac{1}{2}\psi^*\nabla^2\psi\bigg\}.\label{eq:energy-tot}
\end{align}
where in the last step we have used Stokes' Theorem as well as the Poisson equation (\ref{eq:p-adim}) to perform simplifications. Because we are working with the dimensionless quantities defined in equation \ref{eq:dimensionless}, it is easy to see that this quantity is related to the physical energy through multiplication by a constant factor of $\CMcal{L}^5\CMcal{T}^{-4}G^{-1}$. It should be noted that equation \ref{eq:energy-tot} is not equivalent to the expectation value of the Schr{\"o}dinger Hamiltonian, which is itself not a conserved quantity of the SP system and is given by
\begin{equation}
    \langle\hat{H}\rangle=\int_{\mathbb{R}^3}d^3x \ \bigg\{\Phi\vert\psi\vert^2-\frac{1}{2}\psi^*\nabla^2\psi\bigg\}.
\end{equation}

The two terms in the integral \ref{eq:energy-tot} are calculated separately within the code. It is easy to see that the first term represents the gravitational potential energy of the SP system. As discussed in \cite{Hui2016}, the second term may be decomposed further into two contributions which may be considered separately as kinetic and `quantum' energies, but for the purposes of this implementation it is sufficient to consider only their combined contribution. Because P\textsc{y}U\textsc{ltra}L\textsc{ight} provides the option to include the central potential of a point mass located at the centre of the simulation grid, we have additional energy contributions when such a potential is included. Because this central potential does not arise from a ULDM object, we calculate the gravitational potential energy due to self-interaction separately from the gravitational potential energy due to the central potential.

Figures \ref{fig:combined_1} and \ref{fig:binary} demonstrate energy conservation for two possible simulation scenarios. In the first case, we examine the evolution of the energy for a single soliton undergoing significant tidal disruption whilst orbiting within a central potential. We see that as the soliton is disrupted, the kinetic energy increases, while the gravitational energy due to the central potential decreases, as expected. Meanwhile, the gravitational potential energy due to self-interaction gradually increases toward zero as the disruption continues. The sum of the individual energy components is conserved to $10^{-2}\ \%$ in this case, where the simulation was run at $256^3$. Higher grid resolutions result in even better energy conservation. Figure \ref{fig:binary} demonstrates the evolution of the energy of a binary system of solitons in elliptical orbits around their common centre of mass. In this case the simulation ran for approximately three orbital periods at $256^3$. At points of closest approach, it can be clearly seen that the kinetic energy increases as the solitons speed up, while the potential energy due to self-interaction decreases commensurately. In this scenario no central potential has been included, so the gravitational energy due to the central potential stays uniformly zero throughout the simulation. The small scale oscillations that can be observed are artifacts of the relatively low spatial resolution, and smooth out as grid resolution is increased. In this case the total energy is conserved to $10^{-3}\ \%$.

% \begin{figure}[tbp]
% \centering 
% \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{density_diagram1}
% \caption{Density profile of solitonic core at equally spaced times as it undergoes tidal disruption in a potential centred at the red cross.}\label{fig:density_diagram1}
% \end{figure}

% \begin{figure}[tbp]
% \centering 
% \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{energy_diagram1}
% \caption{Evolution of the energy of the system illustrated in Figure \ref{fig:density_diagram1}. Times are indicated in correspondence with the snapshots of the density profile. }\label{fig:energy_diagram_1}
% \end{figure}

\begin{figure}[tbp]
\centering 
\includegraphics[width=1.1\textwidth,trim=2.5cm 0 0 1cm,clip]{combined_energy_and_density1}
\caption{Left: Evolution of the density profile of a solitonic core at equally spaced times as it undergoes tidal disruption in a potential centred at the red cross. Right: Evolution of the energy of the system, where times are indicated in correspondence with the snapshots of the density profile. All quantities are in dimensionless code units.}\label{fig:combined_1}
\end{figure}

\begin{figure}[tbp]
\centering 
\includegraphics[width=1.\textwidth,trim=0 0 2cm 2cm,clip]{energy_diagram2}
\caption{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}\label{fig:binary}
\end{figure}

\section{Discussion and Outlook}

We have demonstrated that P\textsc{y}U\textsc{ltra}L\textsc{ight} is an accurate tool for the study of the dynamics of ultra-light dark matter governed by the Schr{\"o}dinger-Poisson system of equations. The code makes use of a pseudo-spectral symmetrised split-step Fourier methodology, in which all spatial derivatives are treated via explicit multiplication in the Fourier domain, thereby avoiding difficulties associated with finite-differencing methods.

Energy conservation within P\textsc{y}U\textsc{ltra}L\textsc{ight} is excellent, at sub-percent level for simulations run at $128^3$, and even better performance as resolution is increased. The code is able to reproduce complex phenomena resulting from the wave-like properties of ultra-light dark matter as predicted by theoretical models, such as the interference patterns arising during high-velocity collisions of solitonic cores, and the effective forces observed in cases where the colliding cores are out of phase. These phenomena can be clearly observed at relatively low spatial resolution, obviating the need for high-performance computing infrastructure to study the fundamental behaviour of ULDM systems. This makes P\textsc{y}U\textsc{ltra}L\textsc{ight} a useful tool for predicting the overall dynamics of a ULDM system in a computationally efficient manner prior to a rigorous examination using more advanced codes such as Nyx.

P\textsc{y}U\textsc{ltra}L\textsc{ight} is Python-based, and as such is particularly simple to understand and use. The accompanying Jupyter notebook allows for the efficient adjustment of simulation parameters, and offers a useful browser interface for quick visualisation of simulation results. Despite being Python-based, the code makes use of low-level language resources, namely the FFTW libraries through the use of the Pythonic pyFFTW wrapper.

While the current implementation of P\textsc{y}U\textsc{ltra}L\textsc{ight} is already a useful tool for simulating dynamical ULDM systems, there remains scope for substantial improvement. In particular, it is anticipated that future releases may incorporate adaptive mesh refinement capabilities and variable timestep functionality. Additionally, while the current implementation utilises a spatial grid of fixed size and does not provide a means by which to incorporate baryonic density contributions, future releases are likely to allow for cosmic expansion and a means by which to incorporate simplified models of baryonic matter. Further adaptations of the code may also be implemented to allow for greater flexibility in the choice of ULDM model, incorporating possibilities such as explicit self-interaction.

\appendix
\section{Download and Licensing}

To be written.

\acknowledgments

To be written.

% The bibliography will probably be heavily edited during typesetting.
% We'll parse it and, using the arxiv number or the journal data, will
% query inspire, trying to verify the data (this will probalby spot
% eventual typos) and retrive the document DOI and eventual errata.
% We however suggest to always provide author, title and journal data:
% in short all the informations that clearly identify a document.

%\section*{References}
\bibliographystyle{JHEP}
\bibliography{refs}

%\begin{thebibliography}{99}

%\bibitem{a}
%Author, \emph{Title}, \emph{J. Abbrev.} {\bf vol} (year) pg.

%\bibitem{b}
%Author, \emph{Title},
%arxiv:1234.5678.

%\bibitem{c}
%Author, \emph{Title},
%Publisher (year).

% Please avoid comments such as "For a review'', "For some examples",
% "and references therein" or move them in the text. In general,
% please leave only references in the bibliography and move all
% accessory text in footnotes.

% Also, please have only one work for each \bibitem.

%\end{thebibliography}

\end{document}

### 2018-05-09 draft.tex
Initial ShareLaTeX Import

--- 2018-05-09 Backup.tex
@@
 \usepackage{eucal}
-\usepackage{float}
+
+\usepackage[justification=justified]{caption}
+\usepackage{varwidth}
+\usepackage{graphicx}
 
@@
 Neglecting these small deviations, we can verify that the code is well suited to faithfully reproduce the wave interference effects of the ULDM model.
-
-\begin{figure}[tbp]
-\centering % \begin{center}/\end{center} takes some additional vertical space
-\includegraphics{interference_patterns}
-\label{fig:interference}
-\caption{Comparison of theoretical and numerical density profiles at time of maximal interference for head-on collision of two solitons with mass ratio $\mu=2$ and no relative phase difference.
+\begin{center}
+  \includegraphics[trim={0 0 0 0.9cm},clip, scale=0.9]{interference_patterns}
+  \captionof{figure}{Comparison of theoretical and numerical density profiles at time of maximal interference for head-on collision of two solitons with mass ratio $\mu=2$ and no relative phase difference.
 The solitons were chosen with dimensionless masses 5 and 10, with an initial separation of 4 code units and relative velocity of 20 code units.
 The simulation was run at $256^3$ for a box of side-length 8 code units.}
-\end{figure}
+  \label{fig:interference}
+\end{center}
 
@@
 Further discussion of this phenomenon and its possible observational consequences can be found in \cite{Paredes2016}.
+\begin{center}
+  \includegraphics[width=1.\textwidth, trim={0 0 0 0},clip]{phase_comparison}
+  \captionof{figure}{Head-on collisions between two solitons of mass 5, relative velocity 6, and initial separation 4 in code units. Time progresses from left to right across each row.}
+  \label{fig:repulsion}
+\end{center}
 
-\begin{figure}[tbp]
-\centering % \begin{center}/\end{center} takes some additional vertical space
-\includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{phase_comparison}
-\caption{Head-on collisions between two solitons of mass 5, relative velocity 6, and initial separation 4 in code units. Time progresses from left to right across each row.}\label{fig:repulsion}
-\end{figure}
-
-% \subsection{Soliton Mergers}
-
-% Want to show radial profile looking like NFW at the outer edges. See http://zulassung.hlrn.de/kurzbeschreibungen/nip00040.pdf
+\vspace{1em}
 
@@
 Meanwhile the internal velocity field can be seen to be that of an irrotational Riemann-S ellipsoid \cite{Daller2012}.
+\begin{center}
+  \includegraphics[width=1.\textwidth,trim=1cm 1cm 0 1cm,clip]{riemann}
+  \captionof{figure}{Solitonic core after one revolution of a central potential, demonstrating the internal velocity field of an irrotational Riemann-S ellipsoid (where the collective motion of the soliton has been subtracted from the velocity field).
+Deformation of the spherical density profile into that of an ellipsoid can be seen, with extension along the radial direction (red arrow).}
+  \label{fig:riemann}
+\end{center}
 
-\begin{figure}[tbp]
-\centering 
-\includegraphics[width=1.\textwidth,trim=1cm 1cm 0 1cm,clip]{riemann}
-\caption{Solitonic core after one revolution of a central potential, demonstrating the internal velocity field of an irrotational Riemann-S ellipsoid (where the collective motion of the soliton has been subtracted from the velocity field).
-Deformation of the spherical density profile into that of an ellipsoid can be seen, with extension along the radial direction (red arrow).}\label{fig:riemann}
-\end{figure}
+\vspace{1em}
 
@@
 
-\begin{figure}[tbp]
-\centering 
-\includegraphics[width=1.1\textwidth,trim=2.5cm 0 0 1cm,clip]{combined_energy_and_density1}
-\caption{Left: Evolution of the density profile of a solitonic core at equally spaced times as it undergoes tidal disruption in a potential centred at the red cross.
+% \begin{figure}[tbp]
+% \centering 
+% \includegraphics[width=1.1\textwidth,trim=2.5cm 0 0 1cm,clip]{combined_energy_and_density1}
+% \caption{Left: Evolution of the density profile of a solitonic core at equally spaced times as it undergoes tidal disruption in a potential centred at the red cross.
 Right: Evolution of the energy of the system, where times are indicated in correspondence with the snapshots of the density profile.
 All quantities are in dimensionless code units.}\label{fig:combined_1}
-\end{figure}
+% \end{figure}
 
-\begin{figure}[tbp]
-\centering 
-\includegraphics[width=1.\textwidth,trim=0 0 2cm 2cm,clip]{energy_diagram2}
-\caption{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}\label{fig:binary}
-\end{figure}
+% \begin{figure}[tbp]
+% \centering 
+% \includegraphics[width=1.\textwidth,trim=0 0 2cm 2cm,clip]{energy_diagram2}
+% \caption{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}\label{fig:binary}
+% \end{figure}
+
+\begin{center}
+  \includegraphics[width=1.1\textwidth,trim=2.5cm 0 0 1cm,clip]{combined_energy_and_density1}
+  \captionof{figure}{Left: Evolution of the density profile of a solitonic core at equally spaced times as it undergoes tidal disruption in a potential centred at the red cross.
+Right: Evolution of the energy of the system, where times are indicated in correspondence with the snapshots of the density profile.
+All quantities are in dimensionless code units.}
+  \label{fig:combined_1}
+\end{center}
+
+\begin{center}
+  \includegraphics[width=0.9\textwidth,trim=0 0 2cm 2cm,clip]{energy_diagram2}
+  \captionof{figure}{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}
+  \label{fig:binary}
+\end{center}
 

### 2018-05-15 draft.tex
changes from Richard; tweaks to format to work well with standalone LaTeX, tweaks to refs, and rewrite of abstract/intro

@@
 \usepackage[justification=justified]{caption}
-\usepackage{varwidth}
-\usepackage{graphicx}
 
-\title{\boldmath PyUltraLight: A Pseudo-Spectral Code for the Dynamics of Ultra-Light Dark Matter}
+\title{ PyUltraLight: A Pseudo-Spectral Solver for Ultralight Dark Matter Dynamics}
 
@@
 
-\abstract{P\textsc{y}U\textsc{ltra}L\textsc{ight} is a Python based program that solves the Schr{\"o}dinger-Poisson coupled differential equations, which govern the dynamics of ultra-light scalar field dark matter in the non-relativistic regime.
-P\textsc{y}U\textsc{ltra}L\textsc{ight} utilises a symmetrised split-step pseudo-spectral Fourier algorithm to solve these differential equations, and demonstrates excellent energy conservation.
-Being Python-based, the program is particularly easy to set up, and the accompanying Jupyter notebook provides a simple interface in which to augment simulation parameters and create visualisations of output data.
-Using P\textsc{y}U\textsc{ltra}L\textsc{ight}, we are able to reproduce many previously studied dynamical characteristics of ultra-light dark matter, even when running the simulations at relatively low spatial resolution from a standard user-PC.}
+\newcommand{\apj}{Astrophysical J.}
+\newcommand{\prd}{Phys. Rev. D}
+\newcommand{\mnras}{Mon. Notices Royal Astron. Soc}
+\newcommand{\PyUltraLight}{\textsc{PyUltraLight}}
+
+\abstract{\PyUltraLight\ simulates the dynamics of ultralight dark matter in a static, non-expanding background.
+\PyUltraLight\  can describe the evolution of several interacting ultralight dark matter halos  or one or more halos orbiting a central, fixed Newtonian potential, the latter scenario corresponding to dwarf galaxies orbiting a massive central galaxy.
+We verify  \PyUltraLight\  by showing that it reproduces qualitative dynamical features of previously published simulations and demonstrate that it has  excellent energy-conservation properties.
+\PyUltraLight\  is implemented in a Python-based Jupyter notebook, solving the Schr{\"o}dinger-Poisson equation  governing  ultralight scalar field dark matter dynamics in the non-relativistic regime using a symmetrised split-step pseudo-spectral  algorithm.
+The notebook interface makes it simple to specify simulation parameters and  visualise the resulting output.
+\PyUltraLight\  runs on standard desktop hardware and is available on GitHub.
+}
 
@@
 
-The true nature of dark matter (DM) is one of the most important outstanding problems in theoretical physics.
-WIMP models of dark matter are favoured by many, due in part to their consistency with supersymmetric extensions to the standard model of particle physics.
-However, with increasingly advanced direct-detection experiments failing to find evidence for WIMP candidates, the viable parameter space continues to shrink \cite{Tan2016, Akerib2017}.
-Furthermore, the $\Lambda$CDM cosmologies consistent with WIMP models suffer from a number of widely-publicised small-scale problems \cite{Bull2016}.
-While baryonic physics is proposed to offer solutions to a number of these problems, the perceived issues within $\Lambda$CDM are increasingly motivating novel research into alternative DM models, one such class of model being ultra-light dark matter (ULDM), also known as fuzzy dark matter (FDM), or BEC dark matter.
+The realisation that there may be more to the universe than meets the eye is one of the most profound developments in 20$^{\textrm{th}}$ Century astronomy and astrophysics.
+However,  while there are now multiple lines of evidence that dark matter outweighs baryonic matter by a ratio of approximately 5:1 [CITE], we have few clues regarding the physical nature of dark matter.
+Much theoretical and experimental effort has focussed on  WIMP models, motivated by their consistency with supersymmetric extensions to the Standard Model and their relatively simple dynamics but advanced direct-detection experiments are putting increasingly tight constraints on the WIMP parameter space \cite{Tan2016, Akerib2017}.
+Moreover, $\Lambda$CDM cosmology with simple, pressureless, noninteracting dark matter (a class which includes simple WIMP models) is potentially at odds with observations at small astrophysical scales \cite{Bull2016}.
+
+[TO HERE]The potential shortcoming of simple cold dark matter scenarios are increasingly motivating the investigations of more novel dark matter scenarios.
+In particular, ultralight dark matter (ULDM) is an increasingly well-studied possibility, also known as fuzzy dark matter (FDM), or BEC dark matter.
 ULDM models are well motivated by fundamental theories possessing approximate shift symmetries, and offer natural resolutions to the small-scale problems of $\Lambda$CDM without the need to invoke baryonic physics.
@@
 While these modified cosmological simulation codes have produced promising results in the study of ULDM physics, at the time of writing only one such code, \textsc{axion-gadget}, has been made publicly available.
-The aim of this paper is to introduce P\textsc{y}U\textbf{ltra}L\textsc{ight}, a stand-alone Python-based pseudo-spectral SP solver, and to demonstrate that this code is able to reproduce a number of key findings from the more complicated cosmological simulation codes without the need for high-performance computing environments.
-Consequently, we anticipate that as a publicly available resource, P\textsc{y}U\textbf{ltra}L\textsc{ight} will serve as a valuable cross-check for more complex implementations and may serve as a basis for further development of such codes within the computational cosmology community.
+The aim of this paper is to introduce \PyUltraLight, a stand-alone Python-based pseudo-spectral SP solver, and to demonstrate that this code is able to reproduce a number of key findings from the more complicated cosmological simulation codes without the need for high-performance computing environments.
+Consequently, we anticipate that as a publicly available resource, \PyUltraLight\ will serve as a valuable cross-check for more complex implementations and may serve as a basis for further development of such codes within the computational cosmology community.
 
-P\textsc{y}U\textbf{ltra}L\textsc{ight} makes use of a symmetrised-split-step pseudo-spectral Fourier algorithm (also known as leap-frog) to compute the dynamics of the SP system.
+\PyUltraLight\ makes use of a symmetrised-split-step pseudo-spectral Fourier algorithm (also known as leap-frog) to compute the dynamics of the SP system.
 We note that a similar methodology has been described in \cite{Paredes2016}, though to the authors' knowledge this code has not been released publicly.
@@
 As a pseudo-spectral code, linear differential operators are computed by direct multiplication in the Fourier domain, while non-linear terms are evaluated in position space.
-This means that P\textsc{y}U\textbf{ltra}L\textsc{ight} is free from noise associated with finite-differencing methods of computing spatial derivatives, though there is a necessary computational cost associated with the Fourier and inverse Fourier transforms.
-These transforms are optimised in P\textsc{y}U\textbf{ltra}L\textsc{ight} through the use of the pyFFTW pythonic wrapper around the C-based FFWTW subroutine library \cite{pyfftw} \cite{fftw}.
+This means that \PyUltraLight\ is free from noise associated with finite-differencing methods of computing spatial derivatives, though there is a necessary computational cost associated with the Fourier and inverse Fourier transforms.
+These transforms are optimised in \PyUltraLight\ through the use of the pyFFTW pythonic wrapper around the C-based FFWTW subroutine library \cite{pyfftw} \cite{fftw}.
 While the code is currently implemented to run on a singe machine, the FFTW libraries offer full parallelisation, so there is no barrier to running the code in a cluster environment.
@@
 We first provide a short review of the physics of ULDM, including a derivation of the SP equations from an appropriate scalar-field Lagrangian.
-We then describe the details of the implementation of P\textsc{y}U\textbf{ltra}L\textsc{ight}, before moving on to a testing and verification section, in which we reproduce a selection of results from a variety of recent ULDM simulations and discuss energy conservation within the code.
+We then describe the details of the implementation of \PyUltraLight, before moving on to a testing and verification section, in which we reproduce a selection of results from a variety of recent ULDM simulations and discuss energy conservation within the code.
 
@@
 \end{equation}
+%
 where we have taken $c=1$ but retain factors of $\hbar$ at this stage. Applying the variational principle to this action yields the Euler-Lagrange equations
+%
 \begin{equation}\label{eq:e-l}
@@
 The coupled equations \ref{eq:schrodinger} and \ref{eq:poisson} together form the nonlinear Schr{\"o}dinger-Poisson (SP) system which describes the dynamics of ULDM in the non-relativistic regime.
-It is this system of equations which P\textsc{y}U\textsc{ltra}L\textsc{ight} is designed to solve.
+It is this system of equations which \PyUltraLight is designed to solve.
 
@@
 \end{align}
-where $m$ is the mass of the ultra-light scalar field, $H_0$ is the present-day Hubble parameter, $G$ is Newton's gravitational constant and $\Omega_{m_0}$ is the present-day matter fraction of the energy density of the universe.
+where $m$ is the mass of the ultralight scalar field, $H_0$ is the present-day Hubble parameter, $G$ is Newton's gravitational constant and $\Omega_{m_0}$ is the present-day matter fraction of the energy density of the universe.
 We now recast equations \ref{eq:schrodinger} and \ref{eq:poisson} in terms of the following dimensionless quantities:
@@
 \end{equation}
-All computations performed by P\textsc{y}U\textsc{ltra}L\textsc{ight} are carried out using the dimensionless quantities defined in \ref{eq:dimensionless}, while provisions are made within the code for initial conditions and outputs to be converted to and from the desired unit system specified by the user.
+All computations performed by \PyUltraLight\ are carried out using the dimensionless quantities defined in \ref{eq:dimensionless}, while provisions are made within the code for initial conditions and outputs to be converted to and from the desired unit system specified by the user.
 Henceforth, we will often refer to the quantity $\vert\psi\vert^2$ simply as the density, where it is understood that this is in fact a dimensionless quantity related to the physical mass density through a constant of proportionality given by equation \ref{eq:density}.
@@
 
-Dynamical evolution within P\textsc{y}U\textsc{ltra}L\textsc{ight} progresses via a symmetrised split-step Fourier process on an $N\times N\times N$ simulation grid with spatial periodic boundary conditions.
+Dynamical evolution within \PyUltraLight\ progresses via a symmetrised split-step Fourier process on an $N\times N\times N$ simulation grid with spatial periodic boundary conditions.
 To understand this method, we first consider the exact expression for the unitary time evolution of the wavefunction according to equation \ref{eq:s-adim}, namely
@@
 
-Evaluation of equation \ref{eq:split-step} within P\textsc{y}U\textsc{ltra}L\textsc{ight} proceeds as follows:
+Evaluation of equation \ref{eq:split-step} within \PyUltraLight proceeds as follows:
 Initially, the nonlinear term acts in position space for one half-timestep.
@@
 
-In the P\textsc{y}U\textsc{ltra}L\textsc{ight} implementation, we make use of an additional simplification of the symmetrised split-step Fourier method.
+In the \PyUltraLight implementation, we make use of an additional simplification of the symmetrised split-step Fourier method.
 Namely, consecutive half-steps in the nonlinear term may be combined into a single full step in each case such that when iterated, only the first and last operations involve half steps.
@@
 
-The discrete Fourier transform (DFT) and its inverse are handled within P\textsc{y}U\textsc{ltra}L\textsc{ight} using pyFFTW.
+The discrete Fourier transform (DFT) and its inverse are handled within \PyUltraLight using pyFFTW.
 pyFFTW is a pythonic wrapper around the C-based FFTW subroutine library, and allows for efficient evaluation of both real and complex DFTs \cite{pyfftw} \cite{fftw} \cite{Frigo2005}.
-Multithreading is supported by pyFFTW, and is implemented within P\textsc{y}U\textsc{ltra}L\textsc{ight} by using the multiprocessing package to determine the number of online CPUs and feed this directly to the threads argument of the pyfftw.FFTW class.
+Multithreading is supported by pyFFTW, and is implemented within \PyUltraLight by using the multiprocessing package to determine the number of online CPUs and feed this directly to the threads argument of the pyfftw.FFTW class.
 
-In addition to the use of pyFFTW, performance within P\textsc{y}U\textsc{ltra}L\textsc{ight} is further improved through the use of the NumExpr package, which provides routines for the efficient manipulation of array objects within the simulation using a vector-based virtual machine \cite{numexpr}.
+In addition to the use of pyFFTW, performance within \PyUltraLight is further improved through the use of the NumExpr package, which provides routines for the efficient manipulation of array objects within the simulation using a vector-based virtual machine \cite{numexpr}.
 
@@
 
-The initialisation procedure within P\textsc{y}U\textsc{ltra}L\textsc{ight} can be easily augmented to accommodate the requirements of the user.
+The initialisation procedure within \PyUltraLight can be easily augmented to accommodate the requirements of the user.
 In the current implementation, the initial field configuration is built by loading a NumPy array file encoding a solitonic solution to the Schr{\"o}dinger-Poisson system.
@@
 \end{equation}
-To construct the initial field configuration within P\textsc{y}U\textsc{ltra}L\textsc{ight}, the NumPy array encoding the radial profile $f(r)$ for the $f(r=0)=1$ case is loaded, where the spatial resolution of this profile must agree with the expected spatial resolution within the main code.
+To construct the initial field configuration within \PyUltraLight, the NumPy array encoding the radial profile $f(r)$ for the $f(r=0)=1$ case is loaded, where the spatial resolution of this profile must agree with the expected spatial resolution within the main code.
 Equation \ref{eq:covariant-solution} is then used to augment the configuration such that the soliton is furnished with the position, mass, and velocity specified by the user in the accompanying Jupyter notebook.
@@
 
-While the current implementation of P\textsc{y}U\textsc{ltra}L\textsc{ight} is designed such that the initial field configuration is built from solitons, it may be simply modified to take in alternative field configurations in future.
+While the current implementation of \PyUltraLight is designed such that the initial field configuration is built from solitons, it may be simply modified to take in alternative field configurations in future.
 
@@
 As such, this condition is often appears in numerical analysis of ULDM physics governed by the SP system, an example being \cite{Schwabe2016}.
-However, because P\textsc{y}U\textsc{ltra}L\textsc{ight} does not use the method of finite differences to compute spatial derivatives, but rather computes them exactly via explicit multiplication in the Fourier domain, the CFL condition in its usual formulation is not applicable to this code.
+However, because \PyUltraLight does not use the method of finite differences to compute spatial derivatives, but rather computes them exactly via explicit multiplication in the Fourier domain, the CFL condition in its usual formulation is not applicable to this code.
 
-In place of the CFL condition, the constraint on the timestep within P\textsc{y}U\textsc{ltra}L\textsc{ight} is set based on the fluid interpretation of the SP system.
-In dimensionless code units we can write
+In place of the CFL condition, the constraint on the timestep within \PyUltraLight is set based on the fluid interpretation of the SP system. In dimensionless code units we can write 
 \begin{equation}
@@
 During the collision, the wavefunctions describing the solitons are superposed, yielding distinctive interference patterns.
-In this section we demonstrate the ability of P\textsc{y}U\textsc{ltra}L\textsc{ight} to reproduce the interference patterns predicted by theoretical models during binary collisions.
+In this section we demonstrate the ability of \PyUltraLight to reproduce the interference patterns predicted by theoretical models during binary collisions.
 
@@
 In Figure \ref{fig:interference}, we plot the dimensionless density profile at the time of maximal interference for two solitons with mass ratio $\mu=2$, and phase difference $\delta=0$.
-We see that the numerical result obtained using P\textsc{y}U\textsc{ltra}L\textsc{ight} closely matches the theoretical prediction of equation \ref{eq:predicted-interference}.
+We see that the numerical result obtained using \PyUltraLight closely matches the theoretical prediction of equation \ref{eq:predicted-interference}.
 Small disparities between the numerical and theoretical profiles may be attributed to the effect of gravitational contraction not included in the theoretical prediction of equation \ref{eq:predicted-interference}, and to a small offset in the true time of maximal interference.
 Neglecting these small deviations, we can verify that the code is well suited to faithfully reproduce the wave interference effects of the ULDM model.
-\begin{center}
+\begin{figure}
   \includegraphics[trim={0 0 0 0.9cm},clip, scale=0.9]{interference_patterns}
-  \captionof{figure}{Comparison of theoretical and numerical density profiles at time of maximal interference for head-on collision of two solitons with mass ratio $\mu=2$ and no relative phase difference.
+  \caption{Comparison of theoretical and numerical density profiles at time of maximal interference for head-on collision of two solitons with mass ratio $\mu=2$ and no relative phase difference.
 The solitons were chosen with dimensionless masses 5 and 10, with an initial separation of 4 code units and relative velocity of 20 code units.
@@
   \label{fig:interference}
-\end{center}
+\end{figure}
 
@@
 These effective forces arise as a result of interference phenomena, and do not require the ULDM model to incorporate explicit local interactions.
-Figure \ref{fig:repulsion} shows the results of two simulations using P\textsc{y}U\textsc{ltra}L\textsc{ight} in which two solitons of equal mass undergo a head-on collision.
+Figure \ref{fig:repulsion} shows the results of two simulations using \PyUltraLight in which two solitons of equal mass undergo a head-on collision.
 Contours of the density profile along the plane of symmetry are displayed.
@@
 Further discussion of this phenomenon and its possible observational consequences can be found in \cite{Paredes2016}.
-\begin{center}
+\begin{figure}
   \includegraphics[width=1.\textwidth, trim={0 0 0 0},clip]{phase_comparison}
-  \captionof{figure}{Head-on collisions between two solitons of mass 5, relative velocity 6, and initial separation 4 in code units. Time progresses from left to right across each row.}
+  \caption{Head-on collisions between two solitons of mass 5, relative velocity 6, and initial separation 4 in code units. Time progresses from left to right across each row.}
   \label{fig:repulsion}
-\end{center}
+\end{figure}
 
@@
 
-P\textsc{y}U\textsc{ltra}L\textsc{ight} is designed such that a central potential may be included by way of a point-mass located at the centre of the simulation grid.
+\PyUltraLight is designed such that a central potential may be included by way of a point-mass located at the centre of the simulation grid.
 While this setup is somewhat unphysical, it provides a starting point for a study of the stability of satellite dark matter halos orbiting a much larger object.
@@
 An extensive study of the tidal disruption of ULDM solitonic cores orbiting a central potential has recently been undertaken in \cite{Du2018}.
-Here we reproduce just one of their results as a validity check for P\textsc{y}U\textsc{ltra}L\textsc{ight}.
+Here we reproduce just one of their results as a validity check for \PyUltraLight.
 In particular, it is found in \cite{Du2018} that an initially spherical solitonic core without self-rotation will gradually spin up to form a tidally-locked irrotational Riemann -S ellipsoid when orbiting a host mass.
-Figure \ref{fig:riemann} shows the internal velocity field of a solitonic satellite after one complete revolution around a host mass, obtained using P\textsc{y}U\textsc{ltra}L\textsc{ight} on a $256^3$ simulation grid.
+Figure \ref{fig:riemann} shows the internal velocity field of a solitonic satellite after one complete revolution around a host mass, obtained using \PyUltraLight on a $256^3$ simulation grid.
 In this case the ratio of host to satellite mass was $\approx 55$.
@@
 Meanwhile the internal velocity field can be seen to be that of an irrotational Riemann-S ellipsoid \cite{Daller2012}.
-\begin{center}
+\begin{figure}
   \includegraphics[width=1.\textwidth,trim=1cm 1cm 0 1cm,clip]{riemann}
-  \captionof{figure}{Solitonic core after one revolution of a central potential, demonstrating the internal velocity field of an irrotational Riemann-S ellipsoid (where the collective motion of the soliton has been subtracted from the velocity field).
+  \caption{Solitonic core after one revolution of a central potential, demonstrating the internal velocity field of an irrotational Riemann-S ellipsoid (where the collective motion of the soliton has been subtracted from the velocity field).
 Deformation of the spherical density profile into that of an ellipsoid can be seen, with extension along the radial direction (red arrow).}
   \label{fig:riemann}
-\end{center}
+\end{figure}
 
@@
 
-In addition to the evolution of the condensate wavefunction $\psi$ and the potential $\Phi$, P\textsc{y}U\textsc{ltra}L\textsc{ight} also evaluates the energy of the SP system.
+In addition to the evolution of the condensate wavefunction $\psi$ and the potential $\Phi$, \PyUltraLight also evaluates the energy of the SP system.
 An important achievement of the code is sub-percent level energy conservation, even at relatively low spatial resolution, for all dynamical scenarios tested.
@@
 As discussed in \cite{Hui2016}, the second term may be decomposed further into two contributions which may be considered separately as kinetic and `quantum' energies, but for the purposes of this implementation it is sufficient to consider only their combined contribution.
-Because P\textsc{y}U\textsc{ltra}L\textsc{ight} provides the option to include the central potential of a point mass located at the centre of the simulation grid, we have additional energy contributions when such a potential is included.
+Because \PyUltraLight provides the option to include the central potential of a point mass located at the centre of the simulation grid, we have additional energy contributions when such a potential is included.
 Because this central potential does not arise from a ULDM object, we calculate the gravitational potential energy due to self-interaction separately from the gravitational potential energy due to the central potential.
@@
 
-\begin{center}
+\begin{figure}
   \includegraphics[width=1.1\textwidth,trim=2.5cm 0 0 1cm,clip]{combined_energy_and_density1}
-  \captionof{figure}{Left: Evolution of the density profile of a solitonic core at equally spaced times as it undergoes tidal disruption in a potential centred at the red cross.
+  \caption{Left: Evolution of the density profile of a solitonic core at equally spaced times as it undergoes tidal disruption in a potential centred at the red cross.
 Right: Evolution of the energy of the system, where times are indicated in correspondence with the snapshots of the density profile.
@@
   \label{fig:combined_1}
-\end{center}
+\end{figure}
 
-\begin{center}
+\begin{figure}
   \includegraphics[width=0.9\textwidth,trim=0 0 2cm 2cm,clip]{energy_diagram2}
-  \captionof{figure}{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}
+  \caption{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}
   \label{fig:binary}
-\end{center}
+\end{figure}
 
@@
 
-We have demonstrated that P\textsc{y}U\textsc{ltra}L\textsc{ight} is an accurate tool for the study of the dynamics of ultra-light dark matter governed by the Schr{\"o}dinger-Poisson system of equations.
+We have demonstrated that \PyUltraLight is an accurate tool for the study of the dynamics of ultralight dark matter governed by the Schr{\"o}dinger-Poisson system of equations.
 The code makes use of a pseudo-spectral symmetrised split-step Fourier methodology, in which all spatial derivatives are treated via explicit multiplication in the Fourier domain, thereby avoiding difficulties associated with finite-differencing methods.
 
-Energy conservation within P\textsc{y}U\textsc{ltra}L\textsc{ight} is excellent, at sub-percent level for simulations run at $128^3$, and even better performance as resolution is increased.
-The code is able to reproduce complex phenomena resulting from the wave-like properties of ultra-light dark matter as predicted by theoretical models, such as the interference patterns arising during high-velocity collisions of solitonic cores, and the effective forces observed in cases where the colliding cores are out of phase.
+Energy conservation within \PyUltraLight is excellent, at sub-percent level for simulations run at $128^3$, and even better performance as resolution is increased.
+The code is able to reproduce complex phenomena resulting from the wave-like properties of ultralight dark matter as predicted by theoretical models, such as the interference patterns arising during high-velocity collisions of solitonic cores, and the effective forces observed in cases where the colliding cores are out of phase.
 These phenomena can be clearly observed at relatively low spatial resolution, obviating the need for high-performance computing infrastructure to study the fundamental behaviour of ULDM systems.
-This makes P\textsc{y}U\textsc{ltra}L\textsc{ight} a useful tool for predicting the overall dynamics of a ULDM system in a computationally efficient manner prior to a rigorous examination using more advanced codes such as Nyx.
+This makes \PyUltraLight a useful tool for predicting the overall dynamics of a ULDM system in a computationally efficient manner prior to a rigorous examination using more advanced codes such as Nyx.
 
-P\textsc{y}U\textsc{ltra}L\textsc{ight} is Python-based, and as such is particularly simple to understand and use.
+\PyUltraLight is Python-based, and as such is particularly simple to understand and use.
 The accompanying Jupyter notebook allows for the efficient adjustment of simulation parameters, and offers a useful browser interface for quick visualisation of simulation results.
@@
 
-While the current implementation of P\textsc{y}U\textsc{ltra}L\textsc{ight} is already a useful tool for simulating dynamical ULDM systems, there remains scope for substantial improvement.
+While the current implementation of \PyUltraLight is already a useful tool for simulating dynamical ULDM systems, there remains scope for substantial improvement.
 In particular, it is anticipated that future releases may incorporate adaptive mesh refinement capabilities and variable timestep functionality.

### 2018-05-17 draft.tex
Minor formatting change
Rearranged sections as per RE's suggestion. Changes to the beginning of the Testing/Verification section.

@@
 \newcommand{\mnras}{Mon. Notices Royal Astron. Soc}
-\newcommand{\PyUltraLight}{\textsc{PyUltraLight}}
+\newcommand{\PyUltraLight}{\textsc{PyUltraLight }}
 
@@
 
-\section{The Physics of ULDM: The Schr{\"o}dinger-Poisson System}\label{sec:physics}
+\section{The Physics of ULDM}\label{sec:physics}
+
+\subsection{The Schr{\"o}dinger-Poisson System}
 
@@
 
-\section{Algorithm and Implementation}\label{sec:implementation}
-
-Summary of this section TBW
-
 \subsection{Field Rescalings}
@@
 Henceforth, we will often refer to the quantity $\vert\psi\vert^2$ simply as the density, where it is understood that this is in fact a dimensionless quantity related to the physical mass density through a constant of proportionality given by equation \ref{eq:density}.
+
+\section{Algorithm and Implementation}\label{sec:implementation}
+
+In this section we discuss the methodology used to calculate the dynamics of the adimensional SP system (equations \ref{eq:s-adim} and \ref{eq:p-adim}) given user-specified initial conditions.
+We introduce the symmetrised split-step Fourier method, and describe schematically how the system is evolved at each timestep.
 
@@
 
-\section{Testing and Verification}\label{sec:test}
+\section{Verification of ULDM Dynamics}\label{sec:test}
+
+In this section we test the validity of \PyUltraLight as a numerical SP solver by reproducing a number of results from previous studies of ULDM dynamics.
+We are able to reproduce phenomena arising due to the wavelike nature of ULDM, namely interference effects and effective repulsive forces.
+Additionally, we study the evolution of the velocity field within a solitonic core orbiting within a Newtonian central potential, showing that in the absence of complete tidal disruption, the stable orbital configuration is one of an irrotational Riemann-S ellipsoid.
+Finally, we demonstrate sub-percent level energy conservation for a selection of dynamical scenarios.
 

### 2018-05-19 draft.tex
Additional detail regarding Riemann S ellipsoid

@@
 % more complex case: 4 authors, 3 institutions, 2 footnotes
-\author[a,b,1]{F. Irst,\note{Corresponding author.}}
-\author[c]{S. Econd,}
-\author[a,2]{T. Hird\note{Also at Some University.}}
-\author[a,2]{and Fourth}
+\author[]{Faber Edwards,}
+\author[]{Emily Kendall,}
+\author[]{Shaun Hotchkiss,}
+\author[]{and Richard Easther}
 
@@
 
-\affiliation[a]{One University,\\some-street, Country}
-\affiliation[b]{Another University,\\different-address, Country}
-\affiliation[c]{A School for Advanced Studies,\\some-location, Country}
+\affiliation[]{Department of Physics, University of Auckland, Private Bag 92019, Auckland, New Zealand}
 
 % e-mail addresses: one for each author, in the same order as the authors
-\emailAdd{first@one.univ}
-\emailAdd{second@asas.edu}
-\emailAdd{third@one.univ}
-\emailAdd{fourth@one.univ}
+\emailAdd{faberedwards@gmail.com}
+\emailAdd{eken000@aucklanduni.ac.nz}
+\emailAdd{s.hotchkiss@auckland.ac.nz}
+\emailAdd{r.easther@auckland.ac.nz}
 
@@
 An extensive study of the tidal disruption of ULDM solitonic cores orbiting a central potential has recently been undertaken in \cite{Du2018}.
-Here we reproduce just one of their results as a validity check for \PyUltraLight.
-In particular, it is found in \cite{Du2018} that an initially spherical solitonic core without self-rotation will gradually spin up to form a tidally-locked irrotational Riemann -S ellipsoid when orbiting a host mass.
-Figure \ref{fig:riemann} shows the internal velocity field of a solitonic satellite after one complete revolution around a host mass, obtained using \PyUltraLight on a $256^3$ simulation grid.
+Here, we reproduce just one of their results as an additional validity check for \PyUltraLight.
+In particular, we reproduce the characteristic internal velocity field of a tidally-locked soliton orbiting a central potential.
+
+When the SP system is recast as a system of hydrodynamical equations, the fluid velocity is given by the gradient of the smoothly-varying phase at any point.
+This implies that the velocity flow of the SP system is irrotational, $\Nabla\times\vec{v}=0$.
+When a soliton is initialised in a circular orbit around a Newtonian potential, the initially spherical profile is distorted, elongated along the radial direction of the central potential.
+Meanwhile, the velocity field corresponding to the overall orbital motion of the soliton is superposed with the internal velocity field, combining so as to produce a net flow with vanishing curl.
+The family of Riemann-S ellipsoids describe uniformly rotating bodies which satisfy the condition of irrotationality \cite{Chandrasekhar1965}, and it is therefore the characteristic internal velocity field of such an ellipsoid which we expect to arise during the simulation.
+It is found in \cite{Du2018} that an initially spherical solitonic core without self-rotation will gradually spin up to form a tidally-locked irrotational Riemann-S ellipsoid when orbiting a host mass.
+We are able to reproduce this result using \PyUltraLight, with Figure \ref{fig:riemann} showing the internal velocity field of a solitonic satellite after one complete revolution around a host mass, obtained using \PyUltraLight on a $256^3$ simulation grid.
 In this case the ratio of host to satellite mass was $\approx 55$.

### 2018-05-24 draft.tex
Updated energy conservation Figure and explanation of small-scale oscillations. Added download section.

@@
 
-Figures \ref{fig:combined_1} and \ref{fig:binary} demonstrate energy conservation for two possible simulation scenarios.
-In the first case, we examine the evolution of the energy for a single soliton undergoing significant tidal disruption whilst orbiting within a central potential.
+Figures \ref{fig:combined_1} and \ref{fig:binary} demonstrate energy conservation for two example simulation scenarios.
+In the first case, we examine the evolution of the energy for a single soliton undergoing significant tidal disruption within a Newtonian central potential.
 We see that as the soliton is disrupted, the kinetic energy increases, while the gravitational energy due to the central potential decreases, as expected.
@@
 At points of closest approach, it can be clearly seen that the kinetic energy increases as the solitons speed up, while the potential energy due to self-interaction decreases commensurately.
-In this scenario no central potential has been included, so the gravitational energy due to the central potential stays uniformly zero throughout the simulation.
-The small scale oscillations that can be observed are artifacts of the relatively low spatial resolution, and smooth out as grid resolution is increased.
+In this scenario no central potential has been included, so the gravitational energy due to the central potential stays uniformly zero throughout the simulation and is not plotted in the Figure.
+It can also be seen that as the solitons reach the first point of closest approach, they become slightly deformed, exciting oscillatory modes which are manifest in the Figure as small scale oscillations superposed on the global behaviour.
 In this case the total energy is conserved to $10^{-3}\ \%$.
@@
 \begin{figure}
-  \includegraphics[width=0.9\textwidth,trim=0 0 2cm 2cm,clip]{energy_diagram2}
+  \includegraphics[width=0.9\textwidth,trim=0 0 2cm 2cm,clip]{egy_m=22.eps}
   \caption{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}
@@
 
-To be written.
+\PyUltraLight has been made publicly available under a BSD licence.
+The full code repository, including supplementary files such as the code used to generate soliton profiles, is available on GitHub at https://github.com/erckendall/PyUltraLight.
+\PyUltraLight makes use of the pyFFTW pythonic wrapper around the FFTW C-based fast Fourier transform libraries.
+Both pyFFTW and FFTW are freely-available.
+\PyUltraLight has been successfully tested on both Mac OS and Linux operating systems, though not exhaustively.
+The current version of \PyUltraLight, as demonstrated in this work, is capable of reproducing many of the results of previous ULDM simulations, and we plan to extend its capabilities in future releases.
 
@@
 
-To be written.
+TBW
 

### 2018-05-27 draft.tex
small changes from Richard

@@
 \usepackage{eucal}
+\usepackage{xspace}
+\usepackage{color}
 
@@
 \newcommand{\mnras}{Mon. Notices Royal Astron. Soc}
-\newcommand{\PyUltraLight}{\textsc{PyUltraLight }}
+\newcommand{\PyUltraLight}{\textsc{PyUltraLight}\xspace}
+\newcommand{\re}[1]{\textcolor{blue}{[{\bf RE}: #1]}}
 
-\abstract{\PyUltraLight\ simulates the dynamics of ultralight dark matter in a static, non-expanding background.
-\PyUltraLight\  can describe the evolution of several interacting ultralight dark matter halos  or one or more halos orbiting a central, fixed Newtonian potential, the latter scenario corresponding to dwarf galaxies orbiting a massive central galaxy.
-We verify  \PyUltraLight\  by showing that it reproduces qualitative dynamical features of previously published simulations and demonstrate that it has  excellent energy-conservation properties.
-\PyUltraLight\  is implemented in a Python-based Jupyter notebook, solving the Schr{\"o}dinger-Poisson equation  governing  ultralight scalar field dark matter dynamics in the non-relativistic regime using a symmetrised split-step pseudo-spectral  algorithm.
+\abstract{\PyUltraLight  simulates the dynamics of ultralight dark matter in a static, non-expanding background.
+\PyUltraLight can describe the evolution of several interacting ultralight dark matter halos  or one or more halos orbiting a central, fixed Newtonian potential, the latter scenario corresponding to dwarf galaxies orbiting a massive central galaxy.
+We verify  \PyUltraLight  by showing that it reproduces qualitative dynamical features of previously published simulations and demonstrate that it has  excellent energy-conservation properties.
+\PyUltraLight is implemented in a Python-based Jupyter notebook, solving the Schr{\"o}dinger-Poisson equation  governing  ultralight scalar field dark matter dynamics in the non-relativistic regime using a symmetrised split-step pseudo-spectral  algorithm.
 The notebook interface makes it simple to specify simulation parameters and  visualise the resulting output.
-\PyUltraLight\  runs on standard desktop hardware and is available on GitHub.
+\PyUltraLight  runs on standard desktop hardware and is available on GitHub.
 }
@@
 The realisation that there may be more to the universe than meets the eye is one of the most profound developments in 20$^{\textrm{th}}$ Century astronomy and astrophysics.
-However,  while there are now multiple lines of evidence that dark matter outweighs baryonic matter by a ratio of approximately 5:1 [CITE], we have few clues regarding the physical nature of dark matter.
-Much theoretical and experimental effort has focussed on  WIMP models, motivated by their consistency with supersymmetric extensions to the Standard Model and their relatively simple dynamics but advanced direct-detection experiments are putting increasingly tight constraints on the WIMP parameter space \cite{Tan2016, Akerib2017}.
-Moreover, $\Lambda$CDM cosmology with simple, pressureless, noninteracting dark matter (a class which includes simple WIMP models) is potentially at odds with observations at small astrophysical scales \cite{Bull2016}.
+However,  while there are now multiple lines of evidence that dark matter outweighs baryonic matter by a ratio of approximately 5:1 \re{Add a citation to this}, we have few clues regarding the physical nature of dark matter.
+Much theoretical and experimental effort has focussed on  WIMP models, motivated by their consistency with supersymmetric extensions to the Standard Model and their relatively simple dynamics.
+However but advanced direct-detection experiments are putting increasingly tight constraints on the WIMP parameter space \cite{Tan2016, Akerib2017} and $\Lambda$CDM cosmology with simple, pressureless, noninteracting dark matter (a class  including simple WIMP scenarios) is potentially at odds with observations at small astrophysical scales \cite{Bull2016}.
 
-[TO HERE]The potential shortcoming of simple cold dark matter scenarios are increasingly motivating the investigations of more novel dark matter scenarios.
-In particular, ultralight dark matter (ULDM) is an increasingly well-studied possibility, also known as fuzzy dark matter (FDM), or BEC dark matter.
-ULDM models are well motivated by fundamental theories possessing approximate shift symmetries, and offer natural resolutions to the small-scale problems of $\Lambda$CDM without the need to invoke baryonic physics.
-This is essentially because below the de Broglie wavelength of the ULDM particle, the Heisenberg uncertainty principle suppresses gravitational collapse.
-Since the mass of the ULDM particle is postulated to be of order $10^-22$ eV, this amounts to suppression of gravitational collapse on kilo parsec scales \cite{Hu2000}.
-A recent review of ULDM/FDM is given in \cite{Hui2016}.
+The potential shortcoming of simple cold dark matter scenarios   motivate  investigations of more novel dark matter scenarios.
+In particular, ultralight dark matter (ULDM), also known as fuzzy dark matter (FDM), or BEC dark matter, is an increasingly well-studied possibility; for a recent review of the potential advantages and characteristic attributes of this scenario see Ref~\cite{Hui2016}.
+ULDM models are well motivated by fundamental theories possessing approximate shift symmetries.
+Moreover, ULDM can naturally resolve the small-scale problems of $\Lambda$CDM as the Heisenberg uncertainty principle suppresses gravitational collapse on lengthscales shorter than the de Broglie wavelength of the ULDM particle.
+In this regime the mass of the ULDM particle becomes correlated with astrophysical observables; if it is on the order of $10^{-22}$~eV, structure is suppressed at kiloparsec scales and below \cite{Hu2000}.
 
-The computational modelling of ULDM systems in the non-relativistic regime can be distilled into numerically solving the dynamical Schr{\"o}dinger-Poisson (SP) system of coupled differential equations for given initial conditions.
+\re{read to here}The computational modelling of ULDM systems in the non-relativistic regime can be distilled into numerically solving the dynamical Schr{\"o}dinger-Poisson (SP) system of coupled differential equations for given initial conditions.
 Approaches to this problem have been varied, and have given rise to modifications of existing cosmological simulation codes as well as the development of new codes specifically for the modelling of ULDM systems, though public releases of the latter are  lacking.

### 2018-05-28 draft.tex
Added citation as per RE's comment. Edited JHEP.bst to correctly display arXiv references.

@@
 The realisation that there may be more to the universe than meets the eye is one of the most profound developments in 20$^{\textrm{th}}$ Century astronomy and astrophysics.
-However,  while there are now multiple lines of evidence that dark matter outweighs baryonic matter by a ratio of approximately 5:1 \re{Add a citation to this}, we have few clues regarding the physical nature of dark matter.
-Much theoretical and experimental effort has focussed on  WIMP models, motivated by their consistency with supersymmetric extensions to the Standard Model and their relatively simple dynamics.
-However but advanced direct-detection experiments are putting increasingly tight constraints on the WIMP parameter space \cite{Tan2016, Akerib2017} and $\Lambda$CDM cosmology with simple, pressureless, noninteracting dark matter (a class  including simple WIMP scenarios) is potentially at odds with observations at small astrophysical scales \cite{Bull2016}.
+However, while there are now multiple lines of evidence that dark matter outweighs baryonic matter by a ratio of approximately 5:1 \cite{Planck2015}, we have few clues regarding the physical nature of dark matter.
+Much theoretical and experimental effort has focused on  WIMP models, motivated by their consistency with supersymmetric extensions to the Standard Model and their relatively simple dynamics.
+However, advanced direct-detection experiments are putting increasingly tight constraints on the WIMP parameter space \cite{Tan2016, Akerib2017} and $\Lambda$CDM cosmology with simple, pressureless, noninteracting dark matter (a class including simple WIMP scenarios) is potentially at odds with observations at small astrophysical scales \cite{Bull2016}.
 
-The potential shortcoming of simple cold dark matter scenarios   motivate  investigations of more novel dark matter scenarios.
+The potential shortcomings of simple cold dark matter scenarios motivate investigations of more novel dark matter scenarios.
 In particular, ultralight dark matter (ULDM), also known as fuzzy dark matter (FDM), or BEC dark matter, is an increasingly well-studied possibility; for a recent review of the potential advantages and characteristic attributes of this scenario see Ref~\cite{Hui2016}.
 ULDM models are well motivated by fundamental theories possessing approximate shift symmetries.
-Moreover, ULDM can naturally resolve the small-scale problems of $\Lambda$CDM as the Heisenberg uncertainty principle suppresses gravitational collapse on lengthscales shorter than the de Broglie wavelength of the ULDM particle.
+Moreover, ULDM can naturally resolve the small-scale problems of $\Lambda$CDM as the Heisenberg uncertainty principle suppresses gravitational collapse on length scales shorter than the de Broglie wavelength of the ULDM particle.
 In this regime the mass of the ULDM particle becomes correlated with astrophysical observables; if it is on the order of $10^{-22}$~eV, structure is suppressed at kiloparsec scales and below \cite{Hu2000}.
 
-\re{read to here}The computational modelling of ULDM systems in the non-relativistic regime can be distilled into numerically solving the dynamical Schr{\"o}dinger-Poisson (SP) system of coupled differential equations for given initial conditions.
+\re{read to here}The computational modelling of ULDM systems in the non-relativistic regime can be distilled into the problem of numerically solving the dynamical Schr{\"o}dinger-Poisson (SP) system of coupled differential equations for given initial conditions.
 Approaches to this problem have been varied, and have given rise to modifications of existing cosmological simulation codes as well as the development of new codes specifically for the modelling of ULDM systems, though public releases of the latter are  lacking.
@@
 %\section*{References}
-\bibliographystyle{JHEP}
+\bibliographystyle{JHEP-mod}
 \bibliography{refs}

### 2018-05-28 draft.tex
Richard, through to p5

@@
 We verify  \PyUltraLight  by showing that it reproduces qualitative dynamical features of previously published simulations and demonstrate that it has  excellent energy-conservation properties.
-\PyUltraLight is implemented in a Python-based Jupyter notebook, solving the Schr{\"o}dinger-Poisson equation  governing  ultralight scalar field dark matter dynamics in the non-relativistic regime using a symmetrised split-step pseudo-spectral  algorithm.
+\PyUltraLight is implemented in a Python-based Jupyter notebook, solving the Schr{\"o}dinger-Poisson equation  governing  ultralight scalar field dark matter dynamics in the non-relativistic regime using a symmetrised split-step pseudospectral  algorithm.
 The notebook interface makes it simple to specify simulation parameters and  visualise the resulting output.
@@
 
-\re{read to here}The computational modelling of ULDM systems in the non-relativistic regime can be distilled into the problem of numerically solving the dynamical Schr{\"o}dinger-Poisson (SP) system of coupled differential equations for given initial conditions.
-Approaches to this problem have been varied, and have given rise to modifications of existing cosmological simulation codes as well as the development of new codes specifically for the modelling of ULDM systems, though public releases of the latter are  lacking.
-One approach has been to use the Madelung fluid formulation of the SP system \cite{Madelung1926}, in which there arises a quantum pressure term which can be treated numerically in a variety of ways.
-In \cite{Zhang2018}, the cosmological code \textsc{gadget} \cite{Springel2005} has been modified in order to treat this quantum pressure using an effective particle-particle interaction.
-This modified code, \textsc{axion-gadget}, has been made publicly available \cite{axion-gadget}.
-In \cite{Nori2018}, a non-public extension of \textsc{gadget}, \textsc{p-gadget3}, has been modified to treat the quantum pressure term using smoothed-particle hydrodynamics (SPH) routines.
-The SPH approach has also been used in \cite{Mocz2015}, while a particle-mesh approach was implemented in \cite{Veltmaat2016} to treat the quantum pressure term.
-Other cosmological simulation codes which have been recruited to solve ULDM dynamics (though not necessarily using the Madelung formulation) are N\textsc{yx} \cite{Almgren2013}, G\textsc{alacticus} \cite{Benson2012}, \textsc{arepo} \cite{Springel2010}, and \textsc{gamer} \cite{Schive2010, gamer}.
-N\textsc{yx} was modified in \cite{Schwabe2016} to study merging ULDM solitonic cores, G\textsc{alacticus} was modified in \cite{Du2017} to study the effects of tidal stripping and dynamical friction on ULDM halos, \textsc{arepo} was modified in \cite{Mocz2017} to study the core-mass relationship and turbulence characteristics of ULDM halos, and \textsc{gamer} was modified in \cite{Schive2014_b} to perform a detailed study of structure formation in ULDM cosmologies.
+%\re{read to here}
+Given the presence of a fundamental  lengthscale, the behaviour of ULDM is more complex than that of simple dark matter scenarios whose cosmologically relevant interactions are purely gravitational.
+Physically, the effective short-scale pressure and  condensate-like properties of ULDM  create new dynamical possibilities for ULDM scenarios, such as  purely pressure supported soliton-like solutions  \cite{Marsh2015} and superposition or interference during interactions between condensate-like halos \cite{Schwabe2016}.
+Consequently, modelling dark matter dynamics in ULDM scenarios is more challenging than in simple cold dark matter models, but is critical to understanding the physical consequences of ULDM models.
 
-While these modified cosmological simulation codes have produced promising results in the study of ULDM physics, at the time of writing only one such code, \textsc{axion-gadget}, has been made publicly available.
-The aim of this paper is to introduce \PyUltraLight, a stand-alone Python-based pseudo-spectral SP solver, and to demonstrate that this code is able to reproduce a number of key findings from the more complicated cosmological simulation codes without the need for high-performance computing environments.
-Consequently, we anticipate that as a publicly available resource, \PyUltraLight\ will serve as a valuable cross-check for more complex implementations and may serve as a basis for further development of such codes within the computational cosmology community.
+In the non-relativistic regime, the dynamics of ULDM can be reduced to the Schr{\"o}dinger-Poisson  system, where the complex variable $\psi$ describes the local density of ULDM quanta while the Poisson equation describes the local gravitational potential.
+Many approaches have been taken to this problem, including both modifications of existing cosmological simulation codes and the development of new codes specifically designed for ULDM systems.
+One widely used approach is the Madelung fluid formulation of the Schr{\"o}dinger-Poisson system \cite{Madelung1926} which has a quantum pressure term that can be treated numerically in a variety of ways.
+In Ref.~\cite{Zhang2018}, the cosmological code \textsc{gadget} \cite{Springel2005} is modified to treat the quantum pressure as an effective particle-particle interaction and the resulting code, \textsc{axion-gadget} is publicly available \cite{axion-gadget}.
+Ref.~\cite{Nori2018} modifies a non-public extension of \textsc{gadget}, \textsc{p-gadget3} to treat the quantum pressure term via smoothed-particle hydrodynamics (SPH) routines.
+The SPH approach is also  used in Ref.~\cite{Mocz2015}, while a particle-mesh approach was implemented in \cite{Veltmaat2016}.
+N\textsc{yx} \cite{Almgren2013} was modified in \cite{Schwabe2016} to study merging ULDM solitonic cores, G\textsc{alacticus} \cite{Benson2012} was modified in \cite{Du2017} to study the effects of tidal stripping and dynamical friction on ULDM halos, \textsc{arepo} \cite{Springel2010} was modified in \cite{Mocz2017} to study the core-mass relationship and turbulence characteristics of ULDM halos, and \textsc{gamer} \cite{Schive2010, gamer} was modified \cite{Schive2014_b} to perform a detailed study of structure formation in ULDM cosmologies.
 
-\PyUltraLight\ makes use of a symmetrised-split-step pseudo-spectral Fourier algorithm (also known as leap-frog) to compute the dynamics of the SP system.
-We note that a similar methodology has been described in \cite{Paredes2016}, though to the authors' knowledge this code has not been released publicly.
-The symmetrised-split-step method offers $2^{nd}$ order accurate time integration steps and sub-percent level energy conservation while normalisation of the wavefunction is conserved to machine precision.
-As a pseudo-spectral code, linear differential operators are computed by direct multiplication in the Fourier domain, while non-linear terms are evaluated in position space.
-This means that \PyUltraLight\ is free from noise associated with finite-differencing methods of computing spatial derivatives, though there is a necessary computational cost associated with the Fourier and inverse Fourier transforms.
-These transforms are optimised in \PyUltraLight\ through the use of the pyFFTW pythonic wrapper around the C-based FFWTW subroutine library \cite{pyfftw} \cite{fftw}.
-While the code is currently implemented to run on a singe machine, the FFTW libraries offer full parallelisation, so there is no barrier to running the code in a cluster environment.
+While a large number of public codes can solve conventional dark matter scenarios, \textsc{axion-gadget} is the only currently available solver for ULDM dynamics.
+This paper introduces \PyUltraLight, a stand-alone Python-based pseudospectral Schr{\"o}dinger-Poisson solver, and  demonstrates that it reproduces many of the key findings of more complicated cosmological simulation codes within a desktop computing environment.
+We anticipate that as a publicly available resource, \PyUltraLight\ will serve as a valuable cross-check on more complex implementations, serve as a basis for further development of such codes within the computational cosmology community, and facilitate explorations of ULDM dynamics.
+
+\PyUltraLight is based on a symmetrised-split-step (leapfrog) solver for the time evolution, and uses a pseudospectral Fourier algorithm to solve the Poisson equation for the gravitational potential at each step.\footnote{A similar methodology was described in Ref.~\cite{Paredes2016}; to our knowledge this code has not been released.} This algorithm has $2^{nd}$ order accurate time integration steps and sub-percent level energy conservation, while the wavefunction normalisation  is conserved to machine precision.
+As a pseudospectral code, linear differential operators are computed by direct multiplication in the Fourier domain, while non-linear terms are evaluated in position space.
+Consequently, \PyUltraLight is free from noise associated with spatial derivatives computed via finite-differencing.
+There is a necessary computational cost associated with the Fourier and inverse Fourier transforms but these transforms are optimised in \PyUltraLight\ through the use of the pyFFTW pythonic wrapper around the C-based FFWTW subroutine library \cite{pyfftw,fftw}.
+\PyUltraLight is currently takes advantage of multiple cores on a singe machine but the FFTW libraries offer full parallelisation, so there is no barrier to running the code in a cluster environment.
+\re{ Will the ``top level" Python routines run in a full MPI environment; i.e.
+a cluster based shared-memory implementation?
+We would need this to make the full cluster-based approach work.}
 
 This paper is organised as follows.
-We first provide a short review of the physics of ULDM, including a derivation of the SP equations from an appropriate scalar-field Lagrangian.
-We then describe the details of the implementation of \PyUltraLight, before moving on to a testing and verification section, in which we reproduce a selection of results from a variety of recent ULDM simulations and discuss energy conservation within the code.
+We first provide a short review of ULDM physics, including a derivation of the Schr{\"o}dinger-Poisson equations from the underlying scalar-field Lagrangian.
+We then describe their implementation in \PyUltraLight, before moving on to the testing and verification Section in which we reproduce a selection of results from a variety of recent ULDM simulations and discuss energy conservation within the code.
 
@@
 The existence of an extremely light scalar field, minimally coupled to gravity, is the central premise on which ULDM models are predicated.
-Within the ULDM framework, it is proposed that large-scale cosmological structure formation proceeds through the Bose-Einstein condensation of this scalar field.
-In this section we derive the (non-relativistic) dynamical equations of the single macroscopic condensate wavefunction, namely, the Schr{\"o}dinger-Poisson equations.
-We begin by considering the following action functional for a scalar field, $\phi$, minimally coupled to gravity and in the absence of self-interactions:
+Within the ULDM framework, it is proposed that cosmological structure formation proceeds through the Bose-Einstein condensation of this scalar field.
+In this Section  \re{I would nornally give Section a capital letter in this context} we derive the (non-relativistic) dynamical equations of the single macroscopic condensate wavefunction, namely, the Schr{\"o}dinger-Poisson equations.
+We begin with the action functional for a scalar field, $\phi$, minimally coupled to gravity and in the absence of self-interactions,
+%
 \begin{equation}\label{eq:action}
@@
 where $H=\Dot{a}(t)/a(t)$.
-At late times, $H\ll m$ and it is sufficient to set $H=0$ and $a(t)=1$ in equation \ref{eq:linear}.
-In the non-relativistic limit, we may use the WKB approximation to obtain an ansatz solution to equation \ref{eq:linear}.
+We will work in the limit $H\ll m$ and it is sufficient to set $H=0$ and $a(t)=1$ in equation \ref{eq:linear}.
+\footnote{ \re{ Just checking -- we could say we were working in a  non-expanding universe and just set $H=0$ whatever $m$ was?
+Physically, what does this statement mean?
+Can we say when this approximation becomes good for $m=10^{-22}$eV dark matter?
+We go on to say non-relativistic, but is that synonymous with non-expanding?}}  In the non-relativistic limit,  the WKB approximation gives an ansatz for a  solution to equation \ref{eq:linear}.
 We take
@@
 \end{equation}
-where $\psi$ is a slowly varying complex function so that terms involving $\Ddot{\psi}$ may be neglected. We may also use $\nabla^2\phi\ll m^2/\hbar^2\phi$, such that our final result takes the form
+where $\psi$ is assumed to be slowly varying (with respect to intervals $\Delta t\sim \hbar/m$ \re{right?}) complex function so that terms involving $\Ddot{\psi}$ may be neglected.
+We may also use $\nabla^2\phi\ll m^2/\hbar^2\phi$, \re{physical origin of this} such that our final result takes the form
 \begin{equation}\label{eq:schrodinger}
@@
 \end{equation}
-Thus, by making suitable approximations we have derived the Schr{\"o}dinger equation for $\psi$, which we now interpret as the unique macroscopic wavefunction of a Bose-Einstein condensate.
-It follows that the particle density of the condensate is given by $\vert\psi\vert^2$, such that the mass density is simply $m\vert\psi\vert^2$.
-Using this, we may therefore write down the Poisson equation as
+We have thus show that $\psi$ satisfies the Schr{\"o}dinger equation in this limit, which is interpreted as the unique macroscopic wavefunction of a Bose-Einstein condensate.
+It follows that the particle density of the condensate is given by $\vert\psi\vert^2$, so its mass density is simply $m\vert\psi\vert^2$.
+The local gravitational potential thus satisfies the Poisson equation,
 \begin{equation}\label{eq:poisson}
@@
 
-It is helpful to recast the SP system (equations \ref{eq:schrodinger} and \ref{eq:poisson}) in terms of adimensional quantities.
-In keeping with \cite{Schive2014} and \cite{Paredes2016}, we introduce length, time, and mass scales as follows:
+It is helpful to recast the Schr{\"o}dinger-Poisson system (equations \ref{eq:schrodinger} and \ref{eq:poisson}) in terms of adimensional quantities.
+In keeping with Refs~\cite{Schive2014,Paredes2016} we introduce length, time, and mass scales as follows:
 \begin{align}
@@
 where $m$ is the mass of the ultralight scalar field, $H_0$ is the present-day Hubble parameter, $G$ is Newton's gravitational constant and $\Omega_{m_0}$ is the present-day matter fraction of the energy density of the universe.
-We now recast equations \ref{eq:schrodinger} and \ref{eq:poisson} in terms of the following dimensionless quantities:
+We  recast equations \ref{eq:schrodinger} and \ref{eq:poisson} in terms of the  dimensionless quantities
 \begin{equation}\label{eq:dimensionless}
@@
 \end{equation}
-Dropping the primes for notational convenience, we see that the coupled differential equations of the SP system are reduced to 
+Dropping the primes for notational convenience, we see that the coupled differential equations of the  Schr{\"o}dinger-Poisson system  reduce to 
 \begin{align}
@@
 where it is understood that all quantities involved are dimensionless.
-We can recover any dimensionful quantities of interest through simple transformations using equations \ref{eq:length} to \ref{eq:mass}.
+We can recover  dimensionful quantities via the ``dictionary'' provided by equations \ref{eq:length} to \ref{eq:mass}.
 For example, the integrated mass of the system, $M_{tot}$, is given by
@@
 \end{equation}
-All computations performed by \PyUltraLight\ are carried out using the dimensionless quantities defined in \ref{eq:dimensionless}, while provisions are made within the code for initial conditions and outputs to be converted to and from the desired unit system specified by the user.
-Henceforth, we will often refer to the quantity $\vert\psi\vert^2$ simply as the density, where it is understood that this is in fact a dimensionless quantity related to the physical mass density through a constant of proportionality given by equation \ref{eq:density}.
+\PyUltraLight\ works internally with these dimensionless quantities but receives initial conditions and generates output in physical units.
+Henceforth, we will often refer to $\vert\psi\vert^2$  as the density, where it is understood that this is in fact a dimensionless quantity related to the physical mass density via the constant of proportionality given by equation \ref{eq:density}.
 
@@
 
-In this section we discuss the methodology used to calculate the dynamics of the adimensional SP system (equations \ref{eq:s-adim} and \ref{eq:p-adim}) given user-specified initial conditions.
-We introduce the symmetrised split-step Fourier method, and describe schematically how the system is evolved at each timestep.
+In this section we discuss the methodology used to calculate the dynamics of the adimensional  Schr{\"o}dinger-Poisson system (equations \ref{eq:s-adim} and \ref{eq:p-adim}) given user-specified initial conditions.
+We introduce the symmetrised split-step Fourier method, and schematically describe  how the system is evolved at each timestep.
 
@@
 
-Dynamical evolution within \PyUltraLight\ progresses via a symmetrised split-step Fourier process on an $N\times N\times N$ simulation grid with spatial periodic boundary conditions.
-To understand this method, we first consider the exact expression for the unitary time evolution of the wavefunction according to equation \ref{eq:s-adim}, namely
+Dynamical evolution within \PyUltraLight\ progresses via a symmetrised split-step Fourier process on an $N\times N\times N$  grid with  periodic spatial boundary conditions.
+To understand this method,  first consider the exact expression for the unitary time evolution of the wavefunction according to equation \ref{eq:s-adim}, namely
 \begin{equation}\label{eq:exact-time-ev}
@@
 \end{equation}
-where $\mathcal{T}$ is the time-ordering symbol. For sufficiently small timestep $h$, we can use the trapezoidal rule make the approximation
+where $\mathcal{T}$ is the time-ordering symbol. For a sufficiently small timestep $h$,  the trapezoidal rule gives the approximation
 \begin{equation}
@@
 \end{equation}
-It is acknowledged that the exponential in equation \ref{eq:approx-time-ev} omits the time-ordering symbol, and is only equivalent to its time-ordered counterpart to order $h^2$.
+Note that the exponential in equation \ref{eq:approx-time-ev} omits the time-ordering symbol, and is only equivalent to its time-ordered counterpart to order $h^2$. \re{to here}
 
 Given that the linear differential operator in equation \ref{eq:approx-time-ev} acts naturally in Fourier space, while the nonlinear potential term is simplest to evaluate in position space, we are motivated to split the exponential so as to separate these terms.
-This will then allow us to employ a pseudo-spectral algorithm to compute the time evolution.
+This will then allow us to employ a pseudospectral algorithm to compute the time evolution.
 Such a splitting of the exponential is valid when the timestep $h$ is small, as the dispersive and nonlinear terms can be assumed to act independently.
@@
 When the SP system is recast as a system of hydrodynamical equations, the fluid velocity is given by the gradient of the smoothly-varying phase at any point.
-This implies that the velocity flow of the SP system is irrotational, $\Nabla\times\vec{v}=0$.
+This implies that the velocity flow of the SP system is irrotational, $\nabla\times\vec{v}=0$.
 When a soliton is initialised in a circular orbit around a Newtonian potential, the initially spherical profile is distorted, elongated along the radial direction of the central potential.
@@
 We have demonstrated that \PyUltraLight is an accurate tool for the study of the dynamics of ultralight dark matter governed by the Schr{\"o}dinger-Poisson system of equations.
-The code makes use of a pseudo-spectral symmetrised split-step Fourier methodology, in which all spatial derivatives are treated via explicit multiplication in the Fourier domain, thereby avoiding difficulties associated with finite-differencing methods.
+The code makes use of a pseudospectral symmetrised split-step Fourier methodology, in which all spatial derivatives are treated via explicit multiplication in the Fourier domain, thereby avoiding difficulties associated with finite-differencing methods.
 

### 2018-05-29 draft.tex
Richard, through to p5

@@
 where $G$ is Newton's gravitational constant.
-The coupled equations \ref{eq:schrodinger} and \ref{eq:poisson} together form the nonlinear Schr{\"o}dinger-Poisson (SP) system which describes the dynamics of ULDM in the non-relativistic regime.
+The coupled equations \ref{eq:schrodinger} and \ref{eq:poisson} together form the nonlinear Schr{\"o}dinger-Poisson system which describes the dynamics of ULDM in the non-relativistic regime.
 It is this system of equations which \PyUltraLight is designed to solve.
+\re{I try to avoid acronyms like SP; ULDM works since it replaces something very long, and CDM is common usage, but I think papers are more readable if they are used sparingly}
 
@@
 
-Given that the linear differential operator in equation \ref{eq:approx-time-ev} acts naturally in Fourier space, while the nonlinear potential term is simplest to evaluate in position space, we are motivated to split the exponential so as to separate these terms.
-This will then allow us to employ a pseudospectral algorithm to compute the time evolution.
-Such a splitting of the exponential is valid when the timestep $h$ is small, as the dispersive and nonlinear terms can be assumed to act independently.
-In this implementation, we split the exponential as follows:
+The linear differential operator in equation \ref{eq:approx-time-ev} acts naturally in Fourier space, while the nonlinear potential term is simplest to evaluate in position space and by splitting the exponential we can evaluate each term in its natural domain.
+For small timesteps we assume that the dispersive and nonlinear terms act independently,  adopting the specific evolution scheme
+%
 \begin{equation}\label{eq:split-step}
@@
 \end{equation}
+%
 This splitting can be understood thusly: first, a half timestep is taken in which only the nonlinear potential operator acts, followed by a full timestep in the linear term.
 The potential field is then updated, and a final half timestep in the nonlinear term is performed.
-We may use the Baker-Campbell-Hausdorff formula to express the product of exponentials in equation \ref{eq:split-step} as a single exponential.
-Keeping only terms to order $h^2$ we find:
+Using the Baker-Campbell-Hausdorff formula to express the product of exponentials in equation \ref{eq:split-step} as a single exponential and keeping only terms to order $h^2$ we find:
 \begin{equation}\label{eq:BCH}
@@
 \end{equation}
-Making use of the fact that $\Phi(\vec{x},t+h)\approx\Phi(\vec{x},t)+h\Dot{\Phi}(\vec{x},t)$, we see that the commutators in equation \ref{eq:BCH} cancel at $\mathcal{O}(h^2)$, such that this expression is equivalent to \ref{eq:approx-time-ev}, with the dominant error term appearing at $\mathcal{O}(h^3)$.
+Making use of the fact that $\Phi(\vec{x},t+h)\approx\Phi(\vec{x},t)+h\Dot{\Phi}(\vec{x},t)$ we see that the commutators in equation \ref{eq:BCH} cancel at $\mathcal{O}(h^2)$ and the expression matches \ref{eq:approx-time-ev}, with the dominant error term appearing at $\mathcal{O}(h^3)$.
 
-Evaluation of equation \ref{eq:split-step} within \PyUltraLight proceeds as follows:
+Evaluation of equation \ref{eq:split-step} within \PyUltraLight thus proceeds as follows:
 Initially, the nonlinear term acts in position space for one half-timestep.
-The result is then Fourier transformed, and a full timestep is taken in which the differential operator is applied in the Fourier domain.
+The result is  Fourier transformed, and a full timestep is taken with the differential operator  applied in the Fourier domain.
 The potential field is then updated in accordance with equation \ref{eq:p-adim}.
-An inverse Fourier transform is then performed, and a final half timestep is taken in which the updated nonlinear term acts in position space to give the new $\psi$ field configuration.
-This procedure is known as the symmetrised split-step Fourier method, and is used widely in fields such as nonlinear fiber optics \cite{Agrawal2013}.
-It can be represented schematically as follows:
+After an inverse Fourier transform a final half timestep is taken with the updated nonlinear term acting in position space to give the new $\psi$ field configuration.
+This procedure is known as the symmetrised split-step Fourier method, and  used widely in fields such as nonlinear fiber optics \cite{Agrawal2013}.
+\re{do want a couple of other examples/cites?}
+
+The algorithm can be represented schematically as
 \begin{equation}\label{eq:schematic-psi}
@@
 \end{equation}
-where the order of operations is read from right to left, $\CMcal{F}$ and $\CMcal{F}^{-1}$ represent the discrete Fourier transform and its inverse, and $k$ is the wavenumber in the Fourier domain.
-The potential field is updated following the inverse Fourier transform in equation \ref{eq:schematic-psi}, and this update procedure may be represented schematically as
+where the order of operations runs from right to left, $\CMcal{F}$ and $\CMcal{F}^{-1}$ denote the discrete Fourier transform and its inverse, and $k$ is the wavenumber in the Fourier domain.
+The potential field is updated following the inverse Fourier transform in equation \ref{eq:schematic-psi}, via
+%
 \begin{equation}\label{eq:schematic-phi}
@@
 \end{equation}
-where $\psi(\vec{x},t_{i})$ represents the field configuration at this intermediary point in the timestep.
-The final operation in equation \ref{eq:schematic-psi} changes only the phase of the $\psi$ field, and thus we could replace $\psi(\vec{x},t_{i})$ with $\psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)$ in equation \ref{eq:schematic-phi} with no change in meaning.
-
-In the \PyUltraLight implementation, we make use of an additional simplification of the symmetrised split-step Fourier method.
-Namely, consecutive half-steps in the nonlinear term may be combined into a single full step in each case such that when iterated, only the first and last operations involve half steps.
-This can be represented schematically as
+%
+where $\psi(\vec{x},t_{i})$ is the field configuration at this halfway point in the full timestep.
+The final operation in equation \ref{eq:schematic-psi}  only changes the phase of  $\psi$, so we could replace $\psi(\vec{x},t_{i})$ with $\psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)$ in equation \ref{eq:schematic-phi} with no change in meaning.
+\PyUltraLight makes an additional simplification ti the symmetrised split-step Fourier method as the consecutive half-steps in the nonlinear term may be combined into a single full step.
+Consequently, only the first and last operations involve actual half steps.
+Schematically this becomes
+%
 \begin{equation}
@@
 \end{equation}
-where the $\Phi$ field is updated at each step in accordance with equation \ref{eq:schematic-phi} and attention is drawn to the sign difference between the first and last operators.
+%
+where  $\Phi$  is updated at each step via equation \ref{eq:schematic-phi}; attention is drawn to the sign difference between the first and last operators.
 
-The discrete Fourier transform (DFT) and its inverse are handled within \PyUltraLight using pyFFTW.
-pyFFTW is a pythonic wrapper around the C-based FFTW subroutine library, and allows for efficient evaluation of both real and complex DFTs \cite{pyfftw} \cite{fftw} \cite{Frigo2005}.
-Multithreading is supported by pyFFTW, and is implemented within \PyUltraLight by using the multiprocessing package to determine the number of online CPUs and feed this directly to the threads argument of the pyfftw.FFTW class.
-
-In addition to the use of pyFFTW, performance within \PyUltraLight is further improved through the use of the NumExpr package, which provides routines for the efficient manipulation of array objects within the simulation using a vector-based virtual machine \cite{numexpr}.
+From a computational perspective, the numerical Fourier transforms are likely to be rate-limiting step in any pseudospectral code.
+In \PyUltraLight the discrete Fourier transform (DFT) and its inverse are implemented via pyFFTW, a pythonic wrapper for the C-based FFTW subroutine library which efficiently implements both real and complex DFTs \cite{pyfftw,fftw,Frigo2005}.
+This  allows  \PyUltraLight  to combine the flexibility of a notebook based modelling tool with the efficiency of a carefully tuned, compiled numerical library.
+FFTW is fully parallelised and its support for multithreading  is inherited  by pyFFTW and accessed within \PyUltraLight; the number of threads used by the pyfftw.FFTW class is determined by the  Python multiprocessing routines which are used to ascertain the number of available CPU cores.
+In addition, \PyUltraLight uses the NumExpr package to parallelise operations on array objects within the simulation \cite{numexpr}.
 
@@
 
-The initialisation procedure within \PyUltraLight can be easily augmented to accommodate the requirements of the user.
-In the current implementation, the initial field configuration is built by loading a NumPy array file encoding a solitonic solution to the Schr{\"o}dinger-Poisson system.
-Using this method, solitons may be placed at the desired locations within the simulation grid, with the corresponding mass, velocity, and phase parameters each specified by the user within the accompanying Jupyter notebook.
+\PyUltraLight specifies the initial dark matter configuration as a superposition of an arbitrary number of solitonic halos, with arbitrary (user-defined) velocities and phases.
+This is necessarily an idealisation, given that realistic dark matter halos will not map directly to the solitonic solutions, but it provides an excellent ``playground'' in which to explore ULDM dynamics, and the initialisation routines within \PyUltraLight can be easily augmented to accommodate a wider range of scenarios.
+The initial field configuration is built by loading a NumPy array file encoding a solitonic solution to the Schr{\"o}dinger-Poisson system and the corresponding position mass, velocity, and phase parameters each specified by the user within the accompanying Jupyter notebook.
+In practice, only a finite  range of halo masses can be supported within a given simulation -- the  radius of a solitonic halo is inversely proportional to it mass, so resolve a light halo interacting with a very massive halo would require an extremely fine spatial mesh.
+\re{is there a power to this, or $r \sim 1/M_{halo}$?} \re{to here}
 

### 2018-05-29 draft.tex
Richard, through to p8
Richard, through to section 4.2

@@
 
-%\re{read to here}
 Given the presence of a fundamental  lengthscale, the behaviour of ULDM is more complex than that of simple dark matter scenarios whose cosmologically relevant interactions are purely gravitational.
@@
 This procedure is known as the symmetrised split-step Fourier method, and  used widely in fields such as nonlinear fiber optics \cite{Agrawal2013}.
-\re{do want a couple of other examples/cites?}
+\re{do we want a couple of other examples/cites?}
 
@@
 The initial field configuration is built by loading a NumPy array file encoding a solitonic solution to the Schr{\"o}dinger-Poisson system and the corresponding position mass, velocity, and phase parameters each specified by the user within the accompanying Jupyter notebook.
-In practice, only a finite  range of halo masses can be supported within a given simulation -- the  radius of a solitonic halo is inversely proportional to it mass, so resolve a light halo interacting with a very massive halo would require an extremely fine spatial mesh.
-\re{is there a power to this, or $r \sim 1/M_{halo}$?} \re{to here}
 
-The solitonic solution to the SP system is a natural choice of initial field configuration; these stable, localised objects are ubiquitous in ULDM models \cite{Marsh2015}.
+In practice, only a finite  range of halo masses can be supported within a given simulation -- the  radius of a solitonic halo is inversely proportional to its mass, so resolving a light halo interacting with a very massive halo would require an extremely fine spatial mesh.
+However, \PyUltraLight also allows the user to specify a fixed, external potential which does not take part in the dynamics.
+At this point only a central $1/r$ potential is supported but this would be easily generalised.
+\re{new material.
+Also, do we need to say anything about what happens when the solitions cross into the ``next" box?}
 
-%In particular, it is proposed that ULDM halos exhibit Navarro-Frenk-White profiles at large radii, with solitonic cores at smaller radii.
-Because solitons have a constant central density, it is suggested that this offers a solution to the core-cusp problem of $\Lambda$CDM \cite{Marsh2015}.
-
-%Remove this section if we're going to talk about solitons merging to form NFW/soltion profiles, and compare to e.g. Schwabe.
-
-In this section, we discuss the method used to generate the soliton profile. Following \cite{Paredes2016}, we begin by assuming a spherically symmetric solution to the SP system:
+The soliton profile is found by imposing a spherically symmetric profile \cite{Paredes2016} where the radial profile is assumed to be time independent
+%
 \begin{equation}
@@
 \end{equation}
-where $r=\vert\vec{x}\vert$. Introducing $\tilde{\varphi}(r):=\varphi(r)+\beta$, equations \ref{eq:s-adim} and \ref{eq:p-adim} are reduced to
+where $r=\vert\vec{x}\vert$. Introducing $\tilde{\varphi}(r):=\varphi(r)+\beta$, equations \ref{eq:s-adim} and \ref{eq:p-adim}  reduce to
 \begin{align}
@@
 \end{align}
+%
 where primes denote derivatives with respect to $r$.
-To generate profiles for $\tilde{\varphi}(r)$ and $f(r)$ which satisfy these equations, we use a fourth-order Runge-Kutta algorithm, included as the auxiliary soliton\texttt{\_}solution.py program.
-We make the arbitrary choice $f(r=0)$=1, while the smoothness criterion demands that the first derivatives of $f(r)$ and $\tilde{\varphi}(r)$ vanish at the origin.
-We then search for solutions of $f(r)$ and $\varphi(r)$ which satisfy the boundary conditions $\operatorname{lim}(r\rightarrow\infty)\varphi(r)=0$ and $\operatorname{lim}(r\rightarrow\infty)f(r)=0$.
-To do this we use a shooting method, varying $\tilde{\varphi}(r=0)$ until we obtain a solution of $f(r)$ which approaches zero at the maximal specified radius, $r_m$.
+Note that this system contains no arbitrary constants, so the underlying profile will effectively be universal and \PyUltraLight loads a precomputed profile supplied with the code, rather than computing it from scratch each time the code is run.
+However, an auxiliary program {\sc soliton\texttt{\_}solution.py} is supplied with  \PyUltraLight; it uses a  fourth-order Runge-Kutta algorithm to solve the coupled profile equations.
+We set $\left.
+f(r) \right|_{r=0}$=1, while  smoothness requires that first derivatives of $f(r)$ and $\tilde{\varphi}(r)$ vanish at the origin.
+We then use the shooting method to search for solutions of $f(r)$ and $\varphi(r)$  satisfying the boundary conditions $\lim_{r\rightarrow\infty}\varphi(r)=0$ and $\lim_{r\rightarrow\infty}f(r)=0$, varying $\left.
+\tilde{\varphi}(r)\right|_{r=0}$ until we obtain a solution of $f(r)$ which approaches zero at the maximal specified radius, $r_m$.
 The value of $\beta$ is then calculated by assuming that $\varphi(r)$ goes as $-c/r$ at large radii, where $c$ is a constant.
@@
 We thus obtain the full solution $\psi(\vec{x},t)=e^{i\beta t}f(r)$.
-Having initially chosen $f(r=0)=1$, we may then generalise to $f(r=0)=\alpha$, where $\alpha$ is an arbitrary positive real number.
-It is easy to verify that if $e^{i\beta t}f(r)$ is a solution to the spherically symmetric SP system, then $g(r)$ is also a solution, where
+Having initially chosen $\left.
+f(r)\right|_{r=0}=1$, we may then generalise to $\left.
+f(r)\right|_{r=0}=\alpha$, where $\alpha$ is an arbitrary positive real number.
+It is easily verified that if $e^{i\beta t}f(r)$ is a solution to the spherically symmetric Schr{\"o}dinger-Poisson system, then $g(r)$ is also a solution, where
 \begin{equation}
@@
 \end{equation}
-Hence we obtain a family of spherically symmetric soliton solutions to the dimensionless SP system, for which the dimensionless soliton mass is proportional to $\sqrt{\alpha}$ and the full width at half maximum is proportional to $1/\sqrt{\alpha}$.
-Because the size of the soliton scales inversely with the mass, it is important to consider the maximum soliton mass which can be accurately modeled in a simulation with finite spatial resolution.
+We thus have a family of spherically symmetric soliton solutions to the dimensionless Schr{\"o}\-dinger-Poisson system; the dimensionless soliton mass is proportional to $\sqrt{\alpha}$ and the full width at half maximum is proportional to $1/\sqrt{\alpha}$.
+Since the size of the soliton scales inversely with the mass, the most massive soliton in the solution  puts a lower bound on the required spatial resolution.
 
-While the Schr{\"o}dinger equation is not trivially form invariant under Galilean boosts, we can enforce Galilean covariance by augmenting the solutions through the addition of a velocity-dependent phase factor.
-Namely,
+The Schr{\"o}dinger equation is not trivially form invariant under Galilean boosts but we can enforce Galilean covariance through the addition of a velocity-dependent phase factor,
+%
 \begin{equation}\label{eq:covariant-solution}
@@
 \end{equation}
-To construct the initial field configuration within \PyUltraLight, the NumPy array encoding the radial profile $f(r)$ for the $f(r=0)=1$ case is loaded, where the spatial resolution of this profile must agree with the expected spatial resolution within the main code.
-Equation \ref{eq:covariant-solution} is then used to augment the configuration such that the soliton is furnished with the position, mass, and velocity specified by the user in the accompanying Jupyter notebook.
+%
+To construct the initial field configuration,  \PyUltraLight, loads the NumPy array encoding the radial profile $f(r)$ for the $\left.
+f(r)\right|_{r=0}=1$ case.
+Equation \ref{eq:covariant-solution} is then used to transform the this solution into soliton(s) with user-specified values position, mass, and velocity specified, via the accompanying Jupyter notebook.
 The user may also add an additional constant phase factor if desired.
-
-While the current implementation of \PyUltraLight is designed such that the initial field configuration is built from solitons, it may be simply modified to take in alternative field configurations in future.
 
@@
 
-When using finite-differencing methods to solve certain partial differential equations, the causality-preserving Courant-Friedrichs-Lewy (CFL) condition on the size of the simulation timestep is a necessary condition for convergence \cite{Ajaib2013}.
-As such, this condition is often appears in numerical analysis of ULDM physics governed by the SP system, an example being \cite{Schwabe2016}.
-However, because \PyUltraLight does not use the method of finite differences to compute spatial derivatives, but rather computes them exactly via explicit multiplication in the Fourier domain, the CFL condition in its usual formulation is not applicable to this code.
+In many finite-differencing schemes for  partial differential equations the Courant-Friedrichs-Lewy (CFL) condition is a necessary condition for convergence \cite{Ajaib2013}, and puts an upper bound on the timestep as  function of the gridsize whose physical origin is the finite speed of propagating information  in the evolving system.
+Consequently, this condition is often cited in numerical analysis of ULDM physics governed by the Schr{\"o}dinger-Poisson system, see e.g.
+Ref.~\cite{Schwabe2016}.
+Strictly speaking, however, the CFL condition is applicable to codes using finite differencing methods for spatial derivatives, and is thus not applicable to pseudospectral codes like \PyUltraLight.
+We also note that the Poisson equation is acausal in that it responds globally and instantaneously to local changes in the matter distribution.
+However, the Schr{\"o}dinger-Poisson equation is an approximation derived  in the nonrelativistic limit so this is not a shortcoming in regimes where the underlying equations are physically valid.
 
-In place of the CFL condition, the constraint on the timestep within \PyUltraLight is set based on the fluid interpretation of the SP system. In dimensionless code units we can write 
+In place of the CFL condition, the \PyUltraLight timestep can be set via the fluid interpretation of the Schr{\"o}dinger-Poisson system. In dimensionless code units we can write 
 \begin{equation}
@@
 \end{equation}
-From here we can define a `maximum velocity' for the simulation grid.
-This is the velocity at which the fluid appears to move backward, which occurs when the phase change between successive grid points exceeds $\pi$.
-The timestep is then constrained such that a fluid travelling at this maximum grid velocity will traverse a distance of one grid space per timestep.
+We define a `maximum velocity' for the simulation grid at which point the fluid appears to move backward because the phase change between successive grid points exceeds $\pi$.
+An upper bound can then be placed on the timestep so that a fluid travelling at this maximum  velocity  traverse a distance of one grid space per timestep.
+In practice this can yield a very demanding constraint on the timestep and for many scenarios the actual velocity is far less than this hypothetical limit.
+Consequently, the timestep can be  significantly increased at the user's discretion but convergence testing on a case-by-case basis is strongly encouraged.
 
-This proves to be a highly restrictive constraint on the timestep, and can be relaxed significantly at the user's discretion for simulations in which the expected velocities within the system will not approach this maximum grid velocity.
-Convergence testing on a case-by-case basis is encouraged to determine the timestep constraint necessary for a given degree of accuracy.
+\section{ULDM Dynamics with \PyUltraLight}\label{sec:test}
 
-\section{Verification of ULDM Dynamics}\label{sec:test}
-
-In this section we test the validity of \PyUltraLight as a numerical SP solver by reproducing a number of results from previous studies of ULDM dynamics.
-We are able to reproduce phenomena arising due to the wavelike nature of ULDM, namely interference effects and effective repulsive forces.
-Additionally, we study the evolution of the velocity field within a solitonic core orbiting within a Newtonian central potential, showing that in the absence of complete tidal disruption, the stable orbital configuration is one of an irrotational Riemann-S ellipsoid.
-Finally, we demonstrate sub-percent level energy conservation for a selection of dynamical scenarios.
+In this section we validate \PyUltraLight a by reproducing results from previous studies of ULDM dynamics, demonstrating interference effects and effective repulsive forces arising from the wavelike nature of ULDM.
+In addition we study the evolution of the velocity field of a solitonic core orbiting within a Newtonian central potential, showing that the stable orbital configuration is an irrotational Riemann-S ellipsoid.
+Finally, we show demonstrate that \PyUltraLight delivers sub-percent level energy conservation for a selection of dynamical scenarios.
 
@@
 
-Hierarchical structure formation within the ULDM framework proceeds through both gravitational overdensity collapse as well as through sequential binary mergers of solitonic cores to form individual virialised objects \cite{Schive2014}.
-It is therefore important that ULDM simulations are able to accurately reproduce the expected characteristics of binary soliton interactions.
+\begin{figure}
+  \includegraphics[trim={0 0 0 0.9cm},clip, scale=0.9]{interference_patterns}
+  \caption{Comparison of theoretical and numerical density profiles at time of maximal interference for head-on collision of two solitons with mass ratio $\mu=2$ and no relative phase difference.
+The solitons were chosen with dimensionless masses 5 and 10, with an initial separation of 4 code units and relative velocity of 20 code units.
+The simulation was run at $256^3$ for a box of side-length 8 code units.
+\re{do we want to say something about how a code unit is defined for a velocity?}}
+  \label{fig:interference}
+\end{figure}
 
-Soliton collisions can be classified into two categories depending on whether the total energy of the isolated binary system is positive or negative.
-In the case of positive energy, the solitons pass through each other, emerging largely undisturbed from their initial configurations following the interaction.
-During the collision, the wavefunctions describing the solitons are superposed, yielding distinctive interference patterns.
-In this section we demonstrate the ability of \PyUltraLight to reproduce the interference patterns predicted by theoretical models during binary collisions.
+The outcomes of ULDM soliton collisions depend critically on  whether the total energy of the isolated binary system is positive or negative.
+With a positive total energy the solitons pass through each other, emerging largely undisturbed from their initial configurations and the wavefunctions describing the solitons are superposed during the collision,, yielding distinctive interference patterns.
 
 Following \cite{Schwabe2016}, we consider the head-on collision of two solitions with mass ratio $\mu=2$ and high relative velocity.
-Using equation \ref{eq:covariant-solution}, we can write the total wavefunction of the binary system in terms of dimensionless quantities along the collision axis as
+This simple case can be treated approximately.
+Starting from equation \ref{eq:covariant-solution} we  write the total wavefunction of the binary system in terms of dimensionless quantities along the collision axis as
 \begin{align}
@@
 \end{align}
-where $x_1$ and $x_2$ are the initial positions of the centres of the solitions, $v_1$ and $v_2$ are the velocities of the solitons, $\delta$ is a constant relative phase term, and $\alpha_1=1/4 \ \alpha_2$ parameterise the density profiles as discussed in section \ref{sec:soliton-profiles}.
-Because of the symmetry of the configuration, we take $v_1=-v_2$ and $x_1=-x_2$.
-We are interested in the density profile of the system at the time of maximal interference, where we expect the superposed wavefunction to possess pronounced features.
-Therefore, we choose the time at which the two components of the wavefunction are centred at the same location, such that $x_1+v_1t=-x_2-v_2t=0$.
-This corresponds to a time $t_{max}=\vert x_1/v_1\vert = \vert x_2/v_2\vert$.
+where $x_1$ and $x_2$ are the initial centrals positions of the solitions, $v_1$ and $v_2$ are the soliton velocities, $\delta$ is a constant relative phase term and $\alpha_1=1/4 \ \alpha_2$, parameterising the density profiles as discussed in section \ref{sec:soliton-profiles}.
+For convenience we set $v_1=-v_2$ and $x_1=-x_2$.
+We expect that the interference effects will be maximised when two components of the wavefunction are centred at the same location, such that $x_1+v_1t=-x_2-v_2t=0$.
+This corresponds to a time $t_{o}=\vert x_1/v_1\vert = \vert x_2/v_2\vert$.
 The dimensionless density is then given by
 \begin{align}\label{eq:predicted-interference}
-    \vert\psi(x,t_{max})\vert^2= \ \alpha_1^2\bigg[&f(\sqrt{\alpha_1}x)^2+16f(2\sqrt{\alpha_1}x)^2+\nonumber\\
+    \vert\psi(x,t_{o})\vert^2= \ \alpha_1^2\bigg[&f(\sqrt{\alpha_1}x)^2+16f(2\sqrt{\alpha_1}x)^2+\nonumber\\
     &8f(\sqrt{\alpha_1}x)f(2\sqrt{\alpha_1}x)\operatorname{cos}\left(-3\alpha_1\beta\left\vert\frac{x_1}{v_1}\right\vert+2v_1x+\delta \right)\bigg]
 \end{align}
-In Figure \ref{fig:interference}, we plot the dimensionless density profile at the time of maximal interference for two solitons with mass ratio $\mu=2$, and phase difference $\delta=0$.
-We see that the numerical result obtained using \PyUltraLight closely matches the theoretical prediction of equation \ref{eq:predicted-interference}.
-Small disparities between the numerical and theoretical profiles may be attributed to the effect of gravitational contraction not included in the theoretical prediction of equation \ref{eq:predicted-interference}, and to a small offset in the true time of maximal interference.
-Neglecting these small deviations, we can verify that the code is well suited to faithfully reproduce the wave interference effects of the ULDM model.
-\begin{figure}
-  \includegraphics[trim={0 0 0 0.9cm},clip, scale=0.9]{interference_patterns}
-  \caption{Comparison of theoretical and numerical density profiles at time of maximal interference for head-on collision of two solitons with mass ratio $\mu=2$ and no relative phase difference.
-The solitons were chosen with dimensionless masses 5 and 10, with an initial separation of 4 code units and relative velocity of 20 code units.
-The simulation was run at $256^3$ for a box of side-length 8 code units.}
-  \label{fig:interference}
-\end{figure}
+ Figure \ref{fig:interference} shows the dimensionless density profile at the time of maximal interference for two solitons with mass ratio $\mu=2$ and phase difference $\delta=0$.
+The numerical result obtained using \PyUltraLight closely matches the theoretical prediction of equation \ref{eq:predicted-interference}.
+Small disparities between the numerical and theoretical profiles may be attributed to the effect of gravitational contraction not included in the theoretical prediction of equation \ref{eq:predicted-interference} and to a small offset in the true time of maximal interference \re{due to the solitons accelerating as they fall together?}.
+We do not expect an exact match, but we have verified \PyUltraLight qualitatively reproduce the wave interference effects of the ULDM model.
+\re{are there numerical treaments elsewhere we can compare to?}
+
+\re{read to here}
 

### 2018-05-31 draft.tex
Edited the energy conservation image - now 96^3 rather than 64^3 at lowest resolution.

@@
 The outcomes of ULDM soliton collisions depend critically on  whether the total energy of the isolated binary system is positive or negative.
-With a positive total energy the solitons pass through each other, emerging largely undisturbed from their initial configurations and the wavefunctions describing the solitons are superposed during the collision,, yielding distinctive interference patterns.
+With a positive total energy the solitons pass through each other, emerging largely undisturbed from their initial configurations and the wavefunctions describing the solitons are superposed during the collision, yielding distinctive interference patterns.
 
@@
 
+\begin{figure}
+  \includegraphics[width=1.\textwidth, trim={0 0 0 0},clip]{phase_comparison}
+  \caption{Head-on collisions between two solitons of mass 5, relative velocity 6, and initial separation 4 in code units. Time progresses from left to right across each row.}
+  \label{fig:repulsion}
+\end{figure}
+
 As demonstrated in \cite{Paredes2016}, the wavelike properties of ULDM can give rise to effective forces which can dramatically affect the dynamics of core collisions.
@@
 Further discussion of this phenomenon and its possible observational consequences can be found in \cite{Paredes2016}.
-\begin{figure}
-  \includegraphics[width=1.\textwidth, trim={0 0 0 0},clip]{phase_comparison}
-  \caption{Head-on collisions between two solitons of mass 5, relative velocity 6, and initial separation 4 in code units. Time progresses from left to right across each row.}
-  \label{fig:repulsion}
-\end{figure}
 
@@
 \subsection{Tidal Disruption of Solitons Orbiting a Central Potential}\label{sec:disruption}
+
+\begin{figure}
+  \includegraphics[width=1.\textwidth,trim=1cm 1cm 0 1cm,clip]{riemann}
+  \caption{Solitonic core after one revolution of a central potential, demonstrating the internal velocity field of an irrotational Riemann-S ellipsoid (where the collective motion of the soliton has been subtracted from the velocity field).
+Deformation of the spherical density profile into that of an ellipsoid can be seen, with extension along the radial direction (red arrow).}
+  \label{fig:riemann}
+\end{figure}
 
@@
 Meanwhile the internal velocity field can be seen to be that of an irrotational Riemann-S ellipsoid \cite{Daller2012}.
+
+%\vspace{1em}
+
+\subsection{Energy Conservation}\label{sec:energy}
+
 \begin{figure}
-  \includegraphics[width=1.\textwidth,trim=1cm 1cm 0 1cm,clip]{riemann}
-  \caption{Solitonic core after one revolution of a central potential, demonstrating the internal velocity field of an irrotational Riemann-S ellipsoid (where the collective motion of the soliton has been subtracted from the velocity field).
-Deformation of the spherical density profile into that of an ellipsoid can be seen, with extension along the radial direction (red arrow).}
-  \label{fig:riemann}
+  \includegraphics[width=1.1\textwidth,trim=2.5cm 0 0 1cm,clip]{combined_energy_and_density1}
+  \caption{Left: Evolution of the density profile of a solitonic core at equally spaced times as it undergoes tidal disruption in a potential centred at the red cross.
+Right: Evolution of the energy of the system, where times are indicated in correspondence with the snapshots of the density profile.
+All quantities are in dimensionless code units.}
+  \label{fig:combined_1}
 \end{figure}
 
-\vspace{1em}
+\begin{figure}
+  \includegraphics[width=0.9\textwidth,trim=0 0 2cm 2cm,clip]{egy_m=22.eps}
+  \caption{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}
+  \label{fig:binary}
+\end{figure}
 
-\subsection{Energy Conservation}\label{sec:energy}
+\begin{figure}
+  \includegraphics[width=0.9\textwidth,trim=0 0.5cm 0 0,clip]{Total_Energy_Change.eps}
+  \caption{Demonstration of improving energy conservation with simulation resolution for a binary system of solitons orbiting their common centre of mass.}
+  \label{fig:energy_change}
+\end{figure}
 
@@
 In the first case, we examine the evolution of the energy for a single soliton undergoing significant tidal disruption within a Newtonian central potential.
+For this simulation a soliton of mass 12 in code units was initialised at a radial distance of 3 code units from the centre of a Newtonian central potential generated by a central mass of 1000 code units.
 We see that as the soliton is disrupted, the kinetic energy increases, while the gravitational energy due to the central potential decreases, as expected.
@@
 Higher grid resolutions result in even better energy conservation.
+
 Figure \ref{fig:binary} demonstrates the evolution of the energy of a binary system of solitons in elliptical orbits around their common centre of mass.
 In this case the simulation ran for approximately three orbital periods at $256^3$.
-At points of closest approach, it can be clearly seen that the kinetic energy increases as the solitons speed up, while the potential energy due to self-interaction decreases commensurately.
+In dimensionless code units, the soliton mass was 22, the initial separation was 2, and the initial relative velocity was 3.6.
+At points of closest approach, it can be clearly seen that the kinetic energy increases as the solitons speed up, while the potential energy due to self-interaction decreases commensurately such that the total energy is conserved.
 In this scenario no central potential has been included, so the gravitational energy due to the central potential stays uniformly zero throughout the simulation and is not plotted in the Figure.
@@
 
-% \begin{figure}[tbp]
-% \centering 
-% \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{density_diagram1}
-% \caption{Density profile of solitonic core at equally spaced times as it undergoes tidal disruption in a potential centred at the red cross.}\label{fig:density_diagram1}
-% \end{figure}
-
-% \begin{figure}[tbp]
-% \centering 
-% \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{energy_diagram1}
-% \caption{Evolution of the energy of the system illustrated in Figure \ref{fig:density_diagram1}.
-Times are indicated in correspondence with the snapshots of the density profile.
-}\label{fig:energy_diagram_1}
-% \end{figure}
-
-% \begin{figure}[tbp]
-% \centering 
-% \includegraphics[width=1.1\textwidth,trim=2.5cm 0 0 1cm,clip]{combined_energy_and_density1}
-% \caption{Left: Evolution of the density profile of a solitonic core at equally spaced times as it undergoes tidal disruption in a potential centred at the red cross.
-Right: Evolution of the energy of the system, where times are indicated in correspondence with the snapshots of the density profile.
-All quantities are in dimensionless code units.}\label{fig:combined_1}
-% \end{figure}
-
-% \begin{figure}[tbp]
-% \centering 
-% \includegraphics[width=1.\textwidth,trim=0 0 2cm 2cm,clip]{energy_diagram2}
-% \caption{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}\label{fig:binary}
-% \end{figure}
-
-\begin{figure}
-  \includegraphics[width=1.1\textwidth,trim=2.5cm 0 0 1cm,clip]{combined_energy_and_density1}
-  \caption{Left: Evolution of the density profile of a solitonic core at equally spaced times as it undergoes tidal disruption in a potential centred at the red cross.
-Right: Evolution of the energy of the system, where times are indicated in correspondence with the snapshots of the density profile.
-All quantities are in dimensionless code units.}
-  \label{fig:combined_1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=0.9\textwidth,trim=0 0 2cm 2cm,clip]{egy_m=22.eps}
-  \caption{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}
-  \label{fig:binary}
-\end{figure}
+Figure \ref{fig:energy_change} demonstrates the relationship between the total integrated energy and the grid resolution for the same binary system of solitons used to generate Figure \ref{fig:binary}.
+While energy is conserved at sub-percent level even for low resolutions ($96^3$), this is dramatically improved at higher grid resolutions.
 

### 2018-06-01 draft.tex
Responses to RE's notes. More changes to follow.

@@
 
+\re{With regard to the `SP' vs.
+Schr{\"o}dinger-Poisson issue: I also don't like too many initialisms in publications but felt that in this case, due to the sheer number of instances of the term, and given that it seems to have been adopted as a standard by some of the other ULDM folk (e.g.
+arXiv 1801.03507, 1608.00802, 1801.04864), it made sense to compactify things.
+I've removed all instances of `SP' in this version, but feel that somehow the full term can appear rather cumbersome at times.
+What do you think?}
+
 The realisation that there may be more to the universe than meets the eye is one of the most profound developments in 20$^{\textrm{th}}$ Century astronomy and astrophysics.
@@
 Consequently, \PyUltraLight is free from noise associated with spatial derivatives computed via finite-differencing.
-There is a necessary computational cost associated with the Fourier and inverse Fourier transforms but these transforms are optimised in \PyUltraLight\ through the use of the pyFFTW pythonic wrapper around the C-based FFWTW subroutine library \cite{pyfftw,fftw}.
-\PyUltraLight is currently takes advantage of multiple cores on a singe machine but the FFTW libraries offer full parallelisation, so there is no barrier to running the code in a cluster environment.
-\re{ Will the ``top level" Python routines run in a full MPI environment; i.e.
-a cluster based shared-memory implementation?
-We would need this to make the full cluster-based approach work.}
+There is a necessary computational cost associated with the Fourier and inverse Fourier transforms but these transforms are optimised in \PyUltraLight through the use of the pyFFTW pythonic wrapper around the C-based FFWTW subroutine library \cite{pyfftw,fftw}.
+As the FFTW libraries offer full parallelisation, \PyUltraLight is currently designed to take advantage of multiple cores on a user PC or shared-memory environment.
+Full MPI compatibility has not yet been implemented as we have not found a need to run simulations in a distributed-memory cluster environment, however future releases may address this possibility.
 
@@
 We first provide a short review of ULDM physics, including a derivation of the Schr{\"o}dinger-Poisson equations from the underlying scalar-field Lagrangian.
-We then describe their implementation in \PyUltraLight, before moving on to the testing and verification Section in which we reproduce a selection of results from a variety of recent ULDM simulations and discuss energy conservation within the code.
+We then describe their implementation in \PyUltraLight, before moving on to describe testing and verification procedures applied to the code.
+We reproduce a selection of results from a variety of recent ULDM simulations and discuss energy conservation in detail.
 
@@
 Within the ULDM framework, it is proposed that cosmological structure formation proceeds through the Bose-Einstein condensation of this scalar field.
-In this Section  \re{I would nornally give Section a capital letter in this context} we derive the (non-relativistic) dynamical equations of the single macroscopic condensate wavefunction, namely, the Schr{\"o}dinger-Poisson equations.
+In this Section  \re{I would normally give Section a capital letter in this context} \re{(EK) I would be inclined to only capitalise `section' in instances where it appears as a proper noun, i.e.
+"In Section X we showed...", where we are referring to a specifically named section, numbered X.
+It seems to me that in this particular context `section' is a common noun.
+Happy to capitalise all instances of the word if this is your recommendation.} we derive the (non-relativistic) dynamical equations of the single macroscopic condensate wavefunction, namely, the Schr{\"o}dinger-Poisson equations.
 We begin with the action functional for a scalar field, $\phi$, minimally coupled to gravity and in the absence of self-interactions,
@@
 where $H=\Dot{a}(t)/a(t)$.
-We will work in the limit $H\ll m$ and it is sufficient to set $H=0$ and $a(t)=1$ in equation \ref{eq:linear}.
-\footnote{ \re{ Just checking -- we could say we were working in a  non-expanding universe and just set $H=0$ whatever $m$ was?
-Physically, what does this statement mean?
-Can we say when this approximation becomes good for $m=10^{-22}$eV dark matter?
-We go on to say non-relativistic, but is that synonymous with non-expanding?}}  In the non-relativistic limit,  the WKB approximation gives an ansatz for a  solution to equation \ref{eq:linear}.
-We take
+At late times in an expanding universe, $H\ll m$ and it is sufficient to set $H=0$ and $a(t)=1$ in equation \ref{eq:linear}.
+This is a good approximation even at relatively high redshifts, including the epochs of early structure formation.
+Alternatively, if we consider a non-expanding universe, these equalities are true by definition.
+In either case, we can remove the second term in equation \ref{eq:linear} entirely, such that all remaining derivatives are second order in space or time.
+
+In the non-relativistic limit, Equation \ref{eq:linear} can be treated using the WKB approximation\footnote{For a detailed explication of the WKB approximation in the non-relativistic limit, see \cite{Young2015}.}.
+This allows us to write an ansatz solution for the field $\phi$:
 \begin{equation}\label{eq:ansatz}
@@
 \end{equation}
-where $\psi$ is assumed to be slowly varying (with respect to intervals $\Delta t\sim \hbar/m$ \re{right?}) complex function so that terms involving $\Ddot{\psi}$ may be neglected.
-We may also use $\nabla^2\phi\ll m^2/\hbar^2\phi$, \re{physical origin of this} such that our final result takes the form
+where $\psi$ is assumed to be slowly varying in the sense that $m\vert\psi\vert\gg\hbar\vert\dot{\psi}\vert$, $m\vert\dot{\psi}\vert\gg\hbar\vert\Ddot{\psi}\vert$, $m\vert\psi\vert\gg\hbar\vert\nabla{\psi}\vert$, and $m\vert\nabla{\psi}\vert\gg\hbar\vert\nabla^2{\psi}\vert$.
+Direct substitution of the ansatz solution into equation \ref{eq:linear}, discarding the heavily suppressed second derivative terms, yields
 \begin{equation}\label{eq:schrodinger}
@@
 \end{equation}
-We have thus show that $\psi$ satisfies the Schr{\"o}dinger equation in this limit, which is interpreted as the unique macroscopic wavefunction of a Bose-Einstein condensate.
+We have thus shown that $\psi$ satisfies the Schr{\"o}dinger equation in this limit, which is interpreted as the unique macroscopic wavefunction of a Bose-Einstein condensate.
 It follows that the particle density of the condensate is given by $\vert\psi\vert^2$, so its mass density is simply $m\vert\psi\vert^2$.
@@
 It is this system of equations which \PyUltraLight is designed to solve.
-\re{I try to avoid acronyms like SP; ULDM works since it replaces something very long, and CDM is common usage, but I think papers are more readable if they are used sparingly}
 
@@
 \end{equation}
-Note that the exponential in equation \ref{eq:approx-time-ev} omits the time-ordering symbol, and is only equivalent to its time-ordered counterpart to order $h^2$. \re{to here}
+Note that the exponential in equation \ref{eq:approx-time-ev} omits the time-ordering symbol, and is only equivalent to its time-ordered counterpart to order $h^2$.
 
@@
 After an inverse Fourier transform a final half timestep is taken with the updated nonlinear term acting in position space to give the new $\psi$ field configuration.
-This procedure is known as the symmetrised split-step Fourier method, and  used widely in fields such as nonlinear fiber optics \cite{Agrawal2013}.
-\re{do we want a couple of other examples/cites?}
+This procedure is known as the symmetrised split-step Fourier method, and  used widely in fields such as nonlinear fiber optics \cite{Agrawal2013, Zhang2018, Sinkin2003}.
 
@@
 At this point only a central $1/r$ potential is supported but this would be easily generalised.
-\re{new material.
-Also, do we need to say anything about what happens when the solitions cross into the ``next" box?}
+It should be noted that because \PyUltraLight enforces periodic boundary conditions, care must be taken in cases where solitons approach the boundaries of the simulation grid.
+If a soliton were to cross the boundary during a simulation in which a Newtonian central potential is implemented, the forces exerted during the crossing would be unphysical.
+For studies of orbital stability this is unlikely to cause any problems, as in these circumstances material collapses toward the centre of the simulation grid rather than crossing the boundaries.
+However, the user should ensure that solitons are initialised sufficiently far from the boundary for the purposes of each simulation on a case-by-case basis.
+In situations where a significant portion of the total mass is expected to be ejected, such as the merger of multiple solitons to form a larger halo, care should be taken to ensure that mass ejected above the escape velocity is not recaptured as it re-enters the grid from the other side.
+For studies of this kind, an absorbing sponge at the grid boundaries is perhaps more suitable than periodic boundary conditions, though this has not been implemented in \PyUltraLight at this stage.
 
-The soliton profile is found by imposing a spherically symmetric profile \cite{Paredes2016} where the radial profile is assumed to be time independent
+The soliton profile used to generate the initial conditions in \PyUltraLight is found by first imposing spherical symmetry in the Schr{\"o}dinger-Poisson equations and assuming time independence in the radial density profile \cite{Paredes2016}:
 %
@@
 where primes denote derivatives with respect to $r$.
-Note that this system contains no arbitrary constants, so the underlying profile will effectively be universal and \PyUltraLight loads a precomputed profile supplied with the code, rather than computing it from scratch each time the code is run.
-However, an auxiliary program {\sc soliton\texttt{\_}solution.py} is supplied with  \PyUltraLight; it uses a  fourth-order Runge-Kutta algorithm to solve the coupled profile equations.
+Note that this system contains no arbitrary constants, so the underlying profile is effectively universal and is loaded as a pre-computed array by \PyUltraLight, rather than computed from scratch with each code execution.
+The soliton profile numpy array file is included with \PyUltraLight, however, an auxiliary program {\sc soliton\texttt{\_}solution.py} is also supplied, from which this array can be generated; it uses a  fourth-order Runge-Kutta algorithm to solve the coupled profile equations.
 We set $\left.
@@
 The solitons were chosen with dimensionless masses 5 and 10, with an initial separation of 4 code units and relative velocity of 20 code units.
-The simulation was run at $256^3$ for a box of side-length 8 code units.
-\re{do we want to say something about how a code unit is defined for a velocity?}}
+The simulation was run at $256^3$ for a box of side-length 8 code units.}
   \label{fig:interference}
@@
 Following \cite{Schwabe2016}, we consider the head-on collision of two solitions with mass ratio $\mu=2$ and high relative velocity.
-This simple case can be treated approximately.
+While we work in dimensionless code units, it should be noted that a dimensionful velocity can be restored from the code velocity by multiplying through by $\CMcal{L}\CMcal{T}^{-1}$, the scale parameters defined in Equations \ref{eq:length} and \ref{eq:time}.
+This simple case of a head-on soliton collision can be treated approximately.
 Starting from equation \ref{eq:covariant-solution} we write the total wavefunction of the binary system in terms of dimensionless quantities along the collision axis as
@@
     \psi(x,t)=& \ \alpha_1 f\big(\sqrt{\alpha_1}\vert x-x_1-v_1 t\vert\big)e^{i\left(\alpha_1\beta t+v_1(x-x_1)-\frac{1}{2}v_1^2 t+\delta\right)}\nonumber\\
-    &+\alpha_2 f\big(\sqrt{\alpha_2}\vert x-x_2-v_2 t\vert\big)e^{i\left(\alpha_2\beta t+v_2(x-x_2)-\frac{1}{2}v_2^2 t\right)}
+    &+\alpha_2 f\big(\sqrt{\alpha_2}\vert x-x_2-v_2 t\vert\big)e^{i\left(\alpha_2\beta t+v_2(x-x_2)-\frac{1}{2}v_2^2 t\right)},
 \end{align}
-where $x_1$ and $x_2$ are the initial centrals positions of the solitions, $v_1$ and $v_2$ are the soliton velocities, $\delta$ is a constant relative phase term and $\alpha_1=1/4 \ \alpha_2$, parameterising the density profiles as discussed in section \ref{sec:soliton-profiles}.
+where $x_1$ and $x_2$ are the initial centrals positions of the solitions, $v_1$ and $v_2$ are the soliton velocities, $\delta$ is a constant relative phase term and $\alpha_1=1/4 \ \alpha_2$, parameterising the density profiles as discussed in Section \ref{sec:soliton-profiles}.
 For convenience we set $v_1=-v_2$ and $x_1=-x_2$.
@@
     \vert\psi(x,t_{o})\vert^2= \ \alpha_1^2\bigg[&f(\sqrt{\alpha_1}x)^2+16f(2\sqrt{\alpha_1}x)^2+\nonumber\\
-    &8f(\sqrt{\alpha_1}x)f(2\sqrt{\alpha_1}x)\operatorname{cos}\left(-3\alpha_1\beta\left\vert\frac{x_1}{v_1}\right\vert+2v_1x+\delta \right)\bigg]
+    &8f(\sqrt{\alpha_1}x)f(2\sqrt{\alpha_1}x)\operatorname{cos}\left(-3\alpha_1\beta\left\vert\frac{x_1}{v_1}\right\vert+2v_1x+\delta \right)\bigg].
 \end{align}
@@
 The numerical result obtained using \PyUltraLight closely matches the theoretical prediction of equation \ref{eq:predicted-interference}.
-Small disparities between the numerical and theoretical profiles may be attributed to the effect of gravitational contraction not included in the theoretical prediction of equation \ref{eq:predicted-interference} and to a small offset in the true time of maximal interference \re{due to the solitons accelerating as they fall together?}.
-We do not expect an exact match, but we have verified \PyUltraLight qualitatively reproduce the wave interference effects of the ULDM model.
-\re{are there numerical treaments elsewhere we can compare to?}
-
+Small disparities between the numerical and theoretical profiles may be attributed to the effect of gravitational contraction not included in the theoretical prediction of equation \ref{eq:predicted-interference} and to a small offset in the true time of maximal interference due to the solitons accelerating as they fall together.
+We do not expect an exact match, but we have verified that \PyUltraLight qualitatively reproduces the wave interference effects of the ULDM model.
+With the exception of \cite{Schwabe2016}, few studies of ULDM dynamics have investigated the interference patterns generated by colliding solitons in this way.
+In some cases, this is because the algorithm employed to simulate the dynamics is not capable of reproducing such effects.
+An example of this is given in \cite{Veltmaat2016}, where it is demonstrated that the coarse-grained nature of the particle-mesh method renders the algorithm incapable of reproducing detailed interference patterns such as those shown here.
 \re{read to here}
@@
 
-When the SP system is recast as a system of hydrodynamical equations, the fluid velocity is given by the gradient of the smoothly-varying phase at any point.
-This implies that the velocity flow of the SP system is irrotational, $\nabla\times\vec{v}=0$.
+When the Schr{\"o}dinger-Poisson system is recast as a system of hydrodynamical equations, the fluid velocity is given by the gradient of the smoothly-varying phase at any point.
+This implies that the velocity flow of the Schr{\"o}dinger-Poisson system is irrotational, $\nabla\times\vec{v}=0$.
 When a soliton is initialised in a circular orbit around a Newtonian potential, the initially spherical profile is distorted, elongated along the radial direction of the central potential.
@@
 
-In addition to the evolution of the condensate wavefunction $\psi$ and the potential $\Phi$, \PyUltraLight also evaluates the energy of the SP system.
+In addition to the evolution of the condensate wavefunction $\psi$ and the potential $\Phi$, \PyUltraLight also evaluates the energy of the Schr{\"o}dinger-Poisson system.
 An important achievement of the code is sub-percent level energy conservation, even at relatively low spatial resolution, for all dynamical scenarios tested.
-In this section we derive the expression for the total energy of the SP system and discuss its decomposition into individual constituents calculated separately within the code.
+In this section we derive the expression for the total energy of the Schr{\"o}dinger-Poisson system and discuss its decomposition into individual constituents calculated separately within the code.
 We then present the numerical energy evolution of a variety of configurations.
 
-We begin by defining a suitable action which yields the full SP system through its corresponding Euler-Lagrange equations. We find that variation of
+We begin by defining a suitable action which yields the full Schr{\"o}dinger-Poisson system through its corresponding Euler-Lagrange equations. We find that variation of
 \begin{equation}\label{eq:sp-action}
@@
 Because we are working with the dimensionless quantities defined in equation \ref{eq:dimensionless}, it is easy to see that this quantity is related to the physical energy through multiplication by a constant factor of $\CMcal{L}^5\CMcal{T}^{-4}G^{-1}$.
-It should be noted that equation \ref{eq:energy-tot} is not equivalent to the expectation value of the Schr{\"o}dinger Hamiltonian, which is itself not a conserved quantity of the SP system and is given by
+It should be noted that equation \ref{eq:energy-tot} is not equivalent to the expectation value of the Schr{\"o}dinger Hamiltonian, which is itself not a conserved quantity of the Schr{\"o}dinger-Poisson system and is given by
 \begin{equation}
@@
 The two terms in the integral \ref{eq:energy-tot} are calculated separately within the code.
-It is easy to see that the first term represents the gravitational potential energy of the SP system.
+It is easy to see that the first term represents the gravitational potential energy of the Schr{\"o}dinger-Poisson system.
 As discussed in \cite{Hui2016}, the second term may be decomposed further into two contributions which may be considered separately as kinetic and `quantum' energies, but for the purposes of this implementation it is sufficient to consider only their combined contribution.

### 2018-06-03 draft.tex
richard -- read through until end

@@
 \usepackage{color}
-
+\usepackage{url}
 \usepackage[justification=justified]{caption}
@@
 \newcommand{\re}[1]{\textcolor{blue}{[{\bf RE}: #1]}}
+\newcommand{\ek}[1]{\textcolor{red}{[{\bf EK}: #1]}}
 
@@
 \PyUltraLight is implemented in a Python-based Jupyter notebook, solving the Schr{\"o}dinger-Poisson equation  governing  ultralight scalar field dark matter dynamics in the non-relativistic regime using a symmetrised split-step pseudospectral  algorithm.
-The notebook interface makes it simple to specify simulation parameters and  visualise the resulting output.
-\PyUltraLight  runs on standard desktop hardware and is available on GitHub.
-}
+The notebook interface makes it simple to specify simulation parameters and  visualise the resulting output but performance-critical routines are managed via calls to computationally efficient compiled libraries.
+\PyUltraLight  runs on standard desktop hardware with support for shared memory mutlithreading and is available on GitHub.
+\re{small change}}
 
@@
 
-\re{With regard to the `SP' vs.
+\ek{With regard to the `SP' vs.
 Schr{\"o}dinger-Poisson issue: I also don't like too many initialisms in publications but felt that in this case, due to the sheer number of instances of the term, and given that it seems to have been adopted as a standard by some of the other ULDM folk (e.g.
@@
 What do you think?}
+\re{Fair enough, but I think it reads better without -- ULDM works via analogy with CDM; it's not consistent but seems better to me. Also, have added and "EK" comment :-) }
 
@@
 Within the ULDM framework, it is proposed that cosmological structure formation proceeds through the Bose-Einstein condensation of this scalar field.
-In this Section  \re{I would normally give Section a capital letter in this context} \re{(EK) I would be inclined to only capitalise `section' in instances where it appears as a proper noun, i.e.
+In this Section  \re{I would normally give Section a capital letter in this context} \ek{I would be inclined to only capitalise `section' in instances where it appears as a proper noun, i.e.
 "In Section X we showed...", where we are referring to a specifically named section, numbered X.
 It seems to me that in this particular context `section' is a common noun.
-Happy to capitalise all instances of the word if this is your recommendation.} we derive the (non-relativistic) dynamical equations of the single macroscopic condensate wavefunction, namely, the Schr{\"o}dinger-Poisson equations.
+Happy to capitalise all instances of the word if this is your recommendation.}
+\re{I agree, but thought ``this Section'' was an example of it being used as a proper noun. But it is not a huge deal either way.}
+we derive the (non-relativistic) dynamical equations of the single macroscopic condensate wavefunction, namely, the Schr{\"o}dinger-Poisson equations.
 We begin with the action functional for a scalar field, $\phi$, minimally coupled to gravity and in the absence of self-interactions,
@@
 An example of this is given in \cite{Veltmaat2016}, where it is demonstrated that the coarse-grained nature of the particle-mesh method renders the algorithm incapable of reproducing detailed interference patterns such as those shown here.
-\re{read to here}
+%\re{read to here}
 
-\subsection{Effective Forces Due to Destructive Interference}
+\subsection{Effective Forces From Destructive Interference}
 
@@
   \includegraphics[width=1.\textwidth, trim={0 0 0 0},clip]{phase_comparison}
-  \caption{Head-on collisions between two solitons of mass 5, relative velocity 6, and initial separation 4 in code units. Time progresses from left to right across each row.}
+  \caption{Head-on collisions between two solitons of mass 5, relative velocity 6, and initial separation 4 in code units.
+Time progresses from left to right across each row.
+\re{maybe a few more words on what is going on (e.g.
+we see that in the lower panel the solitons slow as they approach one another.
+Also, should we choose masses/separations where it does not take  20+ Gyr for them to approach one another.
+As a test of the dynamics it is not a big deal, but it looks a bit funny if it is several times the age of the universe}}
   \label{fig:repulsion}
@@
 
-As demonstrated in \cite{Paredes2016}, the wavelike properties of ULDM can give rise to effective forces which can dramatically affect the dynamics of core collisions.
-These effective forces arise as a result of interference phenomena, and do not require the ULDM model to incorporate explicit local interactions.
-Figure \ref{fig:repulsion} shows the results of two simulations using \PyUltraLight in which two solitons of equal mass undergo a head-on collision.
+As demonstrated in \cite{Paredes2016}, the wavelike properties of ULDM can give rise to effective forces which  dramatically affect the dynamics of core collisions.
+These effective forces arise as a result of interference phenomena, rather than because of any local interactions the ULDM model might  incorporate.
+Figure \ref{fig:repulsion} shows the results of two \PyUltraLight where two solitons of equal mass undergo head-on collisions.
 Contours of the density profile along the plane of symmetry are displayed.
-In the first simulation, there is no phase offset between the initial solitons, while in the second simulation the phases differ by $\pi$.
-We can see that the result of this phase shift is to create an effective repulsive force between the two solitons as they approach one another.
+In one case (top) there is no phase offset between the initial solitons, while in the second  the phases differ by $\pi$.
+In the latter case, the phase shift creates an effective repulsive force between the two solitons.
 Further discussion of this phenomenon and its possible observational consequences can be found in \cite{Paredes2016}.
@@
   \caption{Solitonic core after one revolution of a central potential, demonstrating the internal velocity field of an irrotational Riemann-S ellipsoid (where the collective motion of the soliton has been subtracted from the velocity field).
-Deformation of the spherical density profile into that of an ellipsoid can be seen, with extension along the radial direction (red arrow).}
+Deformation of the spherical density profile into that of an ellipsoid can be seen, with extension along the radial direction (red arrow).
+The host:satellite mass ratio is approximately 55; simulation resolution is $256^3$.}
   \label{fig:riemann}
@@
 
-\PyUltraLight is designed such that a central potential may be included by way of a point-mass located at the centre of the simulation grid.
-While this setup is somewhat unphysical, it provides a starting point for a study of the stability of satellite dark matter halos orbiting a much larger object.
-Studies of this type can be used to make predictions regarding the lifespans of satellite galaxies around the Milky Way, offering a potential resolution of the so-called missing satellites problem \cite{Weinberg2015}.
+\PyUltraLight allows the inclusion of a static potential as point-mass  at the centre of the simulation region.
+There is no backreaction on this mass as a result of the ULDM dynamics, and its ``mirror images'' within the periodic coordinate systems are not accounted for within the overall gravitational potential.
+The setup is thus somewhat unphysical but it provides a starting point for a study of the stability of satellite dark matter halos orbiting a much larger object.
+In particular, this includes the investigation of lifespans of dwarf satellite galaxies orbiting much larger objects (including the Milky Way) which are key to understanding whether ULDM models can resolve for the so-called missing satellites problem \cite{Weinberg2015}.
 
-An extensive study of the tidal disruption of ULDM solitonic cores orbiting a central potential has recently been undertaken in \cite{Du2018}.
-Here, we reproduce just one of their results as an additional validity check for \PyUltraLight.
-In particular, we reproduce the characteristic internal velocity field of a tidally-locked soliton orbiting a central potential.
+An extensive study of the tidal disruption of ULDM solitonic cores orbiting a central potential has recently been undertaken in \cite{Du2018} and we reproduce just one of their results here.
+When the Schr{\"o}dinger-Poisson system is recast as a system of hydrodynamical equations, the fluid velocity  at any point is the gradient of the smoothly-varying phase.
+Since the curl of a gradient vanishes identically, the velocity field of the Schr{\"o}dinger-Poisson system must be irrotational, and $\nabla\times\vec{v}=0$.
+If a radially symmetric soliton is initialised in a circular orbit around a Newtonian potential, there will be an initial transient as the spherical profile becomes elongated along the radial direction of the central potential.
+Meanwhile, the velocity field corresponding to the overall orbital motion of the soliton is superposed with the internal velocity field, combining so as to produce a net flow with vanishing curl.
 
-When the Schr{\"o}dinger-Poisson system is recast as a system of hydrodynamical equations, the fluid velocity is given by the gradient of the smoothly-varying phase at any point.
-This implies that the velocity flow of the Schr{\"o}dinger-Poisson system is irrotational, $\nabla\times\vec{v}=0$.
-When a soliton is initialised in a circular orbit around a Newtonian potential, the initially spherical profile is distorted, elongated along the radial direction of the central potential.
-Meanwhile, the velocity field corresponding to the overall orbital motion of the soliton is superposed with the internal velocity field, combining so as to produce a net flow with vanishing curl.
 The family of Riemann-S ellipsoids describe uniformly rotating bodies which satisfy the condition of irrotationality \cite{Chandrasekhar1965}, and it is therefore the characteristic internal velocity field of such an ellipsoid which we expect to arise during the simulation.
 It is found in \cite{Du2018} that an initially spherical solitonic core without self-rotation will gradually spin up to form a tidally-locked irrotational Riemann-S ellipsoid when orbiting a host mass.
-We are able to reproduce this result using \PyUltraLight, with Figure \ref{fig:riemann} showing the internal velocity field of a solitonic satellite after one complete revolution around a host mass, obtained using \PyUltraLight on a $256^3$ simulation grid.
-In this case the ratio of host to satellite mass was $\approx 55$.
-It can be seen that the soliton has become elongated along the radial line connecting to the host, indicating that it is tidally locked.
-Meanwhile the internal velocity field can be seen to be that of an irrotational Riemann-S ellipsoid \cite{Daller2012}.
+We a reproduce this result using \PyUltraLight, with Figure \ref{fig:riemann} showing the internal velocity field of a solitonic satellite after one complete revolution around a host mass.
+The soliton has indeed become elongated along the radial line connecting it to the host, indicating that it is tidally locked, while the internal velocity field can be seen to be that of an irrotational Riemann-S ellipsoid \cite{Daller2012}.\re{Did you actually check numerically it is a Riemann-S ellipsoid or is that a qualitative assessment?
+No worries; just wanting to be clear -- as I said previous, I am not sure most people know intuitively what a Riemann-S ellipsoid is.
+Or maybe I am just poorly educated.}
 
@@
 
-In addition to the evolution of the condensate wavefunction $\psi$ and the potential $\Phi$, \PyUltraLight also evaluates the energy of the Schr{\"o}dinger-Poisson system.
-An important achievement of the code is sub-percent level energy conservation, even at relatively low spatial resolution, for all dynamical scenarios tested.
-In this section we derive the expression for the total energy of the Schr{\"o}dinger-Poisson system and discuss its decomposition into individual constituents calculated separately within the code.
-We then present the numerical energy evolution of a variety of configurations.
+Physically, we expect that the overall energy in the system will be conserved.
+This provides a test on the numerical performance of \PyUltraLight, and we find that even at relatively low spatial resolution we see sub-percent level energy conservation for all the dynamical scenarios considered here.
+In this Section we express the energy  the Schr{\"o}dinger-Poisson system in terms of the variables  $\psi$ and $\Phi$ and discuss its decomposition into individual constituents calculated separately within the code.
+We then present results for a variety of configurations.
 
@@
 The two terms in the integral \ref{eq:energy-tot} are calculated separately within the code.
-It is easy to see that the first term represents the gravitational potential energy of the Schr{\"o}dinger-Poisson system.
-As discussed in \cite{Hui2016}, the second term may be decomposed further into two contributions which may be considered separately as kinetic and `quantum' energies, but for the purposes of this implementation it is sufficient to consider only their combined contribution.
-Because \PyUltraLight provides the option to include the central potential of a point mass located at the centre of the simulation grid, we have additional energy contributions when such a potential is included.
-Because this central potential does not arise from a ULDM object, we calculate the gravitational potential energy due to self-interaction separately from the gravitational potential energy due to the central potential.
+The first term is the gravitational potential energy of the Schr{\"o}dinger-Poisson system and as discussed in \cite{Hui2016}, the second term may be decomposed into  contributions which may be considered separately as kinetic and `quantum' energies, but for our purposes it is sufficient to consider only their combined contribution.
+When \PyUltraLight includes the central potential of a point mass located at the centre of the simulation grid we have additional energy contributions and calculate the  gravitational potential energy from self-interactions separately from the gravitational potential energy due to the central potential.
 
-Figures \ref{fig:combined_1} and \ref{fig:binary} demonstrate energy conservation for two example simulation scenarios.
-In the first case, we examine the evolution of the energy for a single soliton undergoing significant tidal disruption within a Newtonian central potential.
+Figures \ref{fig:combined_1} and \ref{fig:binary} demonstrate energy conservation for two scenarios.
+The first case  shows the evolution of the energy of a single soliton undergoing significant tidal disruption within a Newtonian central potential.
 For this simulation a soliton of mass 12 in code units was initialised at a radial distance of 3 code units from the centre of a Newtonian central potential generated by a central mass of 1000 code units.
-We see that as the soliton is disrupted, the kinetic energy increases, while the gravitational energy due to the central potential decreases, as expected.
-Meanwhile, the gravitational potential energy due to self-interaction gradually increases toward zero as the disruption continues.
-The sum of the individual energy components is conserved to $10^{-2}\ \%$ in this case, where the simulation was run at $256^3$.
+As the soliton is disrupted, the kinetic energy increases, while the gravitational energy due to the central potential decreases, as expected.
+Meanwhile, the gravitational potential energy from self-interactions gradually increases toward zero as the disruption continues and the ULDM is halo spread over a greater area.
+In this case the sum of the individual energy components is conserved to $10^{-2}\ \%$  at a resolution of $256^3$.
 Higher grid resolutions result in even better energy conservation.
 
-Figure \ref{fig:binary} demonstrates the evolution of the energy of a binary system of solitons in elliptical orbits around their common centre of mass.
-In this case the simulation ran for approximately three orbital periods at $256^3$.
-In dimensionless code units, the soliton mass was 22, the initial separation was 2, and the initial relative velocity was 3.6.
-At points of closest approach, it can be clearly seen that the kinetic energy increases as the solitons speed up, while the potential energy due to self-interaction decreases commensurately such that the total energy is conserved.
-In this scenario no central potential has been included, so the gravitational energy due to the central potential stays uniformly zero throughout the simulation and is not plotted in the Figure.
-It can also be seen that as the solitons reach the first point of closest approach, they become slightly deformed, exciting oscillatory modes which are manifest in the Figure as small scale oscillations superposed on the global behaviour.
-In this case the total energy is conserved to $10^{-3}\ \%$.
-
-Figure \ref{fig:energy_change} demonstrates the relationship between the total integrated energy and the grid resolution for the same binary system of solitons used to generate Figure \ref{fig:binary}.
-While energy is conserved at sub-percent level even for low resolutions ($96^3$), this is dramatically improved at higher grid resolutions.
+Figure \ref{fig:binary} demonstrates the evolution of the energy of a binary system of solitons in elliptical orbits around their common centre of mass over three orbital periods.
+In dimensionless code units, the soliton masses are 22, the initial separation  2, and the initial relative velocity was 3.6.
+At points of closest approach the kinetic energy increases as the solitons speed up, while the potential energy due to self-interaction decreases commensurately such that the total energy is conserved.
+In this scenario no central potential has been included.
+As the solitons reach the first point of closest approach, they become slightly deformed, exciting oscillatory modes which are manifest in the Figure as small scale oscillations superposed on the global behaviour.
+Figure \ref{fig:energy_change} demonstrates the relationship between the total integrated energy and the grid resolution for the same binary system of solitons used to generate Figure~\ref{fig:binary}.
+Energy is conserved at sub-percent level even at low resolutions ($96^3$), and increasing grid resolution greatly improves accuracy.
 
@@
 
-We have demonstrated that \PyUltraLight is an accurate tool for the study of the dynamics of ultralight dark matter governed by the Schr{\"o}dinger-Poisson system of equations.
+\PyUltraLight is an accurate, flexible and easy to use tool for  studying the dynamics of ultralight dark matter governed by the Schr{\"o}dinger-Poisson system of equations.
 The code makes use of a pseudospectral symmetrised split-step Fourier methodology, in which all spatial derivatives are treated via explicit multiplication in the Fourier domain, thereby avoiding difficulties associated with finite-differencing methods.
 
-Energy conservation within \PyUltraLight is excellent, at sub-percent level for simulations run at $128^3$, and even better performance as resolution is increased.
-The code is able to reproduce complex phenomena resulting from the wave-like properties of ultralight dark matter as predicted by theoretical models, such as the interference patterns arising during high-velocity collisions of solitonic cores, and the effective forces observed in cases where the colliding cores are out of phase.
-These phenomena can be clearly observed at relatively low spatial resolution, obviating the need for high-performance computing infrastructure to study the fundamental behaviour of ULDM systems.
-This makes \PyUltraLight a useful tool for predicting the overall dynamics of a ULDM system in a computationally efficient manner prior to a rigorous examination using more advanced codes such as Nyx.
+Energy conservation within \PyUltraLight is excellent, at sub-percent level for simulations run at $128^3$, with even better performance as resolution is increased.
+The code captures complex phenomena resulting from the wave-like properties of ultralight dark matter, including as the interference patterns arising during high-velocity collisions of solitonic cores, and  effective forces observed in cases where the colliding cores are out of phase.
+These phenomena can be clearly observed at relatively low spatial resolution, avoiding the need for high-performance computing infrastructure to study the fundamental behaviour of ULDM systems in simple configurations.
+This makes \PyUltraLight a useful tool for investigating the  dynamics of ULDM systems.
 
@@
 The accompanying Jupyter notebook allows for the efficient adjustment of simulation parameters, and offers a useful browser interface for quick visualisation of simulation results.
-Despite being Python-based, the code makes use of low-level language resources, namely the FFTW libraries through the use of the Pythonic pyFFTW wrapper.
+While Python-based, the code makes use of low-level language resources, namely the FFTW libraries through the use of the Pythonic pyFFTW wrapper and will operate at $\sim 80\%$ efficiency on a 16 core desktop workstation, suggesting that it is computationally efficient.
 
-While the current implementation of \PyUltraLight is already a useful tool for simulating dynamical ULDM systems, there remains scope for substantial improvement.
-In particular, it is anticipated that future releases may incorporate adaptive mesh refinement capabilities and variable timestep functionality.
-Additionally, while the current implementation utilises a spatial grid of fixed size and does not provide a means by which to incorporate baryonic density contributions, future releases are likely to allow for cosmic expansion and a means by which to incorporate simplified models of baryonic matter.
-Further adaptations of the code may also be implemented to allow for greater flexibility in the choice of ULDM model, incorporating possibilities such as explicit self-interaction.
+The current implementation of \PyUltraLight is already a useful tool for simulating dynamical ULDM systems and exploring their dynamics However, there is much scope for  improvement.
+In particular, future releases may incorporate a variable timestep and more sophisticated physics, including explicit self-interactions in the axion sector or additional matter components.
+\PyUltraLight is publicly available under a BSD license.
+\re{took some stuff out; no  need to tip our hand completely.}
+
+\acknowledgments
+
+We thank Nathan Musoke, Jens Niemeyer (and group -- names?), for valuable discussions, and thank Miro Erkintalo for advice on Schr{\"o}dinger-Poisson solvers  in optical systems.
+We acknowledge support from the Marsden Fund of the Royal Society of New Zealand, the use of the New Zealand eScience Infrastructure (NeSI) high-performance computing facilities, which are funded jointly by NeSI's collaborator institutions and through the Ministry of Business, Innovation \& Employment's Research Infrastructure programme \url{https://www.nesi.org.nz}, (PLUS SPECIFIC NAMES OF PEOPLE AT NESI WHO HELPED??).
+(SHAUN -- PEOPLE WHO PUT YOU UP ON TRAVEL??) (FABER -- any names of yours?)
 
@@
 
-\PyUltraLight has been made publicly available under a BSD licence.
-The full code repository, including supplementary files such as the code used to generate soliton profiles, is available on GitHub at https://github.com/erckendall/PyUltraLight.
+\PyUltraLight is publicly available under a BSD licence.
+The full repository, including supplementary files such as the code used to generate soliton profiles, is available on GitHub at {\url{ https://github.com/erckendall/PyUltraLight}}.
 \PyUltraLight makes use of the pyFFTW pythonic wrapper around the FFTW C-based fast Fourier transform libraries.
-Both pyFFTW and FFTW are freely-available.
-\PyUltraLight has been successfully tested on both Mac OS and Linux operating systems, though not exhaustively.
-The current version of \PyUltraLight, as demonstrated in this work, is capable of reproducing many of the results of previous ULDM simulations, and we plan to extend its capabilities in future releases.
+Both pyFFTW and FFTW are freely-available and \PyUltraLight has been used successfully on both Mac OS and Linux  systems.
+We welcome advice and feedback from users.
 
-\acknowledgments
-
-TBW
+%\url{http://www.nesi.org.nz}
 
@@
 % in short all the informations that clearly identify a document.
+
+\re{Check the bib options so we don't see all the Planck authors :-) }
 

### 2018-06-07 draft.tex
Responses to Richard's comments plus a few small changes

@@
 
-\ek{With regard to the `SP' vs.
-Schr{\"o}dinger-Poisson issue: I also don't like too many initialisms in publications but felt that in this case, due to the sheer number of instances of the term, and given that it seems to have been adopted as a standard by some of the other ULDM folk (e.g.
-arXiv 1801.03507, 1608.00802, 1801.04864), it made sense to compactify things.
-I've removed all instances of `SP' in this version, but feel that somehow the full term can appear rather cumbersome at times.
-What do you think?}
-\re{Fair enough, but I think it reads better without -- ULDM works via analogy with CDM; it's not consistent but seems better to me. Also, have added and "EK" comment :-) }
-
 The realisation that there may be more to the universe than meets the eye is one of the most profound developments in 20$^{\textrm{th}}$ Century astronomy and astrophysics.
@@
 Happy to capitalise all instances of the word if this is your recommendation.}
-\re{I agree, but thought ``this Section'' was an example of it being used as a proper noun. But it is not a huge deal either way.}
+\re{I agree, but thought ``this Section'' was an example of it being used as a proper noun.
+But it is not a huge deal either way.}\ek{The way I would tend to think of it is by analogy with the following: "This person studies physics." vs.
+"Emily studies physics." In the former sentence, "person" is used as a description rather than a title, so would not be capitalised.
+Similar to "Richard is a professor." vs.
+"This is Professor Richard."}
 we derive the (non-relativistic) dynamical equations of the single macroscopic condensate wavefunction, namely, the Schr{\"o}dinger-Poisson equations.
@@
 \begin{equation}\label{eq:pFRW}
-    ds^2=-\big(1+2\Phi(\vec{r})\big)dt^2+a(t)^2\big(1-2\Phi(\vec{r})\big)d\vec{r}^{\,2}.
+    ds^2=-\big(1+2\Phi(\vec{r},t)\big)dt^2+a(t)^2\big(1-2\Phi(\vec{r},t)\big)d\vec{r}^{\,2}.
 \end{equation}
@@
 \begin{equation}\label{eq:linear}
-    \Ddot{\phi}-\frac{\big(1+4\Phi(\vec{r})\big)}{a(t)^2}\nabla^2\phi+3\big(1+2\Phi(\vec{r})\big)H\Dot{\phi}+\big(1+2\Phi(\vec{r})\big)\frac{m^2}{\hbar^2}\phi=0,
+    \Ddot{\phi}-\frac{\big(1+4\Phi\big)}{a(t)^2}\nabla^2\phi+3H\Dot{\phi}-4\dot{\Phi}\dot{\phi}+\big(1+2\Phi\big)\frac{m^2}{\hbar^2}\phi=0,
 \end{equation}
@@
 where $\psi$ is assumed to be slowly varying in the sense that $m\vert\psi\vert\gg\hbar\vert\dot{\psi}\vert$, $m\vert\dot{\psi}\vert\gg\hbar\vert\Ddot{\psi}\vert$, $m\vert\psi\vert\gg\hbar\vert\nabla{\psi}\vert$, and $m\vert\nabla{\psi}\vert\gg\hbar\vert\nabla^2{\psi}\vert$.
-Direct substitution of the ansatz solution into equation \ref{eq:linear}, discarding the heavily suppressed second derivative terms, yields
+Since $\Phi$ is sourced by $\psi$, we also have that $m\vert\Phi\vert\gg\hbar\vert\dot{\Phi}\vert$.
+Direct substitution of the ansatz solution into equation \ref{eq:linear}, discarding heavily suppressed terms, yields
 \begin{equation}\label{eq:schrodinger}
@@
   \includegraphics[width=1.\textwidth, trim={0 0 0 0},clip]{phase_comparison}
-  \caption{Head-on collisions between two solitons of mass 5, relative velocity 6, and initial separation 4 in code units.
-Time progresses from left to right across each row.
-\re{maybe a few more words on what is going on (e.g.
-we see that in the lower panel the solitons slow as they approach one another.
-Also, should we choose masses/separations where it does not take  20+ Gyr for them to approach one another.
-As a test of the dynamics it is not a big deal, but it looks a bit funny if it is several times the age of the universe}}
+  \caption{Head-on collisions between solitons of mass 20, initial relative velocity 20, and initial separation 1.2 in code units.
+Time progresses from left to right across each row and is indicated in code units in each frame.
+The effective repulsive force generated by a $\pi$ phase shift can be seen in the lower panel.}
   \label{fig:repulsion}
@@
 
-As demonstrated in \cite{Paredes2016}, the wavelike properties of ULDM can give rise to effective forces which  dramatically affect the dynamics of core collisions.
+As demonstrated in \cite{Paredes2016}, the wavelike properties of ULDM give rise to effective forces which can dramatically affect the dynamics of core collisions.
 These effective forces arise as a result of interference phenomena, rather than because of any local interactions the ULDM model might  incorporate.
-Figure \ref{fig:repulsion} shows the results of two \PyUltraLight where two solitons of equal mass undergo head-on collisions.
-Contours of the density profile along the plane of symmetry are displayed.
+Figure \ref{fig:repulsion} shows the results of a head-on collision between two solitons, where in one instance the solitons have no initial phase difference, and in the other instance a phase difference of $\pi$ is applied in the initial conditions.
+In this simulation, solitons of mass 20 are initialised with relative velocity 20 and initial separation 1.2 (code units).
+The solitons are allowed to collide, and contours of the density profile along the plane of symmetry are displayed.
 In one case (top) there is no phase offset between the initial solitons, while in the second the phases differ by $\pi$.
 In the latter case, the phase shift creates an effective repulsive force between the two solitons.
+It can be seen in the second frame that as the solitons approach one another, the $\pi$ phase shift results in a slowing of the approach accompanied by a deformation of the density profile, acting so as to avoid contact between the solitons.
+Dissimilarly, in the case where there is no phase shift, the solitons readily collide and merge to form a single contiguous density profile prior to re-separating.
 Further discussion of this phenomenon and its possible observational consequences can be found in \cite{Paredes2016}.
@@
 Since the curl of a gradient vanishes identically, the velocity field of the Schr{\"o}dinger-Poisson system must be irrotational, and $\nabla\times\vec{v}=0$.
-If a radially symmetric soliton is initialised in a circular orbit around a Newtonian potential, there will be an initial transient as the spherical profile becomes elongated along the radial direction of the central potential.
+If a radially symmetric soliton is initialised in a circular orbit around a Newtonian potential, there will be initial transient behaviour as the spherical profile becomes elongated along the radial direction of the central potential.
 Meanwhile, the velocity field corresponding to the overall orbital motion of the soliton is superposed with the internal velocity field, combining so as to produce a net flow with vanishing curl.
@@
 We a reproduce this result using \PyUltraLight, with Figure \ref{fig:riemann} showing the internal velocity field of a solitonic satellite after one complete revolution around a host mass.
-The soliton has indeed become elongated along the radial line connecting it to the host, indicating that it is tidally locked, while the internal velocity field can be seen to be that of an irrotational Riemann-S ellipsoid \cite{Daller2012}.\re{Did you actually check numerically it is a Riemann-S ellipsoid or is that a qualitative assessment?
-No worries; just wanting to be clear -- as I said previous, I am not sure most people know intuitively what a Riemann-S ellipsoid is.
-Or maybe I am just poorly educated.}
+The soliton has become elongated along the radial line connecting it to the host, indicating that it is tidally locked, while the internal velocity field qualitatively matches that of an irrotational Riemann-S ellipsoid within the tidal radius \cite{Daller2012}.
+It should be noted that the wider velocity field is not expected to be accurately predicted in a simulation of this kind.
+This is because the initial soliton density profile is defined only out to a given cutoff radius, beyond which the $\psi$ field value, and by extension its velocity field, is set identically to zero.
+Therefore, the velocity field far from the solitonic centre is not well-modelled in these simulations, and it may be expected that non-trivial self-interaction in the wider field would lead to unpredictable behaviour in these regions.
 
@@
   \label{fig:energy_change}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=0.9\textwidth,trim=0 0.5cm 0 0,clip]{Total_Energy_Change_Disruption.eps}
+  \caption{Demonstration of improving energy conservation with simulation resolution for a soliton undergoing significant tidal disruption in a Newtonian central potential.}
+  \label{fig:energy_change_2}
 \end{figure}
@@
 Meanwhile, the gravitational potential energy from self-interactions gradually increases toward zero as the disruption continues and the ULDM is halo spread over a greater area.
-In this case the sum of the individual energy components is conserved to $10^{-2}\ \%$  at a resolution of $256^3$.
-Higher grid resolutions result in even better energy conservation.
+In this case the sum of the individual energy components is conserved to $10^{-4}\ \%$  at a resolution of $256^3$.
 
@@
 Energy is conserved at sub-percent level even at low resolutions ($96^3$), and increasing grid resolution greatly improves accuracy.
+
+Figure \ref{fig:energy_change_2} demonstrates the improvement in energy conservation with increasing grid resolution for the case of a single soliton becoming tidally disrupted in a Newtonian central potential.
+The simulation setup is the same as that used to generate Figure \ref{fig:combined_1}.
+Namely, a single soliton of mass 12 code units is initialised at a distance of 3 code units from a central mass, $M=1000$.
+The initial velocity of the soliton is $\sqrt{M/r}$ where $r$ is the radial distance of the soliton from the central mass.
+The duration of the simulation is 0.5 code units so that the soliton undergoes significant tidal disruption as demonstrated in Figure \ref{fig:combined_1}.
+While we see that energy is conserved at sub-percent level even for $64^3$ grid resolution, the qualitative behaviour of the mass density distribution in this case is not correct, so we conclude that this resolution is insufficient for convergence despite good energy conservation.
+This highlights the importance of a multifaceted approach to convergence testing.
+At $256^3$, energy is conserved to $10^{-4}\ \%$.
 
@@
 \PyUltraLight is publicly available under a BSD license.
-\re{took some stuff out; no  need to tip our hand completely.}
 
@@
 
-\re{Check the bib options so we don't see all the Planck authors :-) }
+\re{Check the bib options so we don't see all the Planck authors :-) }\ek{I've modified the bibliography style file so that only the first 3 authors will be  displayed if there are more than 3.
+I can't seem to find official JCAP guidance on this - Does it look OK to you in this format?}
 

### 2018-06-12 draft.tex
Edits and comments made by Shaun. Many edits are just suggestions so I haven't actually made the edit, just suggested in a comment.

@@
 \newcommand{\ek}[1]{\textcolor{red}{[{\bf EK}: #1]}}
+\newcommand{\sh}[1]{\textcolor{green}{[{\bf SH}: #1]}}
 
-\abstract{\PyUltraLight  simulates the dynamics of ultralight dark matter in a static, non-expanding background.
+\abstract{\PyUltraLight  simulates the dynamics of ultralight dark matter in a static, non-expanding \sh{are static and non-expanding not exactly rhe same?} background.
 \PyUltraLight can describe the evolution of several interacting ultralight dark matter halos  or one or more halos orbiting a central, fixed Newtonian potential, the latter scenario corresponding to dwarf galaxies orbiting a massive central galaxy.
@@
 In particular, ultralight dark matter (ULDM), also known as fuzzy dark matter (FDM), or BEC dark matter, is an increasingly well-studied possibility; for a recent review of the potential advantages and characteristic attributes of this scenario see Ref~\cite{Hui2016}.
-ULDM models are well motivated by fundamental theories possessing approximate shift symmetries.
+ULDM models are well motivated by fundamental theories possessing approximate shift symmetries\sh{suggest include ", e.g.
+an axion" with references.
+Otherwise this sentence seems a bit out of place, hinting at something without being explicit.}.
 Moreover, ULDM can naturally resolve the small-scale problems of $\Lambda$CDM as the Heisenberg uncertainty principle suppresses gravitational collapse on length scales shorter than the de Broglie wavelength of the ULDM particle.
@@
 Consequently, \PyUltraLight is free from noise associated with spatial derivatives computed via finite-differencing.
-There is a necessary computational cost associated with the Fourier and inverse Fourier transforms but these transforms are optimised in \PyUltraLight through the use of the pyFFTW pythonic wrapper around the C-based FFWTW subroutine library \cite{pyfftw,fftw}.
+There is a necessary computational cost associated with the Fourier and inverse Fourier transforms but these transforms are optimised in \PyUltraLight through the use of the pyFFTW pythonic wrapper around the C-based FFTW subroutine library \cite{pyfftw,fftw}.
 As the FFTW libraries offer full parallelisation, \PyUltraLight is currently designed to take advantage of multiple cores on a user PC or shared-memory environment.
@@
 The existence of an extremely light scalar field, minimally coupled to gravity, is the central premise on which ULDM models are predicated.
-Within the ULDM framework, it is proposed that cosmological structure formation proceeds through the Bose-Einstein condensation of this scalar field.
+Within the ULDM framework, it is proposed that cosmological structure formation proceeds through the Bose-Einstein condensation of this scalar field \sh{I don't understand this sentence.
+There is a Bose-Einstein condensate and it clusters to form structures, but the sentence seems to suggest that the clustering and condensating are the same thing?}.
 In this Section  \re{I would normally give Section a capital letter in this context} \ek{I would be inclined to only capitalise `section' in instances where it appears as a proper noun, i.e.
@@
 where $H=\Dot{a}(t)/a(t)$.
-At late times in an expanding universe, $H\ll m$ and it is sufficient to set $H=0$ and $a(t)=1$ in equation \ref{eq:linear}.
-This is a good approximation even at relatively high redshifts, including the epochs of early structure formation.
+At late times in an expanding universe, $H\ll m$ \sh{I might be confused, but I'm reasonably sure the approximation is $H\ll m/\hbar$} and it is sufficient to set $H=0$ and $a(t)=1$ in equation \ref{eq:linear}.
+This is a good approximation even at relatively high redshifts, including the epochs of early structure formation\sh{maybe worth mention that this approximation is equivalent to saying that the Hubble horizon is much larger than the compton wavelength?
+- I'm not sure this condition is enough to neglect H though because in Nathan's code it is still true, but H is being included.
+It is somehow the de Broglie wavelength that is relevant, not just Compton, so a comparison also needs made to the $\nabla^2 \phi$ term}.
 Alternatively, if we consider a non-expanding universe, these equalities are true by definition.
-In either case, we can remove the second term in equation \ref{eq:linear} entirely, such that all remaining derivatives are second order in space or time.
+In either case, we can remove the second \sh{third?}term in equation \ref{eq:linear} entirely, such that all remaining derivatives are second order in space or time \sh{this sentence confuses me because it seems to imply we've removed terms with higher order derivatives, which isn't true}.
 
@@
 \end{equation}
-We have thus shown that $\psi$ satisfies the Schr{\"o}dinger equation in this limit, which is interpreted as the unique macroscopic wavefunction of a Bose-Einstein condensate.
-It follows that the particle density of the condensate is given by $\vert\psi\vert^2$, so its mass density is simply $m\vert\psi\vert^2$.
+We have thus shown that $\psi$ satisfies the Schr{\"o}dinger equation in this limit, which is interpreted as the unique macroscopic wavefunction of a Bose-Einstein condensate \sh{I'm also not sure I understand this sentence.
+There are other equations that describe other Bose-Einstein condensates (e.g.
+other N-L Schrod equations with different potentials.) I suggest just removing the word unique.}.
+It follows that the particle \sh{suggest include 'number' after particle} density of the condensate is given by $\vert\psi\vert^2$, so its mass density is simply $m\vert\psi\vert^2$.
 The local gravitational potential thus satisfies the Poisson equation,
@@
 It is helpful to recast the Schr{\"o}dinger-Poisson system (equations \ref{eq:schrodinger} and \ref{eq:poisson}) in terms of adimensional quantities.
-In keeping with Refs~\cite{Schive2014,Paredes2016} we introduce length, time, and mass scales as follows:
+In keeping with Refs~\cite{Schive2014,Paredes2016} we introduce length, time, and mass scales as follows\sh{I think I fixed a typo in equation 2.8, but please check.
+I added the 1/2 in the exponent on the RHS}:
 \begin{align}
-    \CMcal{L}&=\left(\frac{8\pi\hbar^2}{3 m^2H_0^2\Omega_{m_0}}\right)^{\frac{1}{4}}\approx121\left(\frac{10^{-23}\operatorname{eV}}{m}\right)\operatorname{kpc},\label{eq:length}\\
+    \CMcal{L}&=\left(\frac{8\pi\hbar^2}{3 m^2H_0^2\Omega_{m_0}}\right)^{\frac{1}{4}}\approx121\left(\frac{10^{-23}\operatorname{eV}}{m}\right)^{1/2}\operatorname{kpc},\label{eq:length}\\
     \CMcal{T}&=\left(\frac{8\pi}{3 H_0^2\Omega_{m_0}}\right)^{\frac{1}{2}}\approx75.5 \operatorname{Gyr},\label{eq:time}\\
@@
 \end{equation}
-where $\mathcal{T}$ is the time-ordering symbol. For a sufficiently small timestep $h$,  the trapezoidal rule gives the approximation
+where $\mathcal{T}$ is the time-ordering symbol \sh{I think I need this explained to me.
+I might just have a hole in my maths knowledge regarding what the $\mathcal{T}$ means in this context.
+I'm happy to go read up and fill that hole though so a reference might be enough.}.
+For a sufficiently small timestep $h$,  the trapezoidal rule gives the approximation
 \begin{equation}
@@
 The linear differential operator in equation \ref{eq:approx-time-ev} acts naturally in Fourier space, while the nonlinear potential term is simplest to evaluate in position space and by splitting the exponential we can evaluate each term in its natural domain.
-For small timesteps we assume that the dispersive and nonlinear terms act independently,  adopting the specific evolution scheme
+For small timesteps we assume that the dispersive and nonlinear terms act independently \sh{I don't understand what 'independently' means in this context.
+Can 3.4 not just be an ansatz that is proven as valid by 3.5?},  adopting the specific evolution scheme
 %
@@
 The final operation in equation \ref{eq:schematic-psi}  only changes the phase of  $\psi$, so we could replace $\psi(\vec{x},t_{i})$ with $\psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)$ in equation \ref{eq:schematic-phi} with no change in meaning.
-\PyUltraLight makes an additional simplification ti the symmetrised split-step Fourier method as the consecutive half-steps in the nonlinear term may be combined into a single full step.
+\PyUltraLight makes an additional simplification to the symmetrised split-step Fourier method by combining the consecutive half-steps in the nonlinear term into a single full step \sh{I changed this sentence a bit to make it more readable, but am happy for it to be reverted if other's prefer the original form}.
 Consequently, only the first and last operations involve actual half steps.
@@
 
-From a computational perspective, the numerical Fourier transforms are likely to be rate-limiting step in any pseudospectral code.
+From a computational perspective, the numerical Fourier transforms are likely to be the rate-limiting step in any pseudospectral code.
 In \PyUltraLight the discrete Fourier transform (DFT) and its inverse are implemented via pyFFTW, a pythonic wrapper for the C-based FFTW subroutine library which efficiently implements both real and complex DFTs \cite{pyfftw,fftw,Frigo2005}.
@@
 
-In many finite-differencing schemes for  partial differential equations the Courant-Friedrichs-Lewy (CFL) condition is a necessary condition for convergence \cite{Ajaib2013}, and puts an upper bound on the timestep as  function of the gridsize whose physical origin is the finite speed of propagating information  in the evolving system.
+In many finite-differencing schemes for  partial differential equations the Courant-Friedrichs-Lewy (CFL) condition is a necessary condition for convergence \cite{Ajaib2013}, and puts an upper bound on the timestep as a function of the gridsize.
+The physical origin of the CFL condition is the finite speed of propagating information  in the evolving system \sh{split one sentence into two}.
 Consequently, this condition is often cited in numerical analysis of ULDM physics governed by the Schr{\"o}dinger-Poisson system, see e.g.
@@
 We define a `maximum velocity' for the simulation grid at which point the fluid appears to move backward because the phase change between successive grid points exceeds $\pi$.
-An upper bound can then be placed on the timestep so that a fluid travelling at this maximum  velocity  traverse a distance of one grid space per timestep.
+An upper bound can then be placed on the timestep so that fluid travelling at this maximum  velocity traverse a distance of one grid space per timestep.
 In practice this can yield a very demanding constraint on the timestep and for many scenarios the actual velocity is far less than this hypothetical limit.
-Consequently, the timestep can be  significantly increased at the user's discretion but convergence testing on a case-by-case basis is strongly encouraged.
+Consequently, the timestep can be  significantly increased at the user's discretion but convergence testing on a case-by-case basis is strongly encouraged.\sh{I wonder whether it is worth mentioning that the two pieces of the split step method are exact to arbitrary large times when the other piece isn't present.
+Also, I wonder whether it is worth including a "boost factor" to the time step in the jupyer-notebook to aid people in making that change?}
 
@@
 
-In this section we validate \PyUltraLight a by reproducing results from previous studies of ULDM dynamics, demonstrating interference effects and effective repulsive forces arising from the wavelike nature of ULDM.
+In this section we validate \PyUltraLight by reproducing results from previous studies of ULDM dynamics, demonstrating interference effects and effective repulsive forces arising from the wavelike nature of ULDM.
 In addition we study the evolution of the velocity field of a solitonic core orbiting within a Newtonian central potential, showing that the stable orbital configuration is an irrotational Riemann-S ellipsoid.
-Finally, we show demonstrate that \PyUltraLight delivers sub-percent level energy conservation for a selection of dynamical scenarios.
+Finally, we demonstrate that \PyUltraLight delivers sub-percent level energy conservation for a selection of dynamical scenarios.
 
@@
 \end{align}
-where $x_1$ and $x_2$ are the initial centrals positions of the solitions, $v_1$ and $v_2$ are the soliton velocities, $\delta$ is a constant relative phase term and $\alpha_1=1/4 \ \alpha_2$, parameterising the density profiles as discussed in Section \ref{sec:soliton-profiles}.
+where $x_1$ and $x_2$ are the initial central positions of the solitions, $v_1$ and $v_2$ are the soliton velocities, $\delta$ is a constant relative phase term and $\alpha_2 = 4\, \alpha_1$, parameterising the density profiles as discussed in Section \ref{sec:soliton-profiles}.
 For convenience we set $v_1=-v_2$ and $x_1=-x_2$.
@@
 This corresponds to a time $t_{o}=\vert x_1/v_1\vert = \vert x_2/v_2\vert$.
-The dimensionless density is then given by
+\sh{Assuming no distortion of either soliton...
+(?)} The dimensionless density is then given by
 \begin{align}\label{eq:predicted-interference}
@@
 \end{align}
- Figure \ref{fig:interference} shows the dimensionless density profile at the time of maximal interference for two solitons with mass ratio $\mu=2$ and phase difference $\delta=0$.
+ Figure \ref{fig:interference} shows the dimensionless density profile at the time of maximal interference for two solitons with mass ratio $\mu=2$\sh{this mass ratio parameter $\mu$ doesn't seem to ever actually be defined, could we not just say mass ratio 2?} and phase difference $\delta=0$.
 The numerical result obtained using \PyUltraLight closely matches the theoretical prediction of equation \ref{eq:predicted-interference}.
@@
 Time progresses from left to right across each row and is indicated in code units in each frame.
-The effective repulsive force generated by a $\pi$ phase shift can be seen in the lower panel.}
+The effective repulsive force generated by a $\pi$ phase shift can be seen in the lower panel.\sh{I assume the contours are density but this is never explicitly stated}}
   \label{fig:repulsion}
@@
   \includegraphics[width=1.\textwidth,trim=1cm 1cm 0 1cm,clip]{riemann}
-  \caption{Solitonic core after one revolution of a central potential, demonstrating the internal velocity field of an irrotational Riemann-S ellipsoid (where the collective motion of the soliton has been subtracted from the velocity field).
+  \caption{Solitonic core after one revolution around a central potential, demonstrating the internal velocity field of an irrotational Riemann-S ellipsoid (where the collective motion of the soliton has been subtracted from the velocity field).
 Deformation of the spherical density profile into that of an ellipsoid can be seen, with extension along the radial direction (red arrow).
-The host:satellite mass ratio is approximately 55; simulation resolution is $256^3$.}
+The host:satellite mass ratio is approximately 55; simulation resolution is $256^3$.
+\sh{I assume the contours are density, but this is never explicitly stated.}}
   \label{fig:riemann}
@@
 
-\PyUltraLight allows the inclusion of a static potential as point-mass  at the centre of the simulation region.
+\PyUltraLight allows the inclusion of a static potential equivalent to a point-mass  at the centre of the simulation region.
 There is no backreaction on this mass as a result of the ULDM dynamics, and its ``mirror images'' within the periodic coordinate systems are not accounted for within the overall gravitational potential.
-The setup is thus somewhat unphysical but it provides a starting point for a study of the stability of satellite dark matter halos orbiting a much larger object.
-In particular, this includes the investigation of lifespans of dwarf satellite galaxies orbiting much larger objects (including the Milky Way) which are key to understanding whether ULDM models can resolve for the so-called missing satellites problem \cite{Weinberg2015}.
+The setup is thus somewhat unphysical but it provides a starting point for a study of the stability of satellite dark matter halos orbiting a much larger object \sh{I'm a bit confused by the use of the word "unphysical" here, is this not more physical than an infinite array of point masses?
+A 1/r potential, which we have, is the *more* physical potential}.
+In particular, this includes the investigation of lifespans of dwarf satellite galaxies orbiting much larger objects (including the Milky Way) which are a key to understanding whether ULDM models can resolve the so-called missing satellites problem \cite{Weinberg2015}.
 
 An extensive study of the tidal disruption of ULDM solitonic cores orbiting a central potential has recently been undertaken in \cite{Du2018} and we reproduce just one of their results here.
-When the Schr{\"o}dinger-Poisson system is recast as a system of hydrodynamical equations, the fluid velocity  at any point is the gradient of the smoothly-varying phase.
+When the Schr{\"o}dinger-Poisson system is recast as a system of hydrodynamical equations, the fluid velocity  at any point is the gradient of the smoothly-varying phase\sh{This correspondence is only approximate though, right?
+So the thing we call 'fluid velocity' must have zero curl, but the actual movement of mass doesn't necessarily need to be?}.
 Since the curl of a gradient vanishes identically, the velocity field of the Schr{\"o}dinger-Poisson system must be irrotational, and $\nabla\times\vec{v}=0$.
@@
 
-The family of Riemann-S ellipsoids describe uniformly rotating bodies which satisfy the condition of irrotationality \cite{Chandrasekhar1965}, and it is therefore the characteristic internal velocity field of such an ellipsoid which we expect to arise during the simulation.
+The family of Riemann-S ellipsoids describe uniformly rotating bodies which satisfy the condition of irrotationality \cite{Chandrasekhar1965} \sh{this sentence is really confusing, but I'm not sure of how to re-phrase it to make it less confusing so we might need to just accept the "rotating bodies...
+that are irrotational" confusingness}.
+Therefore we expect the characteristic internal velocity field of such an ellipsoid to arise during our simulations \sh{I split the previous sentence in two and re-ordered what is now the second sentence}.
 It is found in \cite{Du2018} that an initially spherical solitonic core without self-rotation will gradually spin up to form a tidally-locked irrotational Riemann-S ellipsoid when orbiting a host mass.
-We a reproduce this result using \PyUltraLight, with Figure \ref{fig:riemann} showing the internal velocity field of a solitonic satellite after one complete revolution around a host mass.
-The soliton has become elongated along the radial line connecting it to the host, indicating that it is tidally locked, while the internal velocity field qualitatively matches that of an irrotational Riemann-S ellipsoid within the tidal radius \cite{Daller2012}.
+We reproduce this result using \PyUltraLight.
+Figure \ref{fig:riemann} showing the internal velocity field of a solitonic satellite after one complete revolution around a host mass \sh{I spluit one sentence into two}.
+The soliton has become elongated along the radial line connecting it to the host, indicating that it is tidally locked, while the internal velocity field qualitatively matches that of an irrotational Riemann-S ellipsoid within the tidal radius \cite{Daller2012} \sh{This is stated as if the reader should be able to tell, but I would guess most of our readers haven't heard of a Riemann-S ellipsoid.
+I would suggest at least saying "see figure 2 of ref [37]", and perhaps even some more substantial description of hwat one expects a riemann s ellipsoid to look like.
+alternatively, we could not use that jargon and just appeal to the fact that the velocity field in our figure doesn't look rotational.}.
+
 It should be noted that the wider velocity field is not expected to be accurately predicted in a simulation of this kind.
@@
 Therefore, the velocity field far from the solitonic centre is not well-modelled in these simulations, and it may be expected that non-trivial self-interaction in the wider field would lead to unpredictable behaviour in these regions.
+\sh{This last paragraph was previously part of the preceding one.
+I found it a bit confusing because it reads like a defence of an anomaly in our own figure.
+I was going to suggest removing it but I think I now understand that it was put in as a warning to people who might use our code and get confused about the velocity field far from the centre of the solitons.
+We could either make that more clear or just hope that making it a separate paragraph does this implicitly}
 
@@
 Right: Evolution of the energy of the system, where times are indicated in correspondence with the snapshots of the density profile.
-All quantities are in dimensionless code units.}
+All quantities are in dimensionless code units.\sh{It might be worth repeating here what the energy code units are because it is stated in the text but not in the main 'code units' section.}}
   \label{fig:combined_1}
@@
   \includegraphics[width=0.9\textwidth,trim=0 0 2cm 2cm,clip]{egy_m=22.eps}
-  \caption{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}
+  \caption{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.\sh{Should probably add what energy code units are here too as people might not know to look at other figure}}
   \label{fig:binary}
@@
 This provides a test on the numerical performance of \PyUltraLight, and we find that even at relatively low spatial resolution we see sub-percent level energy conservation for all the dynamical scenarios considered here.
-In this Section we express the energy  the Schr{\"o}dinger-Poisson system in terms of the variables  $\psi$ and $\Phi$ and discuss its decomposition into individual constituents calculated separately within the code.
+In this Section we express the energy of the Schr{\"o}dinger-Poisson system in terms of the variables  $\psi$ and $\Phi$ and discuss its decomposition into individual constituents calculated separately within the code.
 We then present results for a variety of configurations.
@@
 The two terms in the integral \ref{eq:energy-tot} are calculated separately within the code.
-The first term is the gravitational potential energy of the Schr{\"o}dinger-Poisson system and as discussed in \cite{Hui2016}, the second term may be decomposed into  contributions which may be considered separately as kinetic and `quantum' energies, but for our purposes it is sufficient to consider only their combined contribution.
+The first term is the gravitational potential energy of the Schr{\"o}dinger-Poisson system, $E_{GP}$ \sh{I added $E_{GP}$ here because it is labelled like this in the figures and this makes it easier for people looking at the figures to find out what that quantity is.
+same for $E_K$ and $E_Q$ below}.
+As discussed in \cite{Hui2016}, the second term may be decomposed into contributions which may be considered separately as kinetic and `quantum' energies, $E_K$ and $E_Q$.
+However, for our purposes it is sufficient to consider only their combined contribution\sh{split one sentence into three}.
 When \PyUltraLight includes the central potential of a point mass located at the centre of the simulation grid we have additional energy contributions and calculate the  gravitational potential energy from self-interactions separately from the gravitational potential energy due to the central potential.
@@
 As the solitons reach the first point of closest approach, they become slightly deformed, exciting oscillatory modes which are manifest in the Figure as small scale oscillations superposed on the global behaviour.
-Figure \ref{fig:energy_change} demonstrates the relationship between the total integrated energy and the grid resolution for the same binary system of solitons used to generate Figure~\ref{fig:binary}.
+Figure \ref{fig:energy_change} demonstrates the relationship between the total integrated energy and the grid resolution for the same binary system of solitons used to generate Figure~\ref{fig:binary} \sh{I think people can probably work out what the y-axis is in this figure and figure 7, but it isn't ever explicitly stated.
+It might be good to state it here and/or in the figure captions}.
 Energy is conserved at sub-percent level even at low resolutions ($96^3$), and increasing grid resolution greatly improves accuracy.
@@
 Energy conservation within \PyUltraLight is excellent, at sub-percent level for simulations run at $128^3$, with even better performance as resolution is increased.
-The code captures complex phenomena resulting from the wave-like properties of ultralight dark matter, including as the interference patterns arising during high-velocity collisions of solitonic cores, and  effective forces observed in cases where the colliding cores are out of phase.
+The code captures complex phenomena resulting from the wave-like properties of ultralight dark matter, including the interference patterns arising during high-velocity collisions of solitonic cores, and  effective forces observed in cases where the colliding cores are out of phase.
 These phenomena can be clearly observed at relatively low spatial resolution, avoiding the need for high-performance computing infrastructure to study the fundamental behaviour of ULDM systems in simple configurations.
@@
 The current implementation of \PyUltraLight is already a useful tool for simulating dynamical ULDM systems and exploring their dynamics However, there is much scope for  improvement.
-In particular, future releases may incorporate a variable timestep and more sophisticated physics, including explicit self-interactions in the axion sector or additional matter components.
+In particular, future releases may incorporate a variable timestep and more sophisticated physics, including explicit self-interactions in the axion sector or additional matter components\sh{We could also mention the inclusion of other types of central potentials}.
 \PyUltraLight is publicly available under a BSD license.
@@
 
-We thank Nathan Musoke, Jens Niemeyer (and group -- names?), for valuable discussions, and thank Miro Erkintalo for advice on Schr{\"o}dinger-Poisson solvers  in optical systems.
+We thank Nathan Musoke, Jens Niemeyer (and group -- names?)\sh{The only other person I think we've explicitly spoken to about stuff in this paper is Xiaolong Du (and maybe David Marsh by the time we submit if he replies to your email}, for valuable discussions, and thank Miro Erkintalo for advice on Schr{\"o}dinger-Poisson solvers  in optical systems.
 We acknowledge support from the Marsden Fund of the Royal Society of New Zealand, the use of the New Zealand eScience Infrastructure (NeSI) high-performance computing facilities, which are funded jointly by NeSI's collaborator institutions and through the Ministry of Business, Innovation \& Employment's Research Infrastructure programme \url{https://www.nesi.org.nz}, (PLUS SPECIFIC NAMES OF PEOPLE AT NESI WHO HELPED??).
-(SHAUN -- PEOPLE WHO PUT YOU UP ON TRAVEL??) (FABER -- any names of yours?)
+(SHAUN -- PEOPLE WHO PUT YOU UP ON TRAVEL??) (FABER -- any names of yours?)\sh{I don't know if we received any explicit help from particular people at NESI.
+I've been put up in places but not relating to work on this.
+Jens and I discussed this at a conference, but my attendance at the conference was paid for by Auckland}
 
@@
 \PyUltraLight makes use of the pyFFTW pythonic wrapper around the FFTW C-based fast Fourier transform libraries.
-Both pyFFTW and FFTW are freely-available and \PyUltraLight has been used successfully on both Mac OS and Linux  systems.
+Both pyFFTW and FFTW are freely-available and \PyUltraLight has been used successfully on both Mac OS and Linux  systems\sh{I've also used it on the cluster quite easily}.
 We welcome advice and feedback from users.

### 2018-06-13 draft.tex
EK's first responses to Shaun's comments. As yet incomplete - more to follow.

@@
 Alternatively, if we consider a non-expanding universe, these equalities are true by definition.
-In either case, we can remove the second \sh{third?}term in equation \ref{eq:linear} entirely, such that all remaining derivatives are second order in space or time \sh{this sentence confuses me because it seems to imply we've removed terms with higher order derivatives, which isn't true}.
-
-In the non-relativistic limit, Equation \ref{eq:linear} can be treated using the WKB approximation\footnote{For a detailed explication of the WKB approximation in the non-relativistic limit, see \cite{Young2015}.}.
+In either case, we can remove the third term in equation \ref{eq:linear}.
+The resulting equation can then be analysed using WKB methods in the non-relativistic regime \footnote{For a detailed explication of the WKB approximation in the non-relativistic limit, see \cite{Young2015}.}.
 This allows us to write an ansatz solution for the field $\phi$:
@@
 \end{equation}
-We have thus shown that $\psi$ satisfies the Schr{\"o}dinger equation in this limit, which is interpreted as the unique macroscopic wavefunction of a Bose-Einstein condensate \sh{I'm also not sure I understand this sentence.
-There are other equations that describe other Bose-Einstein condensates (e.g.
-other N-L Schrod equations with different potentials.) I suggest just removing the word unique.}.
-It follows that the particle \sh{suggest include 'number' after particle} density of the condensate is given by $\vert\psi\vert^2$, so its mass density is simply $m\vert\psi\vert^2$.
+We have thus shown that $\psi$ satisfies the Schr{\"o}dinger equation in this limit, which is interpreted as the macroscopic wavefunction of a Bose-Einstein condensate.
+It follows that the particle number density of the condensate is given by $\vert\psi\vert^2$, so its mass density is simply $m\vert\psi\vert^2$.
 The local gravitational potential thus satisfies the Poisson equation,
@@
 It is helpful to recast the Schr{\"o}dinger-Poisson system (equations \ref{eq:schrodinger} and \ref{eq:poisson}) in terms of adimensional quantities.
-In keeping with Refs~\cite{Schive2014,Paredes2016} we introduce length, time, and mass scales as follows\sh{I think I fixed a typo in equation 2.8, but please check.
-I added the 1/2 in the exponent on the RHS}:
+In keeping with Refs~\cite{Schive2014,Paredes2016} we introduce length, time, and mass scales as follows:
 \begin{align}
-    \CMcal{L}&=\left(\frac{8\pi\hbar^2}{3 m^2H_0^2\Omega_{m_0}}\right)^{\frac{1}{4}}\approx121\left(\frac{10^{-23}\operatorname{eV}}{m}\right)^{1/2}\operatorname{kpc},\label{eq:length}\\
+    \CMcal{L}&=\left(\frac{8\pi\hbar^2}{3 m^2H_0^2\Omega_{m_0}}\right)^{\frac{1}{4}}\approx121\left(\frac{10^{-23}\operatorname{eV}}{m}\right)^{\frac{1}{2}}\operatorname{kpc},\label{eq:length}\\
     \CMcal{T}&=\left(\frac{8\pi}{3 H_0^2\Omega_{m_0}}\right)^{\frac{1}{2}}\approx75.5 \operatorname{Gyr},\label{eq:time}\\
@@
 \end{equation}
-where $\mathcal{T}$ is the time-ordering symbol \sh{I think I need this explained to me.
-I might just have a hole in my maths knowledge regarding what the $\mathcal{T}$ means in this context.
-I'm happy to go read up and fill that hole though so a reference might be enough.}.
-For a sufficiently small timestep $h$,  the trapezoidal rule gives the approximation
+where $\mathcal{T}$ is the time-ordering symbol. For a sufficiently small timestep $h$,  the trapezoidal rule gives the approximation
 \begin{equation}
@@
 
-The linear differential operator in equation \ref{eq:approx-time-ev} acts naturally in Fourier space, while the nonlinear potential term is simplest to evaluate in position space and by splitting the exponential we can evaluate each term in its natural domain.
-For small timesteps we assume that the dispersive and nonlinear terms act independently \sh{I don't understand what 'independently' means in this context.
-Can 3.4 not just be an ansatz that is proven as valid by 3.5?},  adopting the specific evolution scheme
+The linear differential operator in equation \ref{eq:approx-time-ev} acts naturally in Fourier space, while the nonlinear potential term is simplest to evaluate in position space.
+By splitting the exponential we can evaluate each term in its natural domain.
+Such a splitting is valid when the timestep is small, and is represented as
 %
@@
 The final operation in equation \ref{eq:schematic-psi}  only changes the phase of  $\psi$, so we could replace $\psi(\vec{x},t_{i})$ with $\psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)$ in equation \ref{eq:schematic-phi} with no change in meaning.
-\PyUltraLight makes an additional simplification to the symmetrised split-step Fourier method by combining the consecutive half-steps in the nonlinear term into a single full step \sh{I changed this sentence a bit to make it more readable, but am happy for it to be reverted if other's prefer the original form}.
+\PyUltraLight makes an additional simplification to the symmetrised split-step Fourier method by combining the consecutive half-steps in the nonlinear term into a single full step.
 Consequently, only the first and last operations involve actual half steps.
@@
 In many finite-differencing schemes for  partial differential equations the Courant-Friedrichs-Lewy (CFL) condition is a necessary condition for convergence \cite{Ajaib2013}, and puts an upper bound on the timestep as a function of the gridsize.
-The physical origin of the CFL condition is the finite speed of propagating information  in the evolving system \sh{split one sentence into two}.
+The physical origin of the CFL condition is the finite speed of propagating information  in the evolving system.
 Consequently, this condition is often cited in numerical analysis of ULDM physics governed by the Schr{\"o}dinger-Poisson system, see e.g.
@@
 In practice this can yield a very demanding constraint on the timestep and for many scenarios the actual velocity is far less than this hypothetical limit.
-Consequently, the timestep can be  significantly increased at the user's discretion but convergence testing on a case-by-case basis is strongly encouraged.\sh{I wonder whether it is worth mentioning that the two pieces of the split step method are exact to arbitrary large times when the other piece isn't present.
-Also, I wonder whether it is worth including a "boost factor" to the time step in the jupyer-notebook to aid people in making that change?}
+Consequently, the timestep can be  significantly increased at the user's discretion by changing the `step\textunderscore factor' parameter in the preliminary section of the Jupyter notebook.
+When this is done, convergence testing on a case-by-case basis is strongly encouraged.\sh{I wonder whether it is worth mentioning that the two pieces of the split step method are exact to arbitrary large times when the other piece isn't present.}\ek{I'm not quite sure what you mean by 'when the other piece isn't present', but I have a feeling that bringing this up here might cause confusion}
 
@@
 We expect that the interference effects will be maximised when two components of the wavefunction are centred at the same location, such that $x_1+v_1t=-x_2-v_2t=0$.
-This corresponds to a time $t_{o}=\vert x_1/v_1\vert = \vert x_2/v_2\vert$.
-\sh{Assuming no distortion of either soliton...
-(?)} The dimensionless density is then given by
+This corresponds to a time $t_{o}=\vert x_1/v_1\vert = \vert x_2/v_2\vert$, where in this simplified model we do not account for distortions caused by the accelerating or compactifying effects that the gravitational interaction has on the soliton profiles as they approach one another.
+The dimensionless density is then given by
 \begin{align}\label{eq:predicted-interference}
@@
 \end{align}
- Figure \ref{fig:interference} shows the dimensionless density profile at the time of maximal interference for two solitons with mass ratio $\mu=2$\sh{this mass ratio parameter $\mu$ doesn't seem to ever actually be defined, could we not just say mass ratio 2?} and phase difference $\delta=0$.
+ Figure \ref{fig:interference} shows the dimensionless density profile at the time of maximal interference for two solitons with mass ratio 2 and phase difference $\delta=0$.
 The numerical result obtained using \PyUltraLight closely matches the theoretical prediction of equation \ref{eq:predicted-interference}.
@@
   \caption{Head-on collisions between solitons of mass 20, initial relative velocity 20, and initial separation 1.2 in code units.
+Contours of constant density are plotted to demonstrate the deformation of the soliton profiles as they converge.
 Time progresses from left to right across each row and is indicated in code units in each frame.
-The effective repulsive force generated by a $\pi$ phase shift can be seen in the lower panel.\sh{I assume the contours are density but this is never explicitly stated}}
+The effective repulsive force generated by a $\pi$ phase shift can be seen in the lower panel.}
   \label{fig:repulsion}
@@
 There is no backreaction on this mass as a result of the ULDM dynamics, and its ``mirror images'' within the periodic coordinate systems are not accounted for within the overall gravitational potential.
-The setup is thus somewhat unphysical but it provides a starting point for a study of the stability of satellite dark matter halos orbiting a much larger object \sh{I'm a bit confused by the use of the word "unphysical" here, is this not more physical than an infinite array of point masses?
-A 1/r potential, which we have, is the *more* physical potential}.
+While a potential of this form does not accurately emulate that which we might expect from a realistic galaxy or dark matter halo, it provides a starting point for a study of the stability of satellite dark matter halos orbiting a much larger object \sh{I'm a bit confused by the use of the word "unphysical" here, is this not more physical than an infinite array of point masses?
+A 1/r potential, which we have, is the *more* physical potential}\ek{I've removed the terminology `unphysical', and instead just noted that this is a very idealised form of potential.
+What was meant by `unphysical' was really that this potential is not sourced by matter which takes part in the dynamics, it doesn't have the NFW-like scaling at large r, and that it is vulnerable to yielding nonsensical dynamics if PBCs aren't properly accounted for (although this latter issue isn't so much to do with the potential but rather the intrinsic limitations of PBCs)}.
 In particular, this includes the investigation of lifespans of dwarf satellite galaxies orbiting much larger objects (including the Milky Way) which are a key to understanding whether ULDM models can resolve the so-called missing satellites problem \cite{Weinberg2015}.

### 2018-06-14 draft.tex
Changes made in response to Shaun's comments. A few notes added.

@@
 
-\abstract{\PyUltraLight  simulates the dynamics of ultralight dark matter in a static, non-expanding \sh{are static and non-expanding not exactly rhe same?} background.
+\abstract{\PyUltraLight  simulates the dynamics of ultralight dark matter in a non-expanding background.
 \PyUltraLight can describe the evolution of several interacting ultralight dark matter halos  or one or more halos orbiting a central, fixed Newtonian potential, the latter scenario corresponding to dwarf galaxies orbiting a massive central galaxy.
@@
 In particular, ultralight dark matter (ULDM), also known as fuzzy dark matter (FDM), or BEC dark matter, is an increasingly well-studied possibility; for a recent review of the potential advantages and characteristic attributes of this scenario see Ref~\cite{Hui2016}.
-ULDM models are well motivated by fundamental theories possessing approximate shift symmetries\sh{suggest include ", e.g.
-an axion" with references.
-Otherwise this sentence seems a bit out of place, hinting at something without being explicit.}.
+ULDM models are well motivated by fundamental theories possessing approximate shift symmetries such as the theory of the QCD axion \cite{Kim2010, Marsh2016}.
 Moreover, ULDM can naturally resolve the small-scale problems of $\Lambda$CDM as the Heisenberg uncertainty principle suppresses gravitational collapse on length scales shorter than the de Broglie wavelength of the ULDM particle.
@@
 The existence of an extremely light scalar field, minimally coupled to gravity, is the central premise on which ULDM models are predicated.
-Within the ULDM framework, it is proposed that cosmological structure formation proceeds through the Bose-Einstein condensation of this scalar field \sh{I don't understand this sentence.
-There is a Bose-Einstein condensate and it clusters to form structures, but the sentence seems to suggest that the clustering and condensating are the same thing?}.
-In this Section  \re{I would normally give Section a capital letter in this context} \ek{I would be inclined to only capitalise `section' in instances where it appears as a proper noun, i.e.
-"In Section X we showed...", where we are referring to a specifically named section, numbered X.
-It seems to me that in this particular context `section' is a common noun.
-Happy to capitalise all instances of the word if this is your recommendation.}
-\re{I agree, but thought ``this Section'' was an example of it being used as a proper noun.
-But it is not a huge deal either way.}\ek{The way I would tend to think of it is by analogy with the following: "This person studies physics." vs.
-"Emily studies physics." In the former sentence, "person" is used as a description rather than a title, so would not be capitalised.
-Similar to "Richard is a professor." vs.
-"This is Professor Richard."}
-we derive the (non-relativistic) dynamical equations of the single macroscopic condensate wavefunction, namely, the Schr{\"o}dinger-Poisson equations.
-We begin with the action functional for a scalar field, $\phi$, minimally coupled to gravity and in the absence of self-interactions,
+Within the ULDM framework, it is proposed that this scalar field exists as a Bose-Einstein condensate, described by a single wavefunction which is governed by the Schr{\"o}dinger-Poisson coupled differential equations.
+We begin by deriving this system of equations as a non-relativistic weak-field limit of a more general theory.
+We start from the action functional for a scalar field, $\phi$, minimally coupled to gravity and in the absence of self-interactions,
 %
@@
 \begin{equation}\label{eq:density}
-    \rho=\frac{\CMcal{M}}{\CMcal{L}^3}\vert\psi\vert^2.
+    \rho=\CMcal{M}\CMcal{L}^{-3}\vert\psi\vert^2.
 \end{equation}
-\PyUltraLight\ works internally with these dimensionless quantities but receives initial conditions and generates output in physical units.
+By dimensional analysis, we can easily restore dimensionful units to any of the quantities calculated by the code. In particular, in the following sections it is to be understood that
+\begin{equation}\label{eq:density}
+    E=\CMcal{M}\CMcal{L}^2\CMcal{T}^{-2}\ E_{code}, \quad v=\CMcal{L}\CMcal{T}^{-1}\ v_{code}, 
+\end{equation}
+where $E$ and $v$ represent energy and velocity, respectively.
+\PyUltraLight\ works internally with these dimensionless quantities but can receive initial conditions and generates output in physical units.
 Henceforth, we will often refer to $\vert\psi\vert^2$  as the density, where it is understood that this is in fact a dimensionless quantity related to the physical mass density via the constant of proportionality given by equation \ref{eq:density}.
@@
 
-In place of the CFL condition, the \PyUltraLight timestep can be set via the fluid interpretation of the Schr{\"o}dinger-Poisson system. In dimensionless code units we can write 
-\begin{equation}
+In place of the CFL condition, the \PyUltraLight timestep can be set via the fluid interpretation of the Schr{\"o}dinger-Poisson system, discussed in greater detail in Section \ref{sec:disruption}.
+In dimensionless code units we define
+\begin{equation}\label{eq:madelung}
     \psi\equiv\sqrt{\rho}e^{i\theta}, \quad \vec{v}\equiv\boldsymbol{\nabla}\theta.
@@
 Consequently, the timestep can be  significantly increased at the user's discretion by changing the `step\textunderscore factor' parameter in the preliminary section of the Jupyter notebook.
-When this is done, convergence testing on a case-by-case basis is strongly encouraged.\sh{I wonder whether it is worth mentioning that the two pieces of the split step method are exact to arbitrary large times when the other piece isn't present.}\ek{I'm not quite sure what you mean by 'when the other piece isn't present', but I have a feeling that bringing this up here might cause confusion}
+When this is done, convergence testing on a case-by-case basis is strongly encouraged.\sh{I wonder whether it is worth mentioning that the two pieces of the split step method are exact to arbitrary large times when the other piece isn't present.}\ek{I'm not quite sure what you mean by `when the other piece isn't present', but I have a feeling that bringing this up here might cause confusion}
 
@@
   \includegraphics[width=1.\textwidth,trim=1cm 1cm 0 1cm,clip]{riemann}
-  \caption{Solitonic core after one revolution around a central potential, demonstrating the internal velocity field of an irrotational Riemann-S ellipsoid (where the collective motion of the soliton has been subtracted from the velocity field).
-Deformation of the spherical density profile into that of an ellipsoid can be seen, with extension along the radial direction (red arrow).
-The host:satellite mass ratio is approximately 55; simulation resolution is $256^3$.
-\sh{I assume the contours are density, but this is never explicitly stated.}}
+  \caption{Contours of constant density are shown for a solitonic core after one revolution around a central potential.
+The contours are superimposed upon a background representing the internal velocity field (where bulk motion has been subtracted).
+The velocity field is qualitatively that of an irrotational Riemann-S ellipsoid, and deformation of the density profile of the soliton along the radial direction (red arrow) can be seen.
+The host:satellite mass ratio is approximately 55; simulation resolution is $256^3$.}
   \label{fig:riemann}
@@
 An extensive study of the tidal disruption of ULDM solitonic cores orbiting a central potential has recently been undertaken in \cite{Du2018} and we reproduce just one of their results here.
-When the Schr{\"o}dinger-Poisson system is recast as a system of hydrodynamical equations, the fluid velocity  at any point is the gradient of the smoothly-varying phase\sh{This correspondence is only approximate though, right?
-So the thing we call 'fluid velocity' must have zero curl, but the actual movement of mass doesn't necessarily need to be?}.
-Since the curl of a gradient vanishes identically, the velocity field of the Schr{\"o}dinger-Poisson system must be irrotational, and $\nabla\times\vec{v}=0$.
+To do this, we again adopt the definitions \ref{eq:madelung}, namely
+\begin{equation}\label{eq:madelung_repeated}
+    \psi\equiv\sqrt{\rho}e^{i\theta}, \quad \vec{v}\equiv\boldsymbol{\nabla}\theta,
+\end{equation}
+where we are working in dimensionless code units.
+Using these definitions, the Schr{\"o}dinger-Poisson system can be recast in terms of hydrodynamical quantities in the so-called Madelung representation.
+The Madelung equations resemble the continuity and Euler equations of classical fluid dynamics, with the addition of a `quantum pressure' term accounting for resistance against gravitational collapse.
+The Madelung formalism is discussed in detail in \cite{Suarez2011, Suarez2015, Johnston2010, Kopp2017}.
+Because this hydrodynamical formulation defines the fluid velocity as the gradient of the phase of the field $\psi$, problems arise when $\psi=0$, where the phase is not well defined.
+Because of this issue, the Madelung and Schr{\"o}dinger representations are not strictly equivalent unless a quantisation condition is imposed, as discussed in \cite{Wallstrom1994}.
+We do not consider the subtleties of the Madelung representation here, as it is sufficient for our purposes to consider the fluid velocity in the region of a solitonic core, where no field nodes are present \footnote{It should be noted that, restoring dimensionful units, the fluid velocity $\vec{v}$ is related to the ususal quantum mechanical probability current $\vec{j}$ through $\vert\psi\vert^2\vec{v} = \vec{j} = \hbar/2mi\left[\psi^*\nabla\psi-\psi\nabla\psi^*\right]$}.
+For a discussion of the possible remedies to the `nodes problem', the reader is referred to Chapter 15.3 of \cite{Wyatt2005}.\sh{This correspondence is only approximate though, right?
+So the thing we call 'fluid velocity' must have zero curl, but the actual movement of mass doesn't necessarily need to be?} \ek{The equivalence between the Madelung representation and the SP system is exact in the regime where $\psi\neq 0$.
+In other words, we don't introduce any new approximations to get to the hydrodynamical representation, we just make a definition of a `fluid velocity'.
+So for a system that is governed by the SP equations, it is true to say that the field of these velocity vectors is strictly irrotational.
+If you look at my footnote, I've also written how the fluid velocity is related to the quantum mechanical probability current (or in this case, number density current since $\vert\psi\vert^2=n$).
+Therefore, if you multiply the fluid velocity by the mass density you do get the mass current.
+So yes, the irrotationality applies to the motion of the mass as well.
+In saying that, the SP system itself is only a non-relativistic weak field approximation anyway, so the correspondence of either representation to the fundamental physics is not exact in that sense.}.
+Where the Madelung representation is well defined, i.e.
+where the phase is a smoothly varying function, the velocity field of the Schr{\"o}dinger-Poisson system is strictly irrotational, $\nabla\times\vec{v}=0$.
 If a radially symmetric soliton is initialised in a circular orbit around a Newtonian potential, there will be initial transient behaviour as the spherical profile becomes elongated along the radial direction of the central potential.
-Meanwhile, the velocity field corresponding to the overall orbital motion of the soliton is superposed with the internal velocity field, combining so as to produce a net flow with vanishing curl.
+Meanwhile, the velocity field corresponding to the overall orbital motion of the soliton will be superposed with the internal velocity field, combining so as to produce a net flow with vanishing curl.
 
-The family of Riemann-S ellipsoids describe uniformly rotating bodies which satisfy the condition of irrotationality \cite{Chandrasekhar1965} \sh{this sentence is really confusing, but I'm not sure of how to re-phrase it to make it less confusing so we might need to just accept the "rotating bodies...
-that are irrotational" confusingness}.
-Therefore we expect the characteristic internal velocity field of such an ellipsoid to arise during our simulations \sh{I split the previous sentence in two and re-ordered what is now the second sentence}.
-It is found in \cite{Du2018} that an initially spherical solitonic core without self-rotation will gradually spin up to form a tidally-locked irrotational Riemann-S ellipsoid when orbiting a host mass.
+The family of Riemann-S ellipsoids describe non-axisymmetric uniformly rotating bodies whose internal velocity fields have vanishing curl \cite{Chandrasekhar1965}.
+\ek{Rephrased from previous sentence, removed the word `irrotational' - does this read ok?}.
+Therefore, it is the characteristic internal velocity field of a Riemann-S ellipsoid which we expect to arise during our simulation of a soliton orbiting a central mass.
+It is found in \cite{Du2018} that an initially spherical solitonic core without self-rotation will gradually spin up to form a tidally-locked ellipsoid with an irrotational internal velocity field when orbiting a host mass.
 We reproduce this result using \PyUltraLight.
-Figure \ref{fig:riemann} showing the internal velocity field of a solitonic satellite after one complete revolution around a host mass \sh{I spluit one sentence into two}.
-The soliton has become elongated along the radial line connecting it to the host, indicating that it is tidally locked, while the internal velocity field qualitatively matches that of an irrotational Riemann-S ellipsoid within the tidal radius \cite{Daller2012} \sh{This is stated as if the reader should be able to tell, but I would guess most of our readers haven't heard of a Riemann-S ellipsoid.
-I would suggest at least saying "see figure 2 of ref [37]", and perhaps even some more substantial description of hwat one expects a riemann s ellipsoid to look like.
-alternatively, we could not use that jargon and just appeal to the fact that the velocity field in our figure doesn't look rotational.}.
+Figure \ref{fig:riemann} shows the internal velocity field of a solitonic satellite after one complete revolution around a host mass.
+The soliton has become elongated along the radial line connecting it to the host, indicating that it is tidally locked, while the velocity field within the tidal radius is visibly irrotational and bears the qualitative trademarks of the Riemann-S ellipsoid as presented in Figure 2 of \cite{Daller2012}.
 
-It should be noted that the wider velocity field is not expected to be accurately predicted in a simulation of this kind.
-This is because the initial soliton density profile is defined only out to a given cutoff radius, beyond which the $\psi$ field value, and by extension its velocity field, is set identically to zero.
-Therefore, the velocity field far from the solitonic centre is not well-modelled in these simulations, and it may be expected that non-trivial self-interaction in the wider field would lead to unpredictable behaviour in these regions.
-\sh{This last paragraph was previously part of the preceding one.
-I found it a bit confusing because it reads like a defence of an anomaly in our own figure.
-I was going to suggest removing it but I think I now understand that it was put in as a warning to people who might use our code and get confused about the velocity field far from the centre of the solitons.
-We could either make that more clear or just hope that making it a separate paragraph does this implicitly}
-
-%\vspace{1em}
+It should be noted that the wider velocity field is not expected to be accurately predicted in a simulation of this kind, though the field within the tidal radius is well-modelled.
+This is because the initial soliton density profile is defined only out to a given cutoff radius, beyond which the $\psi$ field value is set identically to zero.
+As mentioned previously, the Madelung hydrodynamical formulation of the Schr{\"o}dinger-Poisson system is not valid where $\psi=0$.
+Because of this, we focus primarily on the internal velocity field within the high density region of the solitonic core.
+As we have seen, in this region \PyUltraLight is able to accurately reproduce the expected velocity field characteristics.\ek{Rephrased this paragraph}
 
@@
 Right: Evolution of the energy of the system, where times are indicated in correspondence with the snapshots of the density profile.
-All quantities are in dimensionless code units.\sh{It might be worth repeating here what the energy code units are because it is stated in the text but not in the main 'code units' section.}}
+All quantities are in dimensionless code units.\sh{It might be worth repeating here what the energy code units are because it is stated in the text but not in the main 'code units' section.}\ek{Now this is explicitly written in the code units section}}
   \label{fig:combined_1}
@@
   \includegraphics[width=0.9\textwidth,trim=0 0 2cm 2cm,clip]{egy_m=22.eps}
-  \caption{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.\sh{Should probably add what energy code units are here too as people might not know to look at other figure}}
+  \caption{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}
   \label{fig:binary}
@@
   \includegraphics[width=0.9\textwidth,trim=0 0.5cm 0 0,clip]{Total_Energy_Change.eps}
-  \caption{Demonstration of improving energy conservation with simulation resolution for a binary system of solitons orbiting their common centre of mass.}
+  \caption{Demonstration of improving energy conservation with simulation resolution for a binary system of solitons orbiting their common centre of mass.
+$\Delta E_{total}$ is the difference between the current integrated energy at a given timestep and the initial integrated energy for the simulation.
+The ratio of this difference to the initial integrated energy is plotted on the y axis for different resolutions.\ek{More explicit description of y axis added}}
   \label{fig:energy_change}
@@
   \includegraphics[width=0.9\textwidth,trim=0 0.5cm 0 0,clip]{Total_Energy_Change_Disruption.eps}
-  \caption{Demonstration of improving energy conservation with simulation resolution for a soliton undergoing significant tidal disruption in a Newtonian central potential.}
+  \caption{Demonstration of improving energy conservation with simulation resolution for a soliton undergoing significant tidal disruption in a Newtonian central potential.
+Again, $\Delta E_{total}$ is the difference between the current integrated energy at a given timestep and the initial integrated energy for the simulation.
+The ratio of this difference to the initial integrated energy is plotted on the y axis for each resolution.}
   \label{fig:energy_change_2}
@@
 The two terms in the integral \ref{eq:energy-tot} are calculated separately within the code.
-The first term is the gravitational potential energy of the Schr{\"o}dinger-Poisson system, $E_{GP}$ \sh{I added $E_{GP}$ here because it is labelled like this in the figures and this makes it easier for people looking at the figures to find out what that quantity is.
-same for $E_K$ and $E_Q$ below}.
+The first term is the gravitational potential energy of the Schr{\"o}dinger-Poisson system, $E_{GP}$.
 As discussed in \cite{Hui2016}, the second term may be decomposed into contributions which may be considered separately as kinetic and `quantum' energies, $E_K$ and $E_Q$.
-However, for our purposes it is sufficient to consider only their combined contribution\sh{split one sentence into three}.
+However, for our purposes it is sufficient to consider only their combined contribution.
 When \PyUltraLight includes the central potential of a point mass located at the centre of the simulation grid we have additional energy contributions and calculate the  gravitational potential energy from self-interactions separately from the gravitational potential energy due to the central potential.
@@
 As the solitons reach the first point of closest approach, they become slightly deformed, exciting oscillatory modes which are manifest in the Figure as small scale oscillations superposed on the global behaviour.
-Figure \ref{fig:energy_change} demonstrates the relationship between the total integrated energy and the grid resolution for the same binary system of solitons used to generate Figure~\ref{fig:binary} \sh{I think people can probably work out what the y-axis is in this figure and figure 7, but it isn't ever explicitly stated.
-It might be good to state it here and/or in the figure captions}.
+Figure \ref{fig:energy_change} demonstrates the relationship between the total integrated energy and the grid resolution for the same binary system of solitons used to generate Figure~\ref{fig:binary}.
+The vertical axis shows the ratio of the deviation in the total energy to the initial value of the energy, where the deviation is measured as the difference between the current and initial values.
 Energy is conserved at sub-percent level even at low resolutions ($96^3$), and increasing grid resolution greatly improves accuracy.
@@
 The current implementation of \PyUltraLight is already a useful tool for simulating dynamical ULDM systems and exploring their dynamics However, there is much scope for  improvement.
-In particular, future releases may incorporate a variable timestep and more sophisticated physics, including explicit self-interactions in the axion sector or additional matter components\sh{We could also mention the inclusion of other types of central potentials}.
+In particular, future releases may incorporate a variable timestep and more sophisticated physics, including explicit self-interactions in the axion sector or additional matter components\sh{We could also mention the inclusion of other types of central potentials}\ek{I had this in here originally but Richard's recommendation was that we leave it out as we don't want to `reveal our hand'}.
 \PyUltraLight is publicly available under a BSD license.
@@
 \PyUltraLight makes use of the pyFFTW pythonic wrapper around the FFTW C-based fast Fourier transform libraries.
-Both pyFFTW and FFTW are freely-available and \PyUltraLight has been used successfully on both Mac OS and Linux  systems\sh{I've also used it on the cluster quite easily}.
+Both pyFFTW and FFTW are freely-available and \PyUltraLight has been used successfully on both Mac OS and Linux systems, as well as a shared-memory cluster environment.
 We welcome advice and feedback from users.

### 2018-06-14 draft.tex
One more small edit.

@@
 where $H=\Dot{a}(t)/a(t)$.
-At late times in an expanding universe, $H\ll m$ \sh{I might be confused, but I'm reasonably sure the approximation is $H\ll m/\hbar$} and it is sufficient to set $H=0$ and $a(t)=1$ in equation \ref{eq:linear}.
+At late times in an expanding universe, $H\ll m/\hbar$ and it is sufficient to set $H=0$ and $a(t)=1$ in equation \ref{eq:linear}.
 This is a good approximation even at relatively high redshifts, including the epochs of early structure formation\sh{maybe worth mention that this approximation is equivalent to saying that the Hubble horizon is much larger than the compton wavelength?

### 2018-06-20 draft.tex
Removed old comments following discussion.

@@
 The notebook interface makes it simple to specify simulation parameters and  visualise the resulting output but performance-critical routines are managed via calls to computationally efficient compiled libraries.
-\PyUltraLight  runs on standard desktop hardware with support for shared memory mutlithreading and is available on GitHub.
-\re{small change}}
+\PyUltraLight  runs on standard desktop hardware with support for shared memory mutlithreading and is available on GitHub.}
 
@@
 At late times in an expanding universe, $H\ll m/\hbar$ and it is sufficient to set $H=0$ and $a(t)=1$ in equation \ref{eq:linear}.
-This is a good approximation even at relatively high redshifts, including the epochs of early structure formation\sh{maybe worth mention that this approximation is equivalent to saying that the Hubble horizon is much larger than the compton wavelength?
-- I'm not sure this condition is enough to neglect H though because in Nathan's code it is still true, but H is being included.
-It is somehow the de Broglie wavelength that is relevant, not just Compton, so a comparison also needs made to the $\nabla^2 \phi$ term}.
+This is a good approximation even at relatively high redshifts, including the epochs of early structure formation.
 Alternatively, if we consider a non-expanding universe, these equalities are true by definition.
@@
 There is no backreaction on this mass as a result of the ULDM dynamics, and its ``mirror images'' within the periodic coordinate systems are not accounted for within the overall gravitational potential.
-While a potential of this form does not accurately emulate that which we might expect from a realistic galaxy or dark matter halo, it provides a starting point for a study of the stability of satellite dark matter halos orbiting a much larger object \sh{I'm a bit confused by the use of the word "unphysical" here, is this not more physical than an infinite array of point masses?
-A 1/r potential, which we have, is the *more* physical potential}\ek{I've removed the terminology `unphysical', and instead just noted that this is a very idealised form of potential.
-What was meant by `unphysical' was really that this potential is not sourced by matter which takes part in the dynamics, it doesn't have the NFW-like scaling at large r, and that it is vulnerable to yielding nonsensical dynamics if PBCs aren't properly accounted for (although this latter issue isn't so much to do with the potential but rather the intrinsic limitations of PBCs)}.
+While a potential of this form does not necessarily accurately emulate that which we might expect from a realistic galaxy or dark matter halo, it provides a starting point for a study of the stability of satellite dark matter halos orbiting a much larger object.
 In particular, this includes the investigation of lifespans of dwarf satellite galaxies orbiting much larger objects (including the Milky Way) which are a key to understanding whether ULDM models can resolve the so-called missing satellites problem \cite{Weinberg2015}.
@@
 We do not consider the subtleties of the Madelung representation here, as it is sufficient for our purposes to consider the fluid velocity in the region of a solitonic core, where no field nodes are present \footnote{It should be noted that, restoring dimensionful units, the fluid velocity $\vec{v}$ is related to the ususal quantum mechanical probability current $\vec{j}$ through $\vert\psi\vert^2\vec{v} = \vec{j} = \hbar/2mi\left[\psi^*\nabla\psi-\psi\nabla\psi^*\right]$}.
-For a discussion of the possible remedies to the `nodes problem', the reader is referred to Chapter 15.3 of \cite{Wyatt2005}.\sh{This correspondence is only approximate though, right?
-So the thing we call 'fluid velocity' must have zero curl, but the actual movement of mass doesn't necessarily need to be?} \ek{The equivalence between the Madelung representation and the SP system is exact in the regime where $\psi\neq 0$.
-In other words, we don't introduce any new approximations to get to the hydrodynamical representation, we just make a definition of a `fluid velocity'.
-So for a system that is governed by the SP equations, it is true to say that the field of these velocity vectors is strictly irrotational.
-If you look at my footnote, I've also written how the fluid velocity is related to the quantum mechanical probability current (or in this case, number density current since $\vert\psi\vert^2=n$).
-Therefore, if you multiply the fluid velocity by the mass density you do get the mass current.
-So yes, the irrotationality applies to the motion of the mass as well.
-In saying that, the SP system itself is only a non-relativistic weak field approximation anyway, so the correspondence of either representation to the fundamental physics is not exact in that sense.}.
+For a discussion of the possible remedies to the `nodes problem', the reader is referred to Chapter 15.3 of \cite{Wyatt2005}.
 Where the Madelung representation is well defined, i.e.
@@
 The family of Riemann-S ellipsoids describe non-axisymmetric uniformly rotating bodies whose internal velocity fields have vanishing curl \cite{Chandrasekhar1965}.
-\ek{Rephrased from previous sentence, removed the word `irrotational' - does this read ok?}.
 Therefore, it is the characteristic internal velocity field of a Riemann-S ellipsoid which we expect to arise during our simulation of a soliton orbiting a central mass.
@@
 Because of this, we focus primarily on the internal velocity field within the high density region of the solitonic core.
-As we have seen, in this region \PyUltraLight is able to accurately reproduce the expected velocity field characteristics.\ek{Rephrased this paragraph}
+As we have seen, in this region \PyUltraLight is able to accurately reproduce the expected velocity field characteristics.
 
@@
 Right: Evolution of the energy of the system, where times are indicated in correspondence with the snapshots of the density profile.
-All quantities are in dimensionless code units.\sh{It might be worth repeating here what the energy code units are because it is stated in the text but not in the main 'code units' section.}\ek{Now this is explicitly written in the code units section}}
+All quantities are in dimensionless code units.}
   \label{fig:combined_1}
@@
 $\Delta E_{total}$ is the difference between the current integrated energy at a given timestep and the initial integrated energy for the simulation.
-The ratio of this difference to the initial integrated energy is plotted on the y axis for different resolutions.\ek{More explicit description of y axis added}}
+The ratio of this difference to the initial integrated energy is plotted on the y axis for different resolutions.}
   \label{fig:energy_change}

### 2018-06-20 draft.tex
Acknowledgements edited.

@@
 Consequently, the timestep can be  significantly increased at the user's discretion by changing the `step\textunderscore factor' parameter in the preliminary section of the Jupyter notebook.
-When this is done, convergence testing on a case-by-case basis is strongly encouraged.\sh{I wonder whether it is worth mentioning that the two pieces of the split step method are exact to arbitrary large times when the other piece isn't present.}\ek{I'm not quite sure what you mean by `when the other piece isn't present', but I have a feeling that bringing this up here might cause confusion}
+When this is done, convergence testing on a case-by-case basis is strongly encouraged.
 
@@
 The current implementation of \PyUltraLight is already a useful tool for simulating dynamical ULDM systems and exploring their dynamics However, there is much scope for  improvement.
-In particular, future releases may incorporate a variable timestep and more sophisticated physics, including explicit self-interactions in the axion sector or additional matter components\sh{We could also mention the inclusion of other types of central potentials}\ek{I had this in here originally but Richard's recommendation was that we leave it out as we don't want to `reveal our hand'}.
+In particular, future releases may incorporate a variable timestep and more sophisticated physics, including explicit self-interactions in the axion sector or additional matter components.
 \PyUltraLight is publicly available under a BSD license.
@@
 
-We thank Nathan Musoke, Jens Niemeyer (and group -- names?)\sh{The only other person I think we've explicitly spoken to about stuff in this paper is Xiaolong Du (and maybe David Marsh by the time we submit if he replies to your email}, for valuable discussions, and thank Miro Erkintalo for advice on Schr{\"o}dinger-Poisson solvers  in optical systems.
-We acknowledge support from the Marsden Fund of the Royal Society of New Zealand, the use of the New Zealand eScience Infrastructure (NeSI) high-performance computing facilities, which are funded jointly by NeSI's collaborator institutions and through the Ministry of Business, Innovation \& Employment's Research Infrastructure programme \url{https://www.nesi.org.nz}, (PLUS SPECIFIC NAMES OF PEOPLE AT NESI WHO HELPED??).
-(SHAUN -- PEOPLE WHO PUT YOU UP ON TRAVEL??) (FABER -- any names of yours?)\sh{I don't know if we received any explicit help from particular people at NESI.
-I've been put up in places but not relating to work on this.
-Jens and I discussed this at a conference, but my attendance at the conference was paid for by Auckland}
+We thank Nathan Musoke, Jens Niemeyer, Xiaolong Du, and David Marsh for valuable discussions, and thank Miro Erkintalo for advice on Schr{\"o}dinger-Poisson solvers  in optical systems.
+We acknowledge support from the Marsden Fund of the Royal Society of New Zealand, the use of the New Zealand eScience Infrastructure (NeSI) high-performance computing facilities, which are funded jointly by NeSI's collaborator institutions and through the Ministry of Business, Innovation \& Employment's Research Infrastructure programme \url{https://www.nesi.org.nz}.
 
@@
 % in short all the informations that clearly identify a document.
-
-\re{Check the bib options so we don't see all the Planck authors :-) }\ek{I've modified the bibliography style file so that only the first 3 authors will be  displayed if there are more than 3.
-I can't seem to find official JCAP guidance on this - Does it look OK to you in this format?}
 

### 2018-06-21 draft.tex
Moved footnotes.

@@
 In either case, we can remove the third term in equation \ref{eq:linear}.
-The resulting equation can then be analysed using WKB methods in the non-relativistic regime \footnote{For a detailed explication of the WKB approximation in the non-relativistic limit, see \cite{Young2015}.}.
-This allows us to write an ansatz solution for the field $\phi$:
+The resulting equation can then be analysed using WKB methods in the non-relativistic regime.
+\footnote{For a detailed explication of the WKB approximation in the non-relativistic limit, see \cite{Young2015}.} This allows us to write an ansatz solution for the field $\phi$:
 \begin{equation}\label{eq:ansatz}
@@
 Because of this issue, the Madelung and Schr{\"o}dinger representations are not strictly equivalent unless a quantisation condition is imposed, as discussed in \cite{Wallstrom1994}.
-We do not consider the subtleties of the Madelung representation here, as it is sufficient for our purposes to consider the fluid velocity in the region of a solitonic core, where no field nodes are present \footnote{It should be noted that, restoring dimensionful units, the fluid velocity $\vec{v}$ is related to the ususal quantum mechanical probability current $\vec{j}$ through $\vert\psi\vert^2\vec{v} = \vec{j} = \hbar/2mi\left[\psi^*\nabla\psi-\psi\nabla\psi^*\right]$}.
-For a discussion of the possible remedies to the `nodes problem', the reader is referred to Chapter 15.3 of \cite{Wyatt2005}.
+We do not consider the subtleties of the Madelung representation here, as it is sufficient for our purposes to consider the fluid velocity in the region of a solitonic core, where no field nodes are present.
+\footnote{It should be noted that, restoring dimensionful units, the fluid velocity $\vec{v}$ is related to the ususal quantum mechanical probability current $\vec{j}$ through $\vert\psi\vert^2\vec{v} = \vec{j} = \hbar/2mi\left[\psi^*\nabla\psi-\psi\nabla\psi^*\right]$} For a discussion of the possible remedies to the `nodes problem', the reader is referred to Chapter 15.3 of \cite{Wyatt2005}.
 Where the Madelung representation is well defined, i.e.

### 2018-06-22 draft.tex
minor changes to acks

@@
 
-We thank Nathan Musoke, Jens Niemeyer, Xiaolong Du, and David Marsh for valuable discussions, and thank Miro Erkintalo for advice on Schr{\"o}dinger-Poisson solvers  in optical systems.
-We acknowledge support from the Marsden Fund of the Royal Society of New Zealand, the use of the New Zealand eScience Infrastructure (NeSI) high-performance computing facilities, which are funded jointly by NeSI's collaborator institutions and through the Ministry of Business, Innovation \& Employment's Research Infrastructure programme \url{https://www.nesi.org.nz}.
+We thank Xiaolong Du, Lam Hui, David Marsh, Nathan Musoke, Jens Niemeyer, and Chanda Prescod-Weinstein for valuable discussions on axion / ULDM cosmology, and thank Miro Erkintalo for advice on Schr{\"o}dinger-Poisson solvers  in optical systems.
+We acknowledge support from the Marsden Fund of the Royal Society of New Zealand, and  the use of the New Zealand eScience Infrastructure (NeSI) high-performance computing facilities, which are funded jointly by NeSI's collaborator institutions and through the Ministry of Business, Innovation \& Employment's Research Infrastructure programme \url{https://www.nesi.org.nz}.
 
@@
 Both pyFFTW and FFTW are freely-available and \PyUltraLight has been used successfully on both Mac OS and Linux systems, as well as a shared-memory cluster environment.
-We welcome advice and feedback from users.
+\re{I don't think we have used this in a SHARED MEMORY environment} We welcome advice and feedback from users.
 

### 2018-07-03 draft.tex
Removed comment about shared mem
Changed repo location in download section

@@
 \PyUltraLight is publicly available under a BSD licence.
-The full repository, including supplementary files such as the code used to generate soliton profiles, is available on GitHub at {\url{ https://github.com/erckendall/PyUltraLight}}.
+The full repository, including supplementary files such as the code used to generate soliton profiles, is available on GitHub at {\url{https://github.com/auckland-cosmo/PyUltraLight}}.
 \PyUltraLight makes use of the pyFFTW pythonic wrapper around the FFTW C-based fast Fourier transform libraries.
 Both pyFFTW and FFTW are freely-available and \PyUltraLight has been used successfully on both Mac OS and Linux systems, as well as a shared-memory cluster environment.
-\re{I don't think we have used this in a SHARED MEMORY environment} We welcome advice and feedback from users.
+We welcome advice and feedback from users.
 

### 2018-07-09 draft.tex
last changes

@@
 
-\PyUltraLight is based on a symmetrised-split-step (leapfrog) solver for the time evolution, and uses a pseudospectral Fourier algorithm to solve the Poisson equation for the gravitational potential at each step.\footnote{A similar methodology was described in Ref.~\cite{Paredes2016}; to our knowledge this code has not been released.} This algorithm has $2^{nd}$ order accurate time integration steps and sub-percent level energy conservation, while the wavefunction normalisation  is conserved to machine precision.
+\PyUltraLight is based on a symmetrised-split-step (leapfrog) solver for the time evolution, and uses a pseudospectral Fourier algorithm to solve the Poisson equation for the gravitational potential at each step.\footnote{A similar methodology was described in Ref.~\cite{Paredes2016}; to our knowledge this code has not been released.
+Spectral methods are often used to solve the Poisson equation in large scale structure simulations, while the {\sc PSpectre} code \cite{Easther:2010qz} provides a pseudospectral solver for the evolution of the inflaton  and fields coupled to it during parametric resonance and preheating after inflation \cite{Amin:2010dc,Amin:2011hj,Zhou:2013tsa}.} This algorithm has $2^{nd}$ order accurate time integration steps and sub-percent level energy conservation, while the wavefunction normalisation  is conserved to machine precision.
 As a pseudospectral code, linear differential operators are computed by direct multiplication in the Fourier domain, while non-linear terms are evaluated in position space.

### 2018-07-11 draft.tex
ADS bib style file
References reformatted

@@
 \usepackage{url}
+\usepackage{aas_macros}
 \usepackage[justification=justified]{caption}
@@
 The realisation that there may be more to the universe than meets the eye is one of the most profound developments in 20$^{\textrm{th}}$ Century astronomy and astrophysics.
-However, while there are now multiple lines of evidence that dark matter outweighs baryonic matter by a ratio of approximately 5:1 \cite{Planck2015}, we have few clues regarding the physical nature of dark matter.
+However, while there are now multiple lines of evidence that dark matter outweighs baryonic matter by a ratio of approximately 5:1 \cite{Ade:2015xua}, we have few clues regarding the physical nature of dark matter.
 Much theoretical and experimental effort has focused on  WIMP models, motivated by their consistency with supersymmetric extensions to the Standard Model and their relatively simple dynamics.
-However, advanced direct-detection experiments are putting increasingly tight constraints on the WIMP parameter space \cite{Tan2016, Akerib2017} and $\Lambda$CDM cosmology with simple, pressureless, noninteracting dark matter (a class including simple WIMP scenarios) is potentially at odds with observations at small astrophysical scales \cite{Bull2016}.
+However, advanced direct-detection experiments are putting increasingly tight constraints on the WIMP parameter space \cite{Tan:2016zwf, Akerib:2016vxi} and $\Lambda$CDM cosmology with simple, pressureless, noninteracting dark matter (a class including simple WIMP scenarios) is potentially at odds with observations at small astrophysical scales \cite{Bull:2015stt}.
 
 The potential shortcomings of simple cold dark matter scenarios motivate investigations of more novel dark matter scenarios.
-In particular, ultralight dark matter (ULDM), also known as fuzzy dark matter (FDM), or BEC dark matter, is an increasingly well-studied possibility; for a recent review of the potential advantages and characteristic attributes of this scenario see Ref~\cite{Hui2016}.
-ULDM models are well motivated by fundamental theories possessing approximate shift symmetries such as the theory of the QCD axion \cite{Kim2010, Marsh2016}.
+In particular, ultralight dark matter (ULDM), also known as fuzzy dark matter (FDM), or BEC dark matter, is an increasingly well-studied possibility; for a recent review of the potential advantages and characteristic attributes of this scenario see Ref~\cite{Hui:2016ltb}.
+ULDM models are well motivated by fundamental theories possessing approximate shift symmetries such as the theory of the QCD axion \cite{Kim:2008hd, Marsh:2015xka}.
 Moreover, ULDM can naturally resolve the small-scale problems of $\Lambda$CDM as the Heisenberg uncertainty principle suppresses gravitational collapse on length scales shorter than the de Broglie wavelength of the ULDM particle.
-In this regime the mass of the ULDM particle becomes correlated with astrophysical observables; if it is on the order of $10^{-22}$~eV, structure is suppressed at kiloparsec scales and below \cite{Hu2000}.
+In this regime the mass of the ULDM particle becomes correlated with astrophysical observables; if it is on the order of $10^{-22}$~eV, structure is suppressed at kiloparsec scales and below \cite{Hu:2000ke}.
 
 Given the presence of a fundamental  lengthscale, the behaviour of ULDM is more complex than that of simple dark matter scenarios whose cosmologically relevant interactions are purely gravitational.
-Physically, the effective short-scale pressure and  condensate-like properties of ULDM  create new dynamical possibilities for ULDM scenarios, such as  purely pressure supported soliton-like solutions  \cite{Marsh2015} and superposition or interference during interactions between condensate-like halos \cite{Schwabe2016}.
+Physically, the effective short-scale pressure and  condensate-like properties of ULDM  create new dynamical possibilities for ULDM scenarios, such as  purely pressure supported soliton-like solutions  \cite{Marsh:2015wka} and superposition or interference during interactions between condensate-like halos \cite{Schwabe:2016rze}.
 Consequently, modelling dark matter dynamics in ULDM scenarios is more challenging than in simple cold dark matter models, but is critical to understanding the physical consequences of ULDM models.
@@
 One widely used approach is the Madelung fluid formulation of the Schr{\"o}dinger-Poisson system \cite{Madelung1926} which has a quantum pressure term that can be treated numerically in a variety of ways.
-In Ref.~\cite{Zhang2018}, the cosmological code \textsc{gadget} \cite{Springel2005} is modified to treat the quantum pressure as an effective particle-particle interaction and the resulting code, \textsc{axion-gadget} is publicly available \cite{axion-gadget}.
-Ref.~\cite{Nori2018} modifies a non-public extension of \textsc{gadget}, \textsc{p-gadget3} to treat the quantum pressure term via smoothed-particle hydrodynamics (SPH) routines.
-The SPH approach is also  used in Ref.~\cite{Mocz2015}, while a particle-mesh approach was implemented in \cite{Veltmaat2016}.
-N\textsc{yx} \cite{Almgren2013} was modified in \cite{Schwabe2016} to study merging ULDM solitonic cores, G\textsc{alacticus} \cite{Benson2012} was modified in \cite{Du2017} to study the effects of tidal stripping and dynamical friction on ULDM halos, \textsc{arepo} \cite{Springel2010} was modified in \cite{Mocz2017} to study the core-mass relationship and turbulence characteristics of ULDM halos, and \textsc{gamer} \cite{Schive2010, gamer} was modified \cite{Schive2014_b} to perform a detailed study of structure formation in ULDM cosmologies.
+In Ref.~\cite{Zhang:2016uiy}, the cosmological code \textsc{gadget} \cite{Springel:2005mi} is modified to treat the quantum pressure as an effective particle-particle interaction and the resulting code, \textsc{axion-gadget} is publicly available \cite{axion-gadget}.
+Ref.~\cite{Nori:2018hud} modifies a non-public extension of \textsc{gadget}, \textsc{p-gadget3} to treat the quantum pressure term via smoothed-particle hydrodynamics (SPH) routines.
+The SPH approach is also  used in Ref.~\cite{Mocz:2015sda}, while a particle-mesh approach was implemented in \cite{Veltmaat:2016rxo}.
+N\textsc{yx} \cite{Almgren:2013sz} was modified in \cite{Schwabe:2016rze} to study merging ULDM solitonic cores, G\textsc{alacticus} \cite{Benson:2010kx} was modified in \cite{Du:2016zcv} to study the effects of tidal stripping and dynamical friction on ULDM halos, \textsc{arepo} \cite{Springel:2009aa} was modified in \cite{Mocz:2017wlg} to study the core-mass relationship and turbulence characteristics of ULDM halos, and \textsc{gamer} \cite{Schive:2009hw, gamer} was modified \cite{Schive:2014dra} to perform a detailed study of structure formation in ULDM cosmologies.
 
@@
 
-\PyUltraLight is based on a symmetrised-split-step (leapfrog) solver for the time evolution, and uses a pseudospectral Fourier algorithm to solve the Poisson equation for the gravitational potential at each step.\footnote{A similar methodology was described in Ref.~\cite{Paredes2016}; to our knowledge this code has not been released.
+\PyUltraLight is based on a symmetrised-split-step (leapfrog) solver for the time evolution, and uses a pseudospectral Fourier algorithm to solve the Poisson equation for the gravitational potential at each step.\footnote{A similar methodology was described in Ref.~\cite{Paredes:2015wga}; to our knowledge this code has not been released.
 Spectral methods are often used to solve the Poisson equation in large scale structure simulations, while the {\sc PSpectre} code \cite{Easther:2010qz} provides a pseudospectral solver for the evolution of the inflaton  and fields coupled to it during parametric resonance and preheating after inflation \cite{Amin:2010dc,Amin:2011hj,Zhou:2013tsa}.} This algorithm has $2^{nd}$ order accurate time integration steps and sub-percent level energy conservation, while the wavefunction normalisation  is conserved to machine precision.
@@
 In either case, we can remove the third term in equation \ref{eq:linear}.
-The resulting equation can then be analysed using WKB methods in the non-relativistic regime.
-\footnote{For a detailed explication of the WKB approximation in the non-relativistic limit, see \cite{Young2015}.} This allows us to write an ansatz solution for the field $\phi$:
+The resulting equation can then be analysed using WKB methods in the non-relativistic regime.\footnote{For a detailed explication of the WKB approximation in the non-relativistic limit, see \cite{Young2015}.} This allows us to write an ansatz solution for the field $\phi$:
 \begin{equation}\label{eq:ansatz}
@@
 It is helpful to recast the Schr{\"o}dinger-Poisson system (equations \ref{eq:schrodinger} and \ref{eq:poisson}) in terms of adimensional quantities.
-In keeping with Refs~\cite{Schive2014,Paredes2016} we introduce length, time, and mass scales as follows:
+In keeping with Refs~\cite{Schive:2014hza,Paredes:2015wga} we introduce length, time, and mass scales as follows:
 \begin{align}
@@
 After an inverse Fourier transform a final half timestep is taken with the updated nonlinear term acting in position space to give the new $\psi$ field configuration.
-This procedure is known as the symmetrised split-step Fourier method, and  used widely in fields such as nonlinear fiber optics \cite{Agrawal2013, Zhang2018, Sinkin2003}.
+This procedure is known as the symmetrised split-step Fourier method, and  used widely in fields such as nonlinear fiber optics \cite{Agrawal2013, Zhang:2008, Sinkin2003}.
 
@@
 
-The soliton profile used to generate the initial conditions in \PyUltraLight is found by first imposing spherical symmetry in the Schr{\"o}dinger-Poisson equations and assuming time independence in the radial density profile \cite{Paredes2016}:
+The soliton profile used to generate the initial conditions in \PyUltraLight is found by first imposing spherical symmetry in the Schr{\"o}dinger-Poisson equations and assuming time independence in the radial density profile \cite{Paredes:2015wga}:
 %
@@
 
-In many finite-differencing schemes for  partial differential equations the Courant-Friedrichs-Lewy (CFL) condition is a necessary condition for convergence \cite{Ajaib2013}, and puts an upper bound on the timestep as a function of the gridsize.
+In many finite-differencing schemes for  partial differential equations the Courant-Friedrichs-Lewy (CFL) condition is a necessary condition for convergence \cite{Ajaib:2013oua}, and puts an upper bound on the timestep as a function of the gridsize.
 The physical origin of the CFL condition is the finite speed of propagating information  in the evolving system.
 Consequently, this condition is often cited in numerical analysis of ULDM physics governed by the Schr{\"o}dinger-Poisson system, see e.g.
-Ref.~\cite{Schwabe2016}.
+Ref.~\cite{Schwabe:2016rze}.
 Strictly speaking, however, the CFL condition is applicable to codes using finite differencing methods for spatial derivatives, and is thus not applicable to pseudospectral codes like \PyUltraLight.
@@
 
-Following \cite{Schwabe2016}, we consider the head-on collision of two solitions with mass ratio $\mu=2$ and high relative velocity.
+Following \cite{Schwabe:2016rze}, we consider the head-on collision of two solitions with mass ratio $\mu=2$ and high relative velocity.
 While we work in dimensionless code units, it should be noted that a dimensionful velocity can be restored from the code velocity by multiplying through by $\CMcal{L}\CMcal{T}^{-1}$, the scale parameters defined in Equations \ref{eq:length} and \ref{eq:time}.
@@
 We do not expect an exact match, but we have verified that \PyUltraLight qualitatively reproduces the wave interference effects of the ULDM model.
-With the exception of \cite{Schwabe2016}, few studies of ULDM dynamics have investigated the interference patterns generated by colliding solitons in this way.
+With the exception of \cite{Schwabe:2016rze}, few studies of ULDM dynamics have investigated the interference patterns generated by colliding solitons in this way.
 In some cases, this is because the algorithm employed to simulate the dynamics is not capable of reproducing such effects.
-An example of this is given in \cite{Veltmaat2016}, where it is demonstrated that the coarse-grained nature of the particle-mesh method renders the algorithm incapable of reproducing detailed interference patterns such as those shown here.
+An example of this is given in \cite{Veltmaat:2016rxo}, where it is demonstrated that the coarse-grained nature of the particle-mesh method renders the algorithm incapable of reproducing detailed interference patterns such as those shown here.
 %\re{read to here}
@@
 
-As demonstrated in \cite{Paredes2016}, the wavelike properties of ULDM give rise to effective forces which can dramatically affect the dynamics of core collisions.
+As demonstrated in \cite{Paredes:2015wga}, the wavelike properties of ULDM give rise to effective forces which can dramatically affect the dynamics of core collisions.
 These effective forces arise as a result of interference phenomena, rather than because of any local interactions the ULDM model might  incorporate.
@@
 Dissimilarly, in the case where there is no phase shift, the solitons readily collide and merge to form a single contiguous density profile prior to re-separating.
-Further discussion of this phenomenon and its possible observational consequences can be found in \cite{Paredes2016}.
+Further discussion of this phenomenon and its possible observational consequences can be found in \cite{Paredes:2015wga}.
 
@@
 While a potential of this form does not necessarily accurately emulate that which we might expect from a realistic galaxy or dark matter halo, it provides a starting point for a study of the stability of satellite dark matter halos orbiting a much larger object.
-In particular, this includes the investigation of lifespans of dwarf satellite galaxies orbiting much larger objects (including the Milky Way) which are a key to understanding whether ULDM models can resolve the so-called missing satellites problem \cite{Weinberg2015}.
+In particular, this includes the investigation of lifespans of dwarf satellite galaxies orbiting much larger objects (including the Milky Way) which are a key to understanding whether ULDM models can resolve the so-called missing satellites problem \cite{Weinberg:2013aya}.
 
-An extensive study of the tidal disruption of ULDM solitonic cores orbiting a central potential has recently been undertaken in \cite{Du2018} and we reproduce just one of their results here.
+An extensive study of the tidal disruption of ULDM solitonic cores orbiting a central potential has recently been undertaken in \cite{Du:2018zrg} and we reproduce just one of their results here.
 To do this, we again adopt the definitions \ref{eq:madelung}, namely
@@
 The Madelung equations resemble the continuity and Euler equations of classical fluid dynamics, with the addition of a `quantum pressure' term accounting for resistance against gravitational collapse.
-The Madelung formalism is discussed in detail in \cite{Suarez2011, Suarez2015, Johnston2010, Kopp2017}.
+The Madelung formalism is discussed in detail in \cite{Suarez:2011yf, Suarez:2015uva, Johnston:2009wz, Kopp:2017hbb}.
 Because this hydrodynamical formulation defines the fluid velocity as the gradient of the phase of the field $\psi$, problems arise when $\psi=0$, where the phase is not well defined.
-Because of this issue, the Madelung and Schr{\"o}dinger representations are not strictly equivalent unless a quantisation condition is imposed, as discussed in \cite{Wallstrom1994}.
-We do not consider the subtleties of the Madelung representation here, as it is sufficient for our purposes to consider the fluid velocity in the region of a solitonic core, where no field nodes are present.
-\footnote{It should be noted that, restoring dimensionful units, the fluid velocity $\vec{v}$ is related to the ususal quantum mechanical probability current $\vec{j}$ through $\vert\psi\vert^2\vec{v} = \vec{j} = \hbar/2mi\left[\psi^*\nabla\psi-\psi\nabla\psi^*\right]$} For a discussion of the possible remedies to the `nodes problem', the reader is referred to Chapter 15.3 of \cite{Wyatt2005}.
+Because of this issue, the Madelung and Schr{\"o}dinger representations are not strictly equivalent unless a quantisation condition is imposed, as discussed in \cite{Wallstrom:1994fp}.
+We do not consider the subtleties of the Madelung representation here, as it is sufficient for our purposes to consider the fluid velocity in the region of a solitonic core, where no field nodes are present.\footnote{It should be noted that, restoring dimensionful units, the fluid velocity $\vec{v}$ is related to the ususal quantum mechanical probability current $\vec{j}$ through $\vert\psi\vert^2\vec{v} = \vec{j} = \hbar/2mi\left[\psi^*\nabla\psi-\psi\nabla\psi^*\right]$} For a discussion of the possible remedies to the `nodes problem', the reader is referred to Chapter 15.3 of \cite{Wyatt:2005uc}.
 Where the Madelung representation is well defined, i.e.
@@
 Therefore, it is the characteristic internal velocity field of a Riemann-S ellipsoid which we expect to arise during our simulation of a soliton orbiting a central mass.
-It is found in \cite{Du2018} that an initially spherical solitonic core without self-rotation will gradually spin up to form a tidally-locked ellipsoid with an irrotational internal velocity field when orbiting a host mass.
+It is found in \cite{Du:2018zrg} that an initially spherical solitonic core without self-rotation will gradually spin up to form a tidally-locked ellipsoid with an irrotational internal velocity field when orbiting a host mass.
 We reproduce this result using \PyUltraLight.
 Figure \ref{fig:riemann} shows the internal velocity field of a solitonic satellite after one complete revolution around a host mass.
-The soliton has become elongated along the radial line connecting it to the host, indicating that it is tidally locked, while the velocity field within the tidal radius is visibly irrotational and bears the qualitative trademarks of the Riemann-S ellipsoid as presented in Figure 2 of \cite{Daller2012}.
+The soliton has become elongated along the radial line connecting it to the host, indicating that it is tidally locked, while the velocity field within the tidal radius is visibly irrotational and bears the qualitative trademarks of the Riemann-S ellipsoid as presented in Figure 2 of \cite{RindlerDaller:2011kx}.
 
@@
 The first term is the gravitational potential energy of the Schr{\"o}dinger-Poisson system, $E_{GP}$.
-As discussed in \cite{Hui2016}, the second term may be decomposed into contributions which may be considered separately as kinetic and `quantum' energies, $E_K$ and $E_Q$.
+As discussed in \cite{Hui:2016ltb}, the second term may be decomposed into contributions which may be considered separately as kinetic and `quantum' energies, $E_K$ and $E_Q$.
 However, for our purposes it is sufficient to consider only their combined contribution.

### 2018-09-24 draft.tex
Responses to referee comments

@@
 \usepackage{url}
-\usepackage{aas_macros}
-\usepackage[justification=justified]{caption}
+%\usepackage{aas_macros}
 
@@
 
-\newcommand{\apj}{Astrophysical J.}
-\newcommand{\prd}{Phys. Rev. D}
-\newcommand{\mnras}{Mon. Notices Royal Astron. Soc}
 \newcommand{\PyUltraLight}{\textsc{PyUltraLight}\xspace}
-\newcommand{\re}[1]{\textcolor{blue}{[{\bf RE}: #1]}}
-\newcommand{\ek}[1]{\textcolor{red}{[{\bf EK}: #1]}}
-\newcommand{\sh}[1]{\textcolor{green}{[{\bf SH}: #1]}}
 
@@
 The coupled equations \ref{eq:schrodinger} and \ref{eq:poisson} together form the nonlinear Schr{\"o}dinger-Poisson system which describes the dynamics of ULDM in the non-relativistic regime.
-It is this system of equations which \PyUltraLight is designed to solve.
+While Equations \ref{eq:schrodinger} and \ref{eq:poisson} are valid for open boundary conditions, \PyUltraLight is designed to solve the Schr{\"o}dinger-Poisson system under periodic boundary conditions.
+In this case the correct form of equation \ref{eq:poisson} is
+\begin{equation}\label{eq:poisson_periodic}
+    \nabla^2\Phi=4\pi G m \big(\vert\psi\vert^2-\langle\vert\psi\vert^2\rangle\big).
+\end{equation}
 
@@
     \CMcal{T}&=\left(\frac{8\pi}{3 H_0^2\Omega_{m_0}}\right)^{\frac{1}{2}}\approx75.5 \operatorname{Gyr},\label{eq:time}\\
-    \CMcal{M}&=\frac{1}{G}\left(\frac{8\pi}{3 H_0^2\Omega_{m_0}}\right)^{-\frac{1}{4}}\left(\frac{\hbar}{m}\right)^{\frac{3}{2}}\approx 7\times 10^7\left(\frac{10^{-23}\operatorname{eV}}{m}\right)\operatorname{M}_{\odot},\label{eq:mass}
+    \CMcal{M}&=\frac{1}{G}\left(\frac{8\pi}{3 H_0^2\Omega_{m_0}}\right)^{-\frac{1}{4}}\left(\frac{\hbar}{m}\right)^{\frac{3}{2}}\approx 7\times 10^7\left(\frac{10^{-23}\operatorname{eV}}{m}\right)^{\frac{3}{2}}\operatorname{M}_{\odot},\label{eq:mass}
 \end{align}
@@
 \end{equation}
-Dropping the primes for notational convenience, we see that the coupled differential equations of the  Schr{\"o}dinger-Poisson system  reduce to 
+Dropping the primes for notational convenience, we see that the coupled differential equations of the  Schr{\"o}dinger-Poisson system under periodic boundary conditions reduce to 
 \begin{align}
     i\Dot{\psi}(\vec{x},t)&=-\frac{1}{2}\nabla^2\psi(\vec{x},t)+\Phi(\vec{x},t)\psi(\vec{x},t),\label{eq:s-adim}\\
-    \nabla^2\Phi(\vec{x},t)&=4\pi\vert\psi(\vec{x},t)\vert^2,\label{eq:p-adim}
+    \nabla^2\Phi(\vec{x},t)&=4\pi\big(\vert\psi(\vec{x},t)\vert^2-\langle\vert\psi(\vec{x},t)\vert^2\rangle\big),\label{eq:p-adim}
 \end{align}
@@
 By dimensional analysis, we can easily restore dimensionful units to any of the quantities calculated by the code. In particular, in the following sections it is to be understood that
-\begin{equation}\label{eq:density}
+\begin{equation}
     E=\CMcal{M}\CMcal{L}^2\CMcal{T}^{-2}\ E_{code}, \quad v=\CMcal{L}\CMcal{T}^{-1}\ v_{code}, 
@@
 %
-where $\psi(\vec{x},t_{i})$ is the field configuration at this halfway point in the full timestep.
-The final operation in equation \ref{eq:schematic-psi}  only changes the phase of  $\psi$, so we could replace $\psi(\vec{x},t_{i})$ with $\psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)$ in equation \ref{eq:schematic-phi} with no change in meaning.
+where $\psi(\vec{x},t_{i})$ is the field configuration at this halfway point in the full timestep.\footnote{Unlike Equation \ref{eq:poisson_periodic}, in Equation \ref{eq:schematic-phi} the global average value of the density is not subtracted from the local value.
+This is because subtraction of a constant affects only the $k=0$ Fourier mode, which is manually set to zero within the code prior to the final inverse Fourier transform.
+Therefore, the subtraction of the average density has no effect on the final result, and is consequently not incorporated into the algorithm.} The final operation in equation \ref{eq:schematic-psi}  only changes the phase of  $\psi$, so we could replace $\psi(\vec{x},t_{i})$ with $\psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)$ in equation \ref{eq:schematic-phi} with no change in meaning.
 \PyUltraLight makes an additional simplification to the symmetrised split-step Fourier method by combining the consecutive half-steps in the nonlinear term into a single full step.
@@
 \begin{equation}
-    \psi(t\hspace*{-.2em}+\hspace*{-.2em}nh)=\operatorname{exp}\left[+\frac{ih}{2}\Phi\right]\left(\prod^n\operatorname{exp}\left[-\frac{ih}{2}\Phi\right]\operatorname{exp}\left[\frac{ih}{2}\nabla^2\right]\right)\operatorname{exp}\left[-\frac{ih}{2}\Phi\right]\psi(t),
+    \psi(t\hspace*{-.2em}+\hspace*{-.2em}nh)=\operatorname{exp}\left[+\frac{ih}{2}\Phi\right]\left(\prod^n\operatorname{exp}\left[-ih\Phi\right]\operatorname{exp}\left[\frac{ih}{2}\nabla^2\right]\right)\operatorname{exp}\left[-\frac{ih}{2}\Phi\right]\psi(t),
 \end{equation}
@@
 
-\subsection{Timestep Constraints}
+\subsection{Choosing the Timestep}
 
-In many finite-differencing schemes for  partial differential equations the Courant-Friedrichs-Lewy (CFL) condition is a necessary condition for convergence \cite{Ajaib:2013oua}, and puts an upper bound on the timestep as a function of the gridsize.
-The physical origin of the CFL condition is the finite speed of propagating information  in the evolving system.
+When employing the method of finite differences to numerically solve systems of partial differential equations, the Courant-Friedrichs-Lewy (CFL) condition is a necessary condition for convergence \cite{Ajaib:2013oua}.
+This condition puts an upper bound on the size of the timestep as a function of the grid spacing.
 Consequently, this condition is often cited in numerical analysis of ULDM physics governed by the Schr{\"o}dinger-Poisson system, see e.g.
 Ref.~\cite{Schwabe:2016rze}.
-Strictly speaking, however, the CFL condition is applicable to codes using finite differencing methods for spatial derivatives, and is thus not applicable to pseudospectral codes like \PyUltraLight.
-We also note that the Poisson equation is acausal in that it responds globally and instantaneously to local changes in the matter distribution.
-However, the Schr{\"o}dinger-Poisson equation is an approximation derived  in the nonrelativistic limit so this is not a shortcoming in regimes where the underlying equations are physically valid.
 
-In place of the CFL condition, the \PyUltraLight timestep can be set via the fluid interpretation of the Schr{\"o}dinger-Poisson system, discussed in greater detail in Section \ref{sec:disruption}.
-In dimensionless code units we define
+However, because \PyUltraLight uses pseudospectral methods to compute spatial derivatives, the algorithm is unconditionally stable \cite{Taha:1984jz} and timesteps need not be subjected to a CFL-like constraint.
+Consequently, the choice of timestep is in principle arbitrary, and is generally selected in accordance with accuracy requirements (our split-step algorithm is accurate to second order in the timestep).
+
+While the user is encouraged validate their choice of timestep through convergence testing on a case-by-case basis, \PyUltraLight does include a default timestep value which is chosen based on maximum fluid velocity arguments.
+In the fluid interpretation of the Schr{\"o}dinger-Poisson system, we define
 \begin{equation}\label{eq:madelung}
@@
 \end{equation}
-We define a `maximum velocity' for the simulation grid at which point the fluid appears to move backward because the phase change between successive grid points exceeds $\pi$.
-An upper bound can then be placed on the timestep so that fluid travelling at this maximum  velocity traverse a distance of one grid space per timestep.
-In practice this can yield a very demanding constraint on the timestep and for many scenarios the actual velocity is far less than this hypothetical limit.
-Consequently, the timestep can be  significantly increased at the user's discretion by changing the `step\textunderscore factor' parameter in the preliminary section of the Jupyter notebook.
-When this is done, convergence testing on a case-by-case basis is strongly encouraged.
+This then allows us to define a maximum fluid velocity that can be successfully modelled on a simulation grid of a given resolution.
+If the phase change between two consecutive grid points exceeds $\pi$, the fluid will appear to move backwards across the grid.
+Hence, we define the maximum velocity as that corresponding to a phase change of $\pi$ between grid points.
+We then set the default timestep, $\Delta t$, such that fluid travelling at this maximum velocity will traverse one grid space, $\Delta x$, per timestep.
+That is to say
+\begin{equation}\label{eq:timestep}
+    \Delta t = \frac{(\Delta x)^2}{\pi}.
+\end{equation}
+We stress that this is merely a choice of timestep, rather than a strict constraint.
+The motivation for this particular choice of timestep is that if the user finds that their fluid is approaching velocities at which this timestep is no longer sufficiently small for convergence, they will also be approaching the point at which they need to refine the simulation grid.
+Therefore, choosing the timestep to be smaller than the above default value offers little advantage.
+In many cases, this choice of timestep will be overly strict, and the user is free to choose a larger timestep by modifying the 'step\textunderscore factor' parameter in the Jupyter notebook.
+
+An alternative method of choosing the timestep, employed for example in \cite{Mocz:2017wlg}, is to ensure that neither of the unitary operators in Equation \ref{eq:schematic-psi} can lead to a phase change of more than $\pi$ for a single grid point over one timestep.
+However, because the pseudo-spectral algorithm does not compare the phase of a single gridpoint at different points in time, this choice of timestep is also not a requirement for stability.
+Using this method of choosing the timestep, we find the following set of constraints:
+\begin{equation}
+    \Delta t < \left[\frac{2\pi}{\Phi_{max}}, \frac{2(\Delta x)^2}{\pi}\right],
+\end{equation}
+where the second of these constraints is generally the stricter of the two, and is equivalent to our previous choice of timestep up to a multiplicative factor of $\mathcal{O}(1)$.
+
+Convergence testing has demonstrated that choosing the timestep according to Equation \ref{eq:timestep} is suitable for the majority of simulation scenarios.
+This is demonstrated in section \ref{sec:resolution}, where we contrast the performance of \PyUltraLight for differing choices of timestep.
 
@@
 
+\section{Convergence Testing Over Spatial and Temporal Resolution Ranges}\label{sec:resolution}
+
+Having demonstrated improving energy conservation with respect to spatial resolution, we now examine the convergence of the $\psi$ field configuration as spatial resolution and timestep are augmented.
+To perform this analysis, we initialise a simulation of two solitons diametrically opposed about a large Newtonian central potential.
+We run the simulation forward until the solitons become significantly tidally disrupted in their orbits, as shown in Figure \ref{fig:disruption}.
+
+\begin{figure}
+  \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{first_and_last}
+  \caption{TBW}
+  \label{fig:disruption}
+\end{figure}
+
+------------------------------
+%Include figure with first and last configuration of the simulation, caption with parameters used
+
+%Mention that tim,estep 1 not much better than timestep 0.1 look at phase
+%Phase appears to be more sensitive than density to resolution
+-----------------------------
+
+To examine the sensitivity of the final $\psi$ field configuration to changes in spatial resolution, we first run this simulation at $256^3$, with the default timestep.
+We then re-run the simulation at resolutions ranging from $64^3$ to $320^3$, fixing the timestep in each case to be the same as the $256^3$ run.
+The final output arrays are then down-sampled such that we obtain five arrays of dimension $64^3$ for comparison.
+We arrange the array values in order of increasing density, and then analyse the deviation of the phase and the magnitude of $\psi$ from the values of the highest resolution run.
+The results of this process are shown in Figures \ref{fig:spatial_magnitudes} and \ref{fig:spatial_phases}.
+
+\begin{figure}
+  \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{spatial_discretisation}
+  \caption{Improving convergence of $\vert\psi\vert$ with increased spatial resolution for the simulation shown in Figure \ref{fig:disruption}.
+Field values are arranged in order of increasing magnitude of the highest resolution run from left to right.}
+  \label{fig:spatial_magnitudes}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{phases}
+  \caption{Deviation of the phase of $\psi$ compared to the highest resolution result ($320^3$).
+Field values are arranged in order of increasing magnitude from left to right.
+A slight improvement in phase convergence can be seen for higher density regions to the right.}
+  \label{fig:spatial_phases}
+\end{figure}
+
+To examine the sensitivity of the final $\psi$ field configuration to the choice of the timestep, we run the simulation at $256^3$, choosing the timestep to be 0.1, 1, 10, and 50 times the default timestep.
+Again we down-sample the output arrays to $64^3$, and arrange the array values in order of $\psi$ field magnitude.
+We then examine how the phase and magnitude of the field values differ in comparison to the case of 0.1 $\times$ the default timestep.
+The results of this analysis are shown in Figures \ref{fig:temporal_magnitudes} and \ref{fig:temporal_phases}.
+
+\begin{figure}
+  \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{temporal_discretisation}
+  \caption{TBW}
+  \label{fig:temporal_magnitudes}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{temporal_phases}
+  \caption{TBW}
+  \label{fig:temporal_phases}
+\end{figure}
+
 \section{Discussion and Outlook}

### 2018-09-24 draft.tex
Second lot of changes in response to ref's comments

@@
 
-Having demonstrated improving energy conservation with respect to spatial resolution, we now examine the convergence of the $\psi$ field configuration as spatial resolution and timestep are augmented.
-To perform this analysis, we initialise a simulation of two solitons diametrically opposed about a large Newtonian central potential.
+Having demonstrated improving energy conservation with respect to spatial resolution, we now examine the overall convergence of the $\psi$ field configuration as the spatial resolution and timestep are augmented in a typical simulation.
+We initialise a simulation of two solitons diametrically opposed about a large Newtonian central potential.
 We run the simulation forward until the solitons become significantly tidally disrupted in their orbits, as shown in Figure \ref{fig:disruption}.
@@
   \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{first_and_last}
-  \caption{TBW}
+  \caption{Visualisation of the configuration used to test $\psi$ convergence under changes to spatial and temporal resolution.
+In code units, two solitons of mass $m=20$ are initialised at radial distances $r=2$ from a central mass $M=1000$.
+The solitons are given initial velocity of $\vert v\vert=\sqrt{M/r}$ in opposing directions, corresponding to a clockwise orbit around the central mass.
+In code units, the total box size is 10, while the duration of the simulation is 0.25.
+Time runs from left to right in the figure.}
   \label{fig:disruption}
 \end{figure}
-
-------------------------------
-%Include figure with first and last configuration of the simulation, caption with parameters used
-
-%Mention that tim,estep 1 not much better than timestep 0.1 look at phase
-%Phase appears to be more sensitive than density to resolution
------------------------------
 
@@
 The results of this process are shown in Figures \ref{fig:spatial_magnitudes} and \ref{fig:spatial_phases}.
+As expected, the lowest resolution run demonstrates poor convergence in both the magnitude and phase of $\psi$, with convergence increasing to acceptable levels at $256^3$.
 
 \begin{figure}
-  \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{spatial_discretisation}
+  \includegraphics[width=1.\textwidth,trim=0 1cm 0 1.2cm,clip]{spatial_discretisation}
   \caption{Improving convergence of $\vert\psi\vert$ with increased spatial resolution for the simulation shown in Figure \ref{fig:disruption}.
@@
 \begin{figure}
-  \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{phases}
+  \includegraphics[width=1.\textwidth,trim=0 0.5cm 0 1.2cm,clip]{phases}
   \caption{Deviation of the phase of $\psi$ compared to the highest resolution result ($320^3$).
@@
 To examine the sensitivity of the final $\psi$ field configuration to the choice of the timestep, we run the simulation at $256^3$, choosing the timestep to be 0.1, 1, 10, and 50 times the default timestep.
-Again we down-sample the output arrays to $64^3$, and arrange the array values in order of $\psi$ field magnitude.
+Again we down-sample the output arrays to $64^3$, and arrange the array values in order of the $\psi$ field magnitude (as determined by the smallest timestep run).
 We then examine how the phase and magnitude of the field values differ in comparison to the case of 0.1 $\times$ the default timestep.
 The results of this analysis are shown in Figures \ref{fig:temporal_magnitudes} and \ref{fig:temporal_phases}.
+We notice that the deviation of the results of the default timestep from those of the timestep which is 10 times smaller are negligible, indicating that our choice of default timestep is suitably small to ensure convergence.
+The results remain reasonably accurate as the timestep is increased within $~10$ times the default value.
+At $50$ times the default value, both phase and magnitude data become significantly less accurate.
 
 \begin{figure}
-  \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{temporal_discretisation}
-  \caption{TBW}
+  \includegraphics[width=1.\textwidth,trim=0 1cm 0 1cm,clip]{temporal_discretisation}
+  \caption{Deviation of the magnitude of $\psi$ as compared to the simulation run with 0.1 times the default timestep.
+The convergence of the results for the default timestep with the results from the smaller timestep run demonstrate the suitability of the chosen default.
+We see relatively good convergence, particularly in high density regions, for a $<10$-fold increase above the default timestep.}
   \label{fig:temporal_magnitudes}
@@
 \begin{figure}
-  \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{temporal_phases}
-  \caption{TBW}
+  \includegraphics[width=1.\textwidth,trim=0 0.5cm 0 1cm,clip]{temporal_phases}
+  \caption{Deviation of the phase of $\psi$ from the results of the simulation in which the timestep is set to 0.1 times the default.
+Again we see excellent convergence for the default timestep, and reasonable convergence for $<10$ times the default, with accuracy improving at high density regions.}
   \label{fig:temporal_phases}
@@
 In particular, future releases may incorporate a variable timestep and more sophisticated physics, including explicit self-interactions in the axion sector or additional matter components.
+Augmented versions of the code may also include higher-order generalisations of the pseudo-spectral method, such as those used in \cite{Levkov:2018kau}.
 \PyUltraLight is publicly available under a BSD license.

### 2018-09-24 draft.tex
Shaun's rewrite of my typo

@@
 
-While the user is encouraged validate their choice of timestep through convergence testing on a case-by-case basis, \PyUltraLight does include a default timestep value which is chosen based on maximum fluid velocity arguments.
+While the user is encouraged to choose their own timestep and validate it through convergence testing on a case-by-case basis, \PyUltraLight does include a default timestep value which is chosen based on maximum fluid velocity arguments.
 In the fluid interpretation of the Schr{\"o}dinger-Poisson system, we define

### 2018-09-26 draft.tex
richard's read through'

--- 2018-09-24 draft.tex
@@
 \usepackage{url}
+\newcommand{\re}[1]{\textcolor{blue}{[{\bf RE}: #1]}}
+
 %\usepackage{aas_macros}
@@
 We then describe their implementation in \PyUltraLight, before moving on to describe testing and verification procedures applied to the code.
-We reproduce a selection of results from a variety of recent ULDM simulations and discuss energy conservation in detail.
+We reproduce a selection of results from a variety of recent ULDM simulations and discuss the energy conservation and accuracy as a function of spatial resolution.
 
@@
 \end{equation}
+\re{do we define $\langle \rangle$? Also -- should we explain in a sentence WHY the form of the equation changes when we go to period b.c.s?}
 
@@
 %
-where $\psi(\vec{x},t_{i})$ is the field configuration at this halfway point in the full timestep.\footnote{Unlike Equation \ref{eq:poisson_periodic}, in Equation \ref{eq:schematic-phi} the global average value of the density is not subtracted from the local value.
-This is because subtraction of a constant affects only the $k=0$ Fourier mode, which is manually set to zero within the code prior to the final inverse Fourier transform.
-Therefore, the subtraction of the average density has no effect on the final result, and is consequently not incorporated into the algorithm.} The final operation in equation \ref{eq:schematic-psi}  only changes the phase of  $\psi$, so we could replace $\psi(\vec{x},t_{i})$ with $\psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)$ in equation \ref{eq:schematic-phi} with no change in meaning.
+where $\psi(\vec{x},t_{i})$ is the field configuration at this halfway point in the full timestep.
+\re{We explicitly set the $k=0$ Fourier mode to zero prior to the final inverse Fourier transform; as a consequence there is no need to subtract the global average density from the local value in Equation \ref{eq:schematic-phi}, in contrast to Equation \ref{eq:poisson_periodic}.} The final operation in equation \ref{eq:schematic-psi}  only changes the phase of  $\psi$, so we could replace $\psi(\vec{x},t_{i})$ with $\psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)$ in equation \ref{eq:schematic-phi} with no change in meaning.
 \PyUltraLight makes an additional simplification to the symmetrised split-step Fourier method by combining the consecutive half-steps in the nonlinear term into a single full step.
@@
 
-When employing the method of finite differences to numerically solve systems of partial differential equations, the Courant-Friedrichs-Lewy (CFL) condition is a necessary condition for convergence \cite{Ajaib:2013oua}.
-This condition puts an upper bound on the size of the timestep as a function of the grid spacing.
-Consequently, this condition is often cited in numerical analysis of ULDM physics governed by the Schr{\"o}dinger-Poisson system, see e.g.
+The Courant-Friedrichs-Lewy (CFL) condition is an upper bound on the timestep (as a function of grid-spacing) that must be satisfied by many partial differential equation solvers based on finite-differencing  \cite{Ajaib:2013oua} and is often cited in numerical analyses of ULDM via the Schr{\"o}dinger-Poisson system, see e.g.
 Ref.~\cite{Schwabe:2016rze}.
+However, the CFL condition expresses a causality constraint, and is generally only strictly applicable to hyperbolic PDEs, whereas the Schr{\"o}dinger-Poisson has only a first order time derivative, even though it is effectively the nonrelativistic limit of the Klein-Gordon equation.
+Moreover, because \PyUltraLight computes spatial derivatives via a pseudospectral method, technically it is unconditionally stable \cite{Taha:1984jz}.
+Our split-step algorithm is  second order in the timestep, and its value will always be an empirical tradeoff between computational cost and convergence to the apparent limit in which step is arbitrarily small.
+Consequently, the user is encouraged to validate their choice  of timestep via case-by-case  convergence testing.
 
-However, because \PyUltraLight uses pseudospectral methods to compute spatial derivatives, the algorithm is unconditionally stable \cite{Taha:1984jz} and timesteps need not be subjected to a CFL-like constraint.
-Consequently, the choice of timestep is in principle arbitrary, and is generally selected in accordance with accuracy requirements (our split-step algorithm is accurate to second order in the timestep).
-
-While the user is encouraged validate their choice of timestep through convergence testing on a case-by-case basis, \PyUltraLight does include a default timestep value which is chosen based on maximum fluid velocity arguments.
-In the fluid interpretation of the Schr{\"o}dinger-Poisson system, we define
+The default timestep in  \PyUltraLight  is fixed with reference to the fluid interpretation of the Schr{\"o}dinger-Poisson system \re{CITATION}. We obtain this via
+%
 \begin{equation}\label{eq:madelung}
-    \psi\equiv\sqrt{\rho}e^{i\theta}, \quad \vec{v}\equiv\boldsymbol{\nabla}\theta.
+    \psi\equiv\sqrt{\rho}e^{i\theta}, \quad \vec{v}\equiv\boldsymbol{\nabla}\theta
 \end{equation}
-This then allows us to define a maximum fluid velocity that can be successfully modelled on a simulation grid of a given resolution.
-If the phase change between two consecutive grid points exceeds $\pi$, the fluid will appear to move backwards across the grid.
-Hence, we define the maximum velocity as that corresponding to a phase change of $\pi$ between grid points.
-We then set the default timestep, $\Delta t$, such that fluid travelling at this maximum velocity will traverse one grid space, $\Delta x$, per timestep.
-That is to say
+%
+\re{Do we need a bit more detail, or a reference?}  If the phase change between two adjacent grid points across a timestep, exceeds $\pi$ the fluid will appear to move backwards across the grid, and we set the default timestep, $\Delta t$, such that fluid travelling at this maximum velocity will traverse one grid space, $\Delta x$, per timestep, or
 \begin{equation}\label{eq:timestep}
@@
 \end{equation}
-We stress that this is merely a choice of timestep, rather than a strict constraint.
-The motivation for this particular choice of timestep is that if the user finds that their fluid is approaching velocities at which this timestep is no longer sufficiently small for convergence, they will also be approaching the point at which they need to refine the simulation grid.
-Therefore, choosing the timestep to be smaller than the above default value offers little advantage.
-In many cases, this choice of timestep will be overly strict, and the user is free to choose a larger timestep by modifying the 'step\textunderscore factor' parameter in the Jupyter notebook.
+%
+This is  a choice, rather than a strict constraint but if the ``fluid'' approaches velocities where the phase appears to switch direction, the configuration will also be approaching the point where the simulation grid is too coarse to fully resolve the dynamics.
+Making the timestep much smaller than this value may offer little practical advantage.
+Conversely, in some cases the breakdown may occur in regions of the simulation volume of little physical interest, and the user is free to choose a larger timestep  via the 'step\textunderscore factor' parameter in the Jupyter notebook.
 
-An alternative method of choosing the timestep, employed for example in \cite{Mocz:2017wlg}, is to ensure that neither of the unitary operators in Equation \ref{eq:schematic-psi} can lead to a phase change of more than $\pi$ for a single grid point over one timestep.
-However, because the pseudo-spectral algorithm does not compare the phase of a single gridpoint at different points in time, this choice of timestep is also not a requirement for stability.
-Using this method of choosing the timestep, we find the following set of constraints:
+Alternatively, Ref.~\cite{Mocz:2017wlg} fixes the timestep by ensuring that neither of the unitary operators in Equation \ref{eq:schematic-psi}  leads to a phase change of more than $\pi$ for a single grid point over one timestep.
+However, because the pseudo-spectral algorithm does not compare the phase of a single gridpoint at different points in time, this choice of timestep is  not a requirement for stability.
+This method  give the following constraints:
 \begin{equation}
@@
 \end{equation}
-where the second of these constraints is generally the stricter of the two, and is equivalent to our previous choice of timestep up to a multiplicative factor of $\mathcal{O}(1)$.
-
-Convergence testing has demonstrated that choosing the timestep according to Equation \ref{eq:timestep} is suitable for the majority of simulation scenarios.
-This is demonstrated in section \ref{sec:resolution}, where we contrast the performance of \PyUltraLight for differing choices of timestep.
+%
+where the second of these constraints is generally the stricter of the two, and is equivalent to the default choice up to a multiplicative factor of $\mathcal{O}(1)$.
+Our experience is that specifying the timestep via Equation \ref{eq:timestep} is suitable for the majority of simulation scenarios, and we explore convergence in more detail in Section \ref{sec:resolution}.
 
@@
 An example of this is given in \cite{Veltmaat:2016rxo}, where it is demonstrated that the coarse-grained nature of the particle-mesh method renders the algorithm incapable of reproducing detailed interference patterns such as those shown here.
-%\re{read to here}
 
@@
 
+\section{Convergence  and Validation}
 \subsection{Energy Conservation}\label{sec:energy}
@@
 
-Figure \ref{fig:energy_change_2} demonstrates the improvement in energy conservation with increasing grid resolution for the case of a single soliton becoming tidally disrupted in a Newtonian central potential.
-The simulation setup is the same as that used to generate Figure \ref{fig:combined_1}.
+Figure \ref{fig:energy_change_2} demonstrates the improvement in energy conservation with increasing grid resolution for a single soliton tidally disrupted in a Newtonian central potential, with the same set up as used in Figure \ref{fig:combined_1}.
 Namely, a single soliton of mass 12 code units is initialised at a distance of 3 code units from a central mass, $M=1000$.
@@
 This highlights the importance of a multifaceted approach to convergence testing.
-At $256^3$, energy is conserved to $10^{-4}\ \%$.
+At $256^3$, energy is conserved to parts in $10^{-6}$.
 
-\section{Convergence Testing Over Spatial and Temporal Resolution Ranges}\label{sec:resolution}
+\subsection{Spatial and Temporal Resolution}\label{sec:resolution}
 
-Having demonstrated improving energy conservation with respect to spatial resolution, we now examine the overall convergence of the $\psi$ field configuration as the spatial resolution and timestep are augmented in a typical simulation.
-We initialise a simulation of two solitons diametrically opposed about a large Newtonian central potential.
-We run the simulation forward until the solitons become significantly tidally disrupted in their orbits, as shown in Figure \ref{fig:disruption}.
+We now examine the  convergence of the $\psi$ field configuration as a function of spatial resolution and timestep in a typical simulation.
+We initialise \PyUltraLight with two  diametrically opposed solitons orbiting a large Newtonian central potential, running until the solitons are tidally disrupted, as shown in Figure \ref{fig:disruption}.
 
@@
   \includegraphics[width=1.\textwidth,trim=0 0 0 0,clip]{first_and_last}
-  \caption{Visualisation of the configuration used to test $\psi$ convergence under changes to spatial and temporal resolution.
-In code units, two solitons of mass $m=20$ are initialised at radial distances $r=2$ from a central mass $M=1000$.
-The solitons are given initial velocity of $\vert v\vert=\sqrt{M/r}$ in opposing directions, corresponding to a clockwise orbit around the central mass.
-In code units, the total box size is 10, while the duration of the simulation is 0.25.
-Time runs from left to right in the figure.}
+  \caption{The configuration used to test the sensitivity of solutions to the spatial and temporal resolution.
+Two solitons of mass $m=20$ are initialised at radial distances $r=2$ from a central mass with $M=1000$ moving in opposite directions with initial speeds $\vert v\vert=\sqrt{M/r}$, corresponding to clockwise orbits around the central mass.
+The  box size is 10, while the total duration is 0.25 (all quantities in code units.
+Time runs from left to right.
+\re{you could actually zoom in, to only include the active region.} }
   \label{fig:disruption}
@@
 
-To examine the sensitivity of the final $\psi$ field configuration to changes in spatial resolution, we first run this simulation at $256^3$, with the default timestep.
-We then re-run the simulation at resolutions ranging from $64^3$ to $320^3$, fixing the timestep in each case to be the same as the $256^3$ run.
-The final output arrays are then down-sampled such that we obtain five arrays of dimension $64^3$ for comparison.
-We arrange the array values in order of increasing density, and then analyse the deviation of the phase and the magnitude of $\psi$ from the values of the highest resolution run.
-The results of this process are shown in Figures \ref{fig:spatial_magnitudes} and \ref{fig:spatial_phases}.
-As expected, the lowest resolution run demonstrates poor convergence in both the magnitude and phase of $\psi$, with convergence increasing to acceptable levels at $256^3$.
+To examine the sensitivity of the  $\psi$ field configuration to the spatial resolution, we first run at $256^3$ with the default timestep.
+We then re-run  at resolutions  from $64^3$ to $320^3$ with the timestep fixed to the  $256^3$ value and downsample the final outputs to $64^3$.
+We  sort the resulting values by  the density at the corresponding spatial location, and plot differences in the phase and the magnitude of $\psi$ relative to the values of the $320^3$ run in Figures \ref{fig:spatial_magnitudes} and \ref{fig:spatial_phases}.
+\re{maybe this could be one combined figure with both plots.}  The convergence is poor  at $\psi$ at $64^3$, but improves with resolution, to the point that there is little difference between the  $256^3$ and $320^3$ cases.
 
@@
 
-To examine the sensitivity of the final $\psi$ field configuration to the choice of the timestep, we run the simulation at $256^3$, choosing the timestep to be 0.1, 1, 10, and 50 times the default timestep.
-Again we down-sample the output arrays to $64^3$, and arrange the array values in order of the $\psi$ field magnitude (as determined by the smallest timestep run).
-We then examine how the phase and magnitude of the field values differ in comparison to the case of 0.1 $\times$ the default timestep.
-The results of this analysis are shown in Figures \ref{fig:temporal_magnitudes} and \ref{fig:temporal_phases}.
-We notice that the deviation of the results of the default timestep from those of the timestep which is 10 times smaller are negligible, indicating that our choice of default timestep is suitably small to ensure convergence.
-The results remain reasonably accurate as the timestep is increased within $~10$ times the default value.
-At $50$ times the default value, both phase and magnitude data become significantly less accurate.
+To examine the sensitivity of the the  $\psi$ field configuration to the timestep, we take the same default simulation at $256^3$, and then compare this to runs with timesteps 0.1, 10, and 50 times the default and down-sample the final output arrays to $64^3$.
+We sort sort the array values in order of the $\psi$ field magnitude in the run with the smallest timestep and in Figures \ref{fig:temporal_magnitudes} and \ref{fig:temporal_phases} \re{maybe also combined into one full page figure?} we show the difference in the phase and magnitude of $\psi$ as a function of the timestep.
+The difference between the results with the default timestep and a value 10 times smaller are negligible; and there is reasonable agreement between the default case and those with the timestep boosted by a factor of 10.
+However, when the timestep is increased by a factor of $50$ the accuracy of both the phase and magnitude data are significantly reduced.
 

### 2018-09-29 draft.tex
Responses to RE's comments

@@
 
-\PyUltraLight is based on a symmetrised-split-step (leapfrog) solver for the time evolution, and uses a pseudospectral Fourier algorithm to solve the Poisson equation for the gravitational potential at each step.\footnote{A similar methodology was described in Ref.~\cite{Paredes:2015wga}; to our knowledge this code has not been released.
+\PyUltraLight is based on a symmetrised-split-step (leapfrog) solver for the time evolution, and uses a pseudospectral Fourier algorithm to solve the Poisson equation for the gravitational potential at each step.\footnote{A similar methodology was described in Ref.~\cite{Paredes:2015wga}; at the time of writing this code has not been released.
 Spectral methods are often used to solve the Poisson equation in large scale structure simulations, while the {\sc PSpectre} code \cite{Easther:2010qz} provides a pseudospectral solver for the evolution of the inflaton  and fields coupled to it during parametric resonance and preheating after inflation \cite{Amin:2010dc,Amin:2011hj,Zhou:2013tsa}.} This algorithm has $2^{nd}$ order accurate time integration steps and sub-percent level energy conservation, while the wavefunction normalisation  is conserved to machine precision.
@@
 \begin{equation}\label{eq:poisson_periodic}
-    \nabla^2\Phi=4\pi G m \big(\vert\psi\vert^2-\langle\vert\psi\vert^2\rangle\big).
+    \nabla^2\Phi=4\pi G m \big(\vert\psi\vert^2-\langle\vert\psi\vert^2\rangle\big),
 \end{equation}
-\re{do we define $\langle \rangle$? Also -- should we explain in a sentence WHY the form of the equation changes when we go to period b.c.s?}
+where we subtract the average density from the right hand side of the Poisson equation.
+The form of Equation \ref{eq:poisson_periodic} is a consequence of Gauss' law and the fact that the surface integral of the gradient of the field is identically zero when periodic boundary conditions are imposed.
+
+\re{do we define $\langle \rangle$?
+Also -- should we explain in a sentence WHY the form of the equation changes when we go to period b.c.s?}\ek{I've given slightly more info here, think this is sufficient to understand the form of 2.8}
 
@@
 where $\psi(\vec{x},t_{i})$ is the field configuration at this halfway point in the full timestep.
-\re{We explicitly set the $k=0$ Fourier mode to zero prior to the final inverse Fourier transform; as a consequence there is no need to subtract the global average density from the local value in Equation \ref{eq:schematic-phi}, in contrast to Equation \ref{eq:poisson_periodic}.} The final operation in equation \ref{eq:schematic-psi}  only changes the phase of  $\psi$, so we could replace $\psi(\vec{x},t_{i})$ with $\psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)$ in equation \ref{eq:schematic-phi} with no change in meaning.
+We explicitly set the $k=0$ Fourier mode to zero prior to the final inverse Fourier transform; as a consequence there is no need to subtract the global average density from the local value in Equation \ref{eq:schematic-phi}, in contrast to Equation \ref{eq:poisson_periodic}.\ek{Incorporated this explanation into main text} The final operation in equation \ref{eq:schematic-psi}  only changes the phase of  $\psi$, so we could replace $\psi(\vec{x},t_{i})$ with $\psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)$ in equation \ref{eq:schematic-phi} with no change in meaning.
 \PyUltraLight makes an additional simplification to the symmetrised split-step Fourier method by combining the consecutive half-steps in the nonlinear term into a single full step.
@@
 
-The default timestep in  \PyUltraLight  is fixed with reference to the fluid interpretation of the Schr{\"o}dinger-Poisson system \re{CITATION}. We obtain this via
-%
+The default timestep in  \PyUltraLight  is fixed with reference to the fluid interpretation of the Schr{\"o}dinger-Poisson system \cite{Hui:2016ltb}.
+The fluid interpretation is often used to recast the Schr{\"o}dinger-Poisson system in the form of the Madelung equations \cite{Suarez:2011yf}, such that the system can then be studied using the principles of hydrodynamics.\ek{Slightly more info and reference given here, though I don't want to go into too much detail as the specifics of the fluid representation aren't needed here.} The first step is to define:
 \begin{equation}\label{eq:madelung}
-    \psi\equiv\sqrt{\rho}e^{i\theta}, \quad \vec{v}\equiv\boldsymbol{\nabla}\theta
+    \psi\equiv\sqrt{\rho}e^{i\theta}, \quad \vec{v}\equiv\boldsymbol{\nabla}\theta.
 \end{equation}
-%
-\re{Do we need a bit more detail, or a reference?}  If the phase change between two adjacent grid points across a timestep, exceeds $\pi$ the fluid will appear to move backwards across the grid, and we set the default timestep, $\Delta t$, such that fluid travelling at this maximum velocity will traverse one grid space, $\Delta x$, per timestep, or
+From this definition of the fluid velocity, we see that if the phase change between two adjacent grid points exceeds $\pi$,\ek{not over a timestep, but at the same point in time} the fluid will appear to move backwards across the grid.
+Therefore, we set the default timestep, $\Delta t$, such that fluid travelling at this maximum velocity will traverse one grid space, $\Delta x$, per timestep, or
 \begin{equation}\label{eq:timestep}
@@
 \end{equation}
-%
-This is  a choice, rather than a strict constraint but if the ``fluid'' approaches velocities where the phase appears to switch direction, the configuration will also be approaching the point where the simulation grid is too coarse to fully resolve the dynamics.
-Making the timestep much smaller than this value may offer little practical advantage.
+This is a choice, rather than a strict constraint on the timestep.
+We motivate this choice of timestep by recognising that if the ``fluid'' approaches velocities where the phase appears to switch direction, the configuration will also be approaching the point where the simulation grid is too coarse to fully resolve the dynamics.
+Hence, making the timestep much smaller than this value may offer little practical advantage.
 Conversely, in some cases the breakdown may occur in regions of the simulation volume of little physical interest, and the user is free to choose a larger timestep  via the 'step\textunderscore factor' parameter in the Jupyter notebook.
 
-Alternatively, Ref.~\cite{Mocz:2017wlg} fixes the timestep by ensuring that neither of the unitary operators in Equation \ref{eq:schematic-psi}  leads to a phase change of more than $\pi$ for a single grid point over one timestep.
+Alternatively, Ref.~\cite{Mocz:2017wlg} fixes the timestep by ensuring that neither of the unitary operators in Equation \ref{eq:schematic-psi}  lead to a phase change of more than $\pi$ for a single grid point over one timestep.
 However, because the pseudo-spectral algorithm does not compare the phase of a single gridpoint at different points in time, this choice of timestep is not a requirement for stability.
-This method  give the following constraints:
+This method gives the following constraints:
 \begin{equation}
@@
 %
-where the second of these constraints is generally the stricter of the two, and is equivalent to the default choice up to a multiplicative factor of $\mathcal{O}(1)$.
+where the second of these constraints is generally the stricter of the two, and is equivalent to our default choice of timestep up to a multiplicative factor of $\mathcal{O}(1)$.
 Our experience is that specifying the timestep via Equation \ref{eq:timestep} is suitable for the majority of simulation scenarios, and we explore convergence in more detail in Section \ref{sec:resolution}.
@@
 Time runs from left to right.
-\re{you could actually zoom in, to only include the active region.} }
+\re{you could actually zoom in, to only include the active region.}\ek{I'm just concerned about misrepesenting the size of the simulation grid - I wanted to accurately represent the buffer zone, because insufficient buffer will affect the dynamics, and people trying to reproduce these results might get confused.
+In any case, the visual details of the density distribution aren't important here, the purpose of the figure is just to give an idea of the setup} }
   \label{fig:disruption}
@@
 We then re-run  at resolutions  from $64^3$ to $320^3$ with the timestep fixed to the  $256^3$ value and downsample the final outputs to $64^3$.
-We  sort the resulting values by  the density at the corresponding spatial location, and plot differences in the phase and the magnitude of $\psi$ relative to the values of the $320^3$ run in Figures \ref{fig:spatial_magnitudes} and \ref{fig:spatial_phases}.
-\re{maybe this could be one combined figure with both plots.}  The convergence is poor  at $\psi$ at $64^3$, but improves with resolution, to the point that there is little difference between the  $256^3$ and $320^3$ cases.
-
-\begin{figure}
-  \includegraphics[width=1.\textwidth,trim=0 1cm 0 1.2cm,clip]{spatial_discretisation}
-  \caption{Improving convergence of $\vert\psi\vert$ with increased spatial resolution for the simulation shown in Figure \ref{fig:disruption}.
-Field values are arranged in order of increasing magnitude of the highest resolution run from left to right.}
-  \label{fig:spatial_magnitudes}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=1.\textwidth,trim=0 0.5cm 0 1.2cm,clip]{phases}
-  \caption{Deviation of the phase of $\psi$ compared to the highest resolution result ($320^3$).
-Field values are arranged in order of increasing magnitude from left to right.
-A slight improvement in phase convergence can be seen for higher density regions to the right.}
-  \label{fig:spatial_phases}
-\end{figure}
+We  sort the resulting values by  the density at the corresponding spatial location, and plot differences in the phase and the magnitude of $\psi$ relative to the values of the $320^3$ run as shown in Figure \ref{fig:spatial} (bottom).
+\re{maybe this could be one combined figure with both plots.}\ek{Done.} The convergence is poor at $64^3$, but improves with resolution, to the point that there is little difference between the  $256^3$ and $320^3$ cases.
 
 To examine the sensitivity of the the  $\psi$ field configuration to the timestep, we take the same default simulation at $256^3$, and then compare this to runs with timesteps 0.1, 10, and 50 times the default and down-sample the final output arrays to $64^3$.
-We sort sort the array values in order of the $\psi$ field magnitude in the run with the smallest timestep and in Figures \ref{fig:temporal_magnitudes} and \ref{fig:temporal_phases} \re{maybe also combined into one full page figure?} we show the difference in the phase and magnitude of $\psi$ as a function of the timestep.
+We sort sort the array values in order of the $\psi$ field magnitude in the run with the smallest timestep and in Figure \ref{fig:temporal} \re{maybe also combined into one full page figure?}\ek{Done.} we show the difference in the phase and magnitude of $\psi$ as a function of the timestep.
 The difference between the results with the default timestep and a value 10 times smaller are negligible; and there is reasonable agreement between the default case and those with the timestep boosted by a factor of 10.
@@
 \begin{figure}
-  \includegraphics[width=1.\textwidth,trim=0 1cm 0 1cm,clip]{temporal_discretisation}
-  \caption{Deviation of the magnitude of $\psi$ as compared to the simulation run with 0.1 times the default timestep.
-The convergence of the results for the default timestep with the results from the smaller timestep run demonstrate the suitability of the chosen default.
-We see relatively good convergence, particularly in high density regions, for a $<10$-fold increase above the default timestep.}
-  \label{fig:temporal_magnitudes}
+  \includegraphics[width=1.\textwidth,trim=0 1cm 0 2cm,clip]{combined2}
+  \caption{Top: Deviation of the phase of $\psi$ compared to the highest resolution result ($320^3$).
+Field values are arranged in order of increasing magnitude from left to right.
+A slight improvement in phase convergence can be seen for higher density regions to the right.
+Bottom: Improving convergence of $\vert\psi\vert$ with increased spatial resolution for the simulation shown in Figure \ref{fig:disruption}}
+  \label{fig:spatial}
 \end{figure}
@@
 \begin{figure}
-  \includegraphics[width=1.\textwidth,trim=0 0.5cm 0 1cm,clip]{temporal_phases}
-  \caption{Deviation of the phase of $\psi$ from the results of the simulation in which the timestep is set to 0.1 times the default.
-Again we see excellent convergence for the default timestep, and reasonable convergence for $<10$ times the default, with accuracy improving at high density regions.}
-  \label{fig:temporal_phases}
+  \includegraphics[width=1.\textwidth,trim=0 1cm 0 2cm,clip]{combined}
+  \caption{Top: Deviation of the phase of $\psi$ from the results of the simulation in which the timestep is set to 0.1 times the default.
+We see excellent convergence for the default timestep, and reasonable convergence for $<10$ times the default, with accuracy improving at high density regions.
+Bottom: Deviation of the magnitude of $\psi$ as compared to the simulation run with 0.1 times the default timestep.
+The convergence of the results for the default timestep with the results from the smaller timestep run demonstrate the suitability of the chosen default.
+We see relatively good convergence, particularly in high density regions, for a $<10$-fold increase above the default timestep.}
+  \label{fig:temporal}
+\end{figure}
+
+\ek{added new content from here to the end of the section}
+To supplement our analysis of the dependence of the field configuration on spatial and temporal resolutions, we include Figures \ref{fig:comparison} and \ref{fig:comparison2}.
+These figures represent the density profiles down the axis of symmetry of the initial configuration (vertical axis in Figure \ref{fig:disruption}), however rather than the profiles at $t=0$, these are the profiles after approximately half a revolution around the central potential at t=0.28 code units.
+At this point, the two initial solitons have become distorted due to tidal forces, but have yet to become completely disrupted.
+Figure \ref{fig:comparison} demonstrates the effect of increasing the timestep from 0.1 times the default value to a maximum of 50 times the default value.
+We see that the profiles for the default timestep and 0.1 times the default timestep are virtually indistinguishable, with 10 times the default timestep performing resonably well, with small deviations at high densities.
+Meanwhile, at 50 times the default timestep the results are significantly distorted.
+Figure \ref{fig:comparison2} demonstrates the effect of decreasing the spatial resolution from $320^3$ to $64^3$.
+As expected, the lowest resolution performs poorly, whereas we see high convergence of the density profile at resolutions above $192^3$.
+
+\begin{figure}
+  \includegraphics[width=1.\textwidth,trim=0 0.8cm 0 1.2cm,clip]{comparison}
+  \caption{Effect of increasing the timestep on the density profile at half a revolution.}
+  \label{fig:comparison}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=1.\textwidth,trim=0 0.8cm 0 1.2cm,clip]{comparison2}
+  \caption{Effect of decreasing the spatial resolution on the density profile at half a revolution.}
+  \label{fig:comparison2}
 \end{figure}

### 2018-09-29 draft.tex
minor changes

@@
 
-\re{do we define $\langle \rangle$?
-Also -- should we explain in a sentence WHY the form of the equation changes when we go to period b.c.s?}\ek{I've given slightly more info here, think this is sufficient to understand the form of 2.8}
+\ek{I've given slightly more info here, think this is sufficient to understand the form of 2.8} \re{I think this is mostly there, but we should say which surface integral and may cite some of the literature here}
 
@@
 where $\psi(\vec{x},t_{i})$ is the field configuration at this halfway point in the full timestep.
-We explicitly set the $k=0$ Fourier mode to zero prior to the final inverse Fourier transform; as a consequence there is no need to subtract the global average density from the local value in Equation \ref{eq:schematic-phi}, in contrast to Equation \ref{eq:poisson_periodic}.\ek{Incorporated this explanation into main text} The final operation in equation \ref{eq:schematic-psi}  only changes the phase of  $\psi$, so we could replace $\psi(\vec{x},t_{i})$ with $\psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)$ in equation \ref{eq:schematic-phi} with no change in meaning.
+We explicitly set the $k=0$ Fourier mode to zero prior to the final inverse Fourier transform; as a consequence there is no need to subtract the global average density from the local value in Equation \ref{eq:schematic-phi}, in contrast to Equation \ref{eq:poisson_periodic}.
+The final operation in equation \ref{eq:schematic-psi}  only changes the phase of  $\psi$, so we could replace $\psi(\vec{x},t_{i})$ with $\psi(\vec{x},t\hspace*{-.2em}+\hspace*{-.2em}h)$ in equation \ref{eq:schematic-phi} with no change in meaning.
 \PyUltraLight makes an additional simplification to the symmetrised split-step Fourier method by combining the consecutive half-steps in the nonlinear term into a single full step.
@@
 The default timestep in  \PyUltraLight  is fixed with reference to the fluid interpretation of the Schr{\"o}dinger-Poisson system \cite{Hui:2016ltb}.
-The fluid interpretation is often used to recast the Schr{\"o}dinger-Poisson system in the form of the Madelung equations \cite{Suarez:2011yf}, such that the system can then be studied using the principles of hydrodynamics.\ek{Slightly more info and reference given here, though I don't want to go into too much detail as the specifics of the fluid representation aren't needed here.} The first step is to define:
+The fluid interpretation is often used to recast the Schr{\"o}dinger-Poisson system in the form of the Madelung equations \cite{Suarez:2011yf}, which a hydrodynamical representation of the system.
+The first step is to define
 \begin{equation}\label{eq:madelung}
@@
 \end{equation}
-From this definition of the fluid velocity, we see that if the phase change between two adjacent grid points exceeds $\pi$,\ek{not over a timestep, but at the same point in time} the fluid will appear to move backwards across the grid.
-Therefore, we set the default timestep, $\Delta t$, such that fluid travelling at this maximum velocity will traverse one grid space, $\Delta x$, per timestep, or
+and to treat $\vec{v}$ as a fluid velocity.
+From this perspective, if the phase difference between two adjacent grid points exceeds $\pi$ the fluid will appear to move ``backwards'' across the grid.
+We thus set the default timestep, $\Delta t$, so that fluid travelling at this maximum velocity traverses one grid space, $\Delta x$, per timestep, or
 \begin{equation}\label{eq:timestep}
@@
 \end{equation}
-This is a choice, rather than a strict constraint on the timestep.
-We motivate this choice of timestep by recognising that if the ``fluid'' approaches velocities where the phase appears to switch direction, the configuration will also be approaching the point where the simulation grid is too coarse to fully resolve the dynamics.
-Hence, making the timestep much smaller than this value may offer little practical advantage.
-Conversely, in some cases the breakdown may occur in regions of the simulation volume of little physical interest, and the user is free to choose a larger timestep  via the 'step\textunderscore factor' parameter in the Jupyter notebook.
+\re{Tweaked a little; changed phase change to phase difference; the latter sounds more explicitly spatial}
+This is a choice, rather than a strict constraint on $\Delta t$.
+However, if the ``fluid'' approaches velocities where the phase appears to switch direction, the configuration is approaching the point where the simulation grid is too coarse to fully resolve the dynamics.
+Hence, a timestep much smaller than this value may offer little practical advantage.
+However, in some cases the breakdown may occur in regions of the simulation volume that are of little physical interest, and the user is free to choose a larger timestep  via the 'step\textunderscore factor' parameter in the Jupyter notebook.
 
@@
   \caption{Comparison of theoretical and numerical density profiles at time of maximal interference for head-on collision of two solitons with mass ratio $\mu=2$ and no relative phase difference.
-The solitons were chosen with dimensionless masses 5 and 10, with an initial separation of 4 code units and relative velocity of 20 code units.
-The simulation was run at $256^3$ for a box of side-length 8 code units.}
+The solitons have dimensionless masses 5 and 10, with an initial separation of 4 code units and relative velocity of 20 code units.
+The simulation resolution is $256^3$ in a box of side-length 8.}
   \label{fig:interference}
@@
   \caption{Head-on collisions between solitons of mass 20, initial relative velocity 20, and initial separation 1.2 in code units.
-Contours of constant density are plotted to demonstrate the deformation of the soliton profiles as they converge.
-Time progresses from left to right across each row and is indicated in code units in each frame.
-The effective repulsive force generated by a $\pi$ phase shift can be seen in the lower panel.}
+Plots show contours of constant density; time progresses from left to right across each row and is indicated in code units for each frame.
+The upper panel shows solutons initially in phase; the effective repulsive force generated by a $\pi$ phase shift can be seen in the lower panel.}
   \label{fig:repulsion}
@@
   \includegraphics[width=1.\textwidth,trim=1cm 1cm 0 1cm,clip]{riemann}
-  \caption{Contours of constant density are shown for a solitonic core after one revolution around a central potential.
-The contours are superimposed upon a background representing the internal velocity field (where bulk motion has been subtracted).
-The velocity field is qualitatively that of an irrotational Riemann-S ellipsoid, and deformation of the density profile of the soliton along the radial direction (red arrow) can be seen.
+  \caption{Contours of constant density for a solitonic core after one revolution around a central potential.
+Contours are superimposed upon the internal velocity field (with the bulk motion subtracted).
+This velocity field is qualitatively that of an irrotational Riemann-S ellipsoid, and the deformation of the density profile of the soliton along the radial direction (red arrow) is visible.
 The host:satellite mass ratio is approximately 55; simulation resolution is $256^3$.}
@@
   \caption{Left: Evolution of the density profile of a solitonic core at equally spaced times as it undergoes tidal disruption in a potential centred at the red cross.
-Right: Evolution of the energy of the system, where times are indicated in correspondence with the snapshots of the density profile.
+Right: Evolution of the components of the total energy of the system; times  correspond to the labeled density profile snapshots.
 All quantities are in dimensionless code units.}
@@
   \includegraphics[width=0.9\textwidth,trim=0 0 2cm 2cm,clip]{egy_m=22.eps}
-  \caption{Evolution of the energy for a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}
+  \caption{Evolution of the components of the total energy of a binary soliton system with each soliton in an elliptical orbit around the common centre of mass.}
   \label{fig:binary}
@@
   \includegraphics[width=0.9\textwidth,trim=0 0.5cm 0 0,clip]{Total_Energy_Change.eps}
-  \caption{Demonstration of improving energy conservation with simulation resolution for a binary system of solitons orbiting their common centre of mass.
-$\Delta E_{total}$ is the difference between the current integrated energy at a given timestep and the initial integrated energy for the simulation.
-The ratio of this difference to the initial integrated energy is plotted on the y axis for different resolutions.}
+  \caption{Energy conservation as a function  of  simulation resolution for a binary solitons orbiting their common centre of mass.
+$\Delta E_{total}$ is the difference between the current total energy and the initial total energy for the configuration.
+The ratio of this difference to the initial integrated energy is plotted on the y axis for each resolution.}
   \label{fig:energy_change}
@@
   \includegraphics[width=0.9\textwidth,trim=0 0.5cm 0 0,clip]{Total_Energy_Change_Disruption.eps}
-  \caption{Demonstration of improving energy conservation with simulation resolution for a soliton undergoing significant tidal disruption in a Newtonian central potential.
-Again, $\Delta E_{total}$ is the difference between the current integrated energy at a given timestep and the initial integrated energy for the simulation.
+  \caption{Energy conservation as a function of simulation resolution for a soliton undergoing significant tidal disruption in a Newtonian central potential.
+$\Delta E_{total}$ is the difference between the current total energy and the initial total energy for the configuration.
 The ratio of this difference to the initial integrated energy is plotted on the y axis for each resolution.}
@@
 Two solitons of mass $m=20$ are initialised at radial distances $r=2$ from a central mass with $M=1000$ moving in opposite directions with initial speeds $\vert v\vert=\sqrt{M/r}$, corresponding to clockwise orbits around the central mass.
-The  box size is 10, while the total duration is 0.25 (all quantities in code units.
+The  box size is 10, while the total duration is 0.25 (all quantities in code units).
 Time runs from left to right.
-\re{you could actually zoom in, to only include the active region.}\ek{I'm just concerned about misrepesenting the size of the simulation grid - I wanted to accurately represent the buffer zone, because insufficient buffer will affect the dynamics, and people trying to reproduce these results might get confused.
-In any case, the visual details of the density distribution aren't important here, the purpose of the figure is just to give an idea of the setup} }
-  \label{fig:disruption}
+}  \label{fig:disruption}
 \end{figure}
@@
 We  sort the resulting values by  the density at the corresponding spatial location, and plot differences in the phase and the magnitude of $\psi$ relative to the values of the $320^3$ run as shown in Figure \ref{fig:spatial} (bottom).
-\re{maybe this could be one combined figure with both plots.}\ek{Done.} The convergence is poor at $64^3$, but improves with resolution, to the point that there is little difference between the  $256^3$ and $320^3$ cases.
+The convergence is poor at $64^3$, but improves with resolution, to the point that there is little difference between the  $256^3$ and $320^3$ cases.
 
 To examine the sensitivity of the the  $\psi$ field configuration to the timestep, we take the same default simulation at $256^3$, and then compare this to runs with timesteps 0.1, 10, and 50 times the default and down-sample the final output arrays to $64^3$.
-We sort sort the array values in order of the $\psi$ field magnitude in the run with the smallest timestep and in Figure \ref{fig:temporal} \re{maybe also combined into one full page figure?}\ek{Done.} we show the difference in the phase and magnitude of $\psi$ as a function of the timestep.
+We sort sort the array values in order of the $\psi$ field magnitude in the run with the smallest timestep and in Figure \ref{fig:temporal}  we show the difference in the phase and magnitude of $\psi$ as a function of the timestep.
 The difference between the results with the default timestep and a value 10 times smaller are negligible; and there is reasonable agreement between the default case and those with the timestep boosted by a factor of 10.
@@
   \includegraphics[width=1.\textwidth,trim=0 1cm 0 2cm,clip]{combined}
-  \caption{Top: Deviation of the phase of $\psi$ from the results of the simulation in which the timestep is set to 0.1 times the default.
-We see excellent convergence for the default timestep, and reasonable convergence for $<10$ times the default, with accuracy improving at high density regions.
-Bottom: Deviation of the magnitude of $\psi$ as compared to the simulation run with 0.1 times the default timestep.
-The convergence of the results for the default timestep with the results from the smaller timestep run demonstrate the suitability of the chosen default.
-We see relatively good convergence, particularly in high density regions, for a $<10$-fold increase above the default timestep.}
+  \caption{Top: Phase deviation of $\psi$, relative to solution with timestep 0.1 times default, sorted by the density.
+There is excellent agreement with the default timestep, and reasonable convergence at steps up to $10$ times the default, with better accuracy in high density regions.
+Bottom: Difference in magnitude of $\psi$, relative to the  solution with timestep 0.1 times default.
+Again we see good convergence with the default timestep, and tolerable agreement in high density regions when the step is a factor of $10$ or less than default.
+\re{Do we want to the put the ``key'' at the bottom not the middle; also I think the $x$-axis label could be clearer, or dropped.}}
   \label{fig:temporal}
@@
 
-\ek{added new content from here to the end of the section}
-To supplement our analysis of the dependence of the field configuration on spatial and temporal resolutions, we include Figures \ref{fig:comparison} and \ref{fig:comparison2}.
-These figures represent the density profiles down the axis of symmetry of the initial configuration (vertical axis in Figure \ref{fig:disruption}), however rather than the profiles at $t=0$, these are the profiles after approximately half a revolution around the central potential at t=0.28 code units.
-At this point, the two initial solitons have become distorted due to tidal forces, but have yet to become completely disrupted.
-Figure \ref{fig:comparison} demonstrates the effect of increasing the timestep from 0.1 times the default value to a maximum of 50 times the default value.
-We see that the profiles for the default timestep and 0.1 times the default timestep are virtually indistinguishable, with 10 times the default timestep performing resonably well, with small deviations at high densities.
-Meanwhile, at 50 times the default timestep the results are significantly distorted.
-Figure \ref{fig:comparison2} demonstrates the effect of decreasing the spatial resolution from $320^3$ to $64^3$.
-As expected, the lowest resolution performs poorly, whereas we see high convergence of the density profile at resolutions above $192^3$.
+Figures \ref{fig:comparison} and \ref{fig:comparison2} show profiles of the density through the simulation volume, as a function of spatial resolution and timestep.
+The figures represent the density profiles down the axis of symmetry of the initial configuration (vertical axis in Figure \ref{fig:disruption}) after approximately half a revolution around the central potential, or t=0.28 code units -- slightly after the final frame in Figure \ref{fig:disruption} - when the solitons have become distorted due to tidal forces, but are not yet completely disrupted.
+Figure \ref{fig:comparison} shows results obtained when the timestep is varied from 0.1 to 50 times the default value.
+The results with the default and the shorter timestep are virtually indistinguishable, and results are still reasonably accurate at 10 times the default timestep, with small deviations at high densities.
+However, the results are significantly distorted at 50 times the default timestep.
+Figure \ref{fig:comparison2} shows results at resolutions from $64^3$ to $320^3$.
+The lowest resolution performs poorly, but there is good convergence  at resolutions of $192^3$ and above.
+\re{minor edits; we could do the two figures as a single image}
 

### 2018-10-01 draft.tex
Latest revisions according to RE's comments

@@
 \usepackage{url}
+\usepackage{afterpage}
 \newcommand{\re}[1]{\textcolor{blue}{[{\bf RE}: #1]}}
@@
 where we subtract the average density from the right hand side of the Poisson equation.
-The form of Equation \ref{eq:poisson_periodic} is a consequence of Gauss' law and the fact that the surface integral of the gradient of the field is identically zero when periodic boundary conditions are imposed.
+The form of Equation \ref{eq:poisson_periodic} is a consequence of Gauss' law and the fact that the surface integral of the gradient of the field around the perimeter of the simulation grid is identically zero when periodic boundary conditions are imposed \cite{Dabo:2008}.
 
-\ek{I've given slightly more info here, think this is sufficient to understand the form of 2.8} \re{I think this is mostly there, but we should say which surface integral and may cite some of the literature here}
+\re{I think this is mostly there, but we should say which surface integral and may cite some of the literature here}\ek{How's this?
+It seems to be difficult to find many references which explain the subtraction of the average in detail, a lot of people just seem to use it without explanation.}
 
@@
 \end{equation}
-\re{Tweaked a little; changed phase change to phase difference; the latter sounds more explicitly spatial}
 This is a choice, rather than a strict constraint on $\Delta t$.
@@
 
+Figure \ref{fig:comparison} shows profiles of the density through the simulation volume, as a function of spatial resolution and timestep.
+Each plot represents the density profile down the axis of symmetry of the initial configuration (vertical axis in Figure \ref{fig:disruption}) after approximately half a revolution around the central potential, or t=0.28 code units -- slightly after the final frame in Figure \ref{fig:disruption} - when the solitons have become distorted due to tidal forces, but are not yet completely disrupted.
+We see that as the timestep is varied from 0.1 to 50 times the default value, the results with the default and the shorter timestep are virtually indistinguishable, and results are still reasonably accurate at 10 times the default timestep, with small deviations at high densities.
+However, the results are significantly distorted at 50 times the default timestep.
+We also see that as the spatial resolution is decreased from $320^3$ to $64^3$, the lowest resolution performs poorly, but there is good convergence  at resolutions of $192^3$ and above.
+\re{minor edits; we could do the two figures as a single image}\ek{How's this?}
+
 \begin{figure}
@@
 \end{figure}
+\clearpage
 
@@
 Again we see good convergence with the default timestep, and tolerable agreement in high density regions when the step is a factor of $10$ or less than default.
-\re{Do we want to the put the ``key'' at the bottom not the middle; also I think the $x$-axis label could be clearer, or dropped.}}
+\re{Do we want to the put the ``key'' at the bottom not the middle; also I think the $x$-axis label could be clearer, or dropped.}}\ek{How's this?}
   \label{fig:temporal}
 \end{figure}
-
-Figures \ref{fig:comparison} and \ref{fig:comparison2} show profiles of the density through the simulation volume, as a function of spatial resolution and timestep.
-The figures represent the density profiles down the axis of symmetry of the initial configuration (vertical axis in Figure \ref{fig:disruption}) after approximately half a revolution around the central potential, or t=0.28 code units -- slightly after the final frame in Figure \ref{fig:disruption} - when the solitons have become distorted due to tidal forces, but are not yet completely disrupted.
-Figure \ref{fig:comparison} shows results obtained when the timestep is varied from 0.1 to 50 times the default value.
-The results with the default and the shorter timestep are virtually indistinguishable, and results are still reasonably accurate at 10 times the default timestep, with small deviations at high densities.
-However, the results are significantly distorted at 50 times the default timestep.
-Figure \ref{fig:comparison2} shows results at resolutions from $64^3$ to $320^3$.
-The lowest resolution performs poorly, but there is good convergence  at resolutions of $192^3$ and above.
-\re{minor edits; we could do the two figures as a single image}
+\clearpage
 
 \begin{figure}
-  \includegraphics[width=1.\textwidth,trim=0 0.8cm 0 1.2cm,clip]{comparison}
-  \caption{Effect of increasing the timestep on the density profile at half a revolution.}
+  \includegraphics[width=1.\textwidth,trim=0 1.2cm 0 2.0cm,clip]{comparison}
+  \caption{Top: Effect of decreasing the spatial resolution on the density profile at half a revolution. Bottom: Effect of increasing the timestep on the density profile at half a revolution.}
   \label{fig:comparison}
 \end{figure}
-
-\begin{figure}
-  \includegraphics[width=1.\textwidth,trim=0 0.8cm 0 1.2cm,clip]{comparison2}
-  \caption{Effect of decreasing the spatial resolution on the density profile at half a revolution.}
-  \label{fig:comparison2}
-\end{figure}
+\clearpage
 

### 2018-10-03 draft.tex
removed comments from text

@@
 The form of Equation \ref{eq:poisson_periodic} is a consequence of Gauss' law and the fact that the surface integral of the gradient of the field around the perimeter of the simulation grid is identically zero when periodic boundary conditions are imposed \cite{Dabo:2008}.
-
-\re{I think this is mostly there, but we should say which surface integral and may cite some of the literature here}\ek{How's this?
-It seems to be difficult to find many references which explain the subtraction of the average in detail, a lot of people just seem to use it without explanation.}
 
@@
 We also see that as the spatial resolution is decreased from $320^3$ to $64^3$, the lowest resolution performs poorly, but there is good convergence  at resolutions of $192^3$ and above.
-\re{minor edits; we could do the two figures as a single image}\ek{How's this?}
 
@@
 Again we see good convergence with the default timestep, and tolerable agreement in high density regions when the step is a factor of $10$ or less than default.
-\re{Do we want to the put the ``key'' at the bottom not the middle; also I think the $x$-axis label could be clearer, or dropped.}}\ek{How's this?}
   \label{fig:temporal}
